### 2022-11-04 data_cleaning.Rmd
initial version of the plots. need to make the code cleaner and substract repetative parts into functions.

---
title: "Data cleaning"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Load data

```{r}
# raw data
raw <- readr::read_csv("data/ECAW_dataCut_221102.csv")
```

# Clean variable names

```{r}
raw <- janitor::clean_names(raw)
```

# Removing Qualtrics labels

```{r}
label <-
  raw%>% 
  slice(1:2)

raw <-
  raw %>% 
  slice(-(1:2))
```

# Creating a simple codebook

```{r}
codebook <-
  label %>% 
  slice(1) %>% 
  pivot_longer(cols = everything(), names_to = "original_var_name", values_to = "description") %>% 
  mutate(human_readable_var_name = case_when(
    original_var_name == "q3_1_1" ~ "typically_trustworthy",
    original_var_name == "q3_1_2" ~ "typically_reproducible",
    original_var_name == "q4_1_1" ~ "method_preregistered",
    original_var_name == "q4_1_2" ~ "method_blind",
    original_var_name == "q4_1_3" ~ "method_script",
    original_var_name == "q4_1_4" ~ "method_confirmatory",
    original_var_name == "q4_1_5" ~ "method_exploratory",
    original_var_name == "q8_1_1" ~ "ecaw_trustworthy",
    original_var_name == "q8_1_2" ~ "ecaw_reproducible",
    original_var_name == "q9_1_1" ~ "likert_less_willing",
    original_var_name == "q9_1_2" ~ "likert_opt_in",
    original_var_name == "q9_1_3" ~ "likert_study",
    original_var_name == "q9_1_4" ~ "likert_prefer_ecaw",
    original_var_name == "q10_3" ~ "contact",
    original_var_name == "q11_1_1" ~ "concerned",
    original_var_name == "q11_2" ~ "n_studies",
    original_var_name == "q11_3" ~ "programming_language",
    original_var_name == "q11_3_5_text" ~ "programming_language_other",
    TRUE ~ original_var_name
  ))

# Save the codebook
readr::write_tsv(codebook, "data/ECAW_raw_codebook.tsv")
```

# Variable renaming

```{r}
colnames(raw) <- dplyr::recode(
  colnames(raw), 
  !!!setNames(as.character(codebook$human_readable_var_name), codebook$original_var_name)
)
```

# Save clean daset

```{r}
readr::write_tsv(raw, "data/ECAW_raw_data.tsv")
```

### 2022-11-04 initial_report.Rmd
initial version of the plots. need to make the code cleaner and substract repetative parts into functions.

---
title: "Initial report"
output: html_document
date: '`r Sys.Date()`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Read data
raw <- readr::read_tsv("data/ECAW_raw_data.tsv")
```

# Reproducibility and trustworthiness

The reproducible and trustworthiness ratings with the current system.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 20}
typically_likert_data <-
  raw %>% 
    select(starts_with("typically_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("time", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
            "Strongly disagree",
            "Somewhat disagree",
            "Neither agree nor disagree",
            "Somewhat agree",
            "Strongly agree",
            "Missing",
            "I don't understand the question")
            ),
      type = str_to_title(type)
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

typically_missing_data <-
  typically_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

typically_likert_data %>% 
  likert_plot_odd(.,
                  title = "Typically, studies that analyze preexisting\nobservational datasets, are ...",
                  questions = type,
                  levels = scale,
                  n = n,
                  levels_to_drop = c("Missing", "I don't understand the question"),
                  positive_side = "lower",
                  text_push = 0.05) +
  geom_text(typically_missing_data, mapping = aes(x = type, y = 1.15, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 10, l = 0),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19)
    )
```

The reproducible and trustworthiness ratings of ECAW compared to the current system.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 20}
ecaw_likert_data <-
  raw %>% 
    select(starts_with("ecaw_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("time", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Much less",
          "Somewhat less",
          "About the same",
          "Somewhat more",
          "Much more",
          "Missing",
          "I don't understand the question")
            ),
      type = str_to_title(type)
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

ecaw_missing_data <-
  ecaw_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

ecaw_likert_data %>% 
  likert_plot_odd(.,
                  title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...",
                  questions = type,
                  levels = scale,
                  n = n,
                  levels_to_drop = c("Missing", "I don't understand the question"),
                  positive_side = "lower",
                  text_push = 0.05) +
  geom_text(typically_missing_data, mapping = aes(x = type, y = 1.15, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 10, l = 0),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19))
```

# ALSPAC

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 22, fig.height = 8}
alspac_likert_data <-
  raw %>% 
    select(starts_with("likert_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "type",
      values_to = "scale",
    ) %>%
    mutate(
      type = str_remove(type, "^[^_]*_")
    ) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree",
          "Missing",
          "I don't understand the question",
          "Unsure")
            ),
      type = case_when(
        type == "study" ~ "ALSPAC should\nrun a study",
        type == "prefer_ecaw" ~ "I would\nprefer ECAW",
        type == "opt_in" ~ "I would opt in",
        type == "less_willing" ~ "I would be\nless willing to use",
      )
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

alspac_missing_data <-
  alspac_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question", "Unsure")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing",
      scale == "Unsure" ~ "Unsure"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

alspac_likert_data %>% 
  likert_plot_odd(.,
                  title = "Thinking about a study you may run with ALSPAC data ...",
                  questions = type,
                  levels = scale,
                  n = n,
                  levels_to_drop = c("Missing", "I don't understand the question", "Unsure"),
                  positive_side = "lower",
                  text_push = 0.08,
                  limits = c(-1, 1)) +
  geom_text(alspac_missing_data, mapping = aes(x = type, y = 1.2, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 10, l = 0),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19))
```

# Current methods

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
method_data <-
  raw %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind analysis",
        type == "confirmatory" ~ "Confirmatory analysis",
        type == "exploratory" ~ "Exploratory analysis",
        type == "preregistered" ~ "Preregistration",
        type == "script" ~ "Analysis script",
      )
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

# method_missing_data <-
#   method_data %>% 
#   filter(scale %in% c("Missing", "I don't understand the question")) %>% 
#   mutate(
#     scale = case_when(
#       scale == "I don't understand the question" ~ "Don't understand",
#       scale == "Missing" ~ "Missing"),
#     text = paste0(scale, ": ", n)
#     ) %>% 
#   group_by(type) %>% 
#   summarise(text = stringr::str_c(text, collapse = "\n"))

method_data %>% 
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  labs(
    title = "The studies using preexisting observational data that I am involved in...",
    x = "Percentage",
    fill = "Response levels"
  ) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```

# How many respondents are open to  be interviewed?

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8}
raw %>% 
  replace_na(list(contact = "Missing")) %>% 
  count(contact) %>% 
  mutate(
    contact = case_when(
      str_detect(contact, "Yes.") ~ "Yes",
      TRUE ~ contact),
    contact = factor(contact, levels = c("Yes", "No", "Missing")),
    sum_n = sum(n),
    percentage = n / sum_n,
    text_label = paste0(n, "/", sum_n)
  ) %>%
  ggplot() +
  aes(
    x = contact,
    y = percentage
  ) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = text_label), size = 7, vjust = -1) +
  scale_y_continuous(expand = c(0, 0), labels = scales::label_percent(), limits = c(0, 0.55)) +
  labs(
    x = "May the research team contact you about potential focus groups?",
    y = "Percentage"
  ) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 15)
  )
```

# Concerned with rigour and reproducibility on the field

```{r, echo = FALSE, message = FALSE, warning = FALSE, , fig.width = 20, fig.height = 8}
concerned_data <-
  raw %>% 
  replace_na(list(concerned = "missing")) %>%
  count(concerned) %>% 
  mutate(
    concerned = tolower(concerned),
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned",
      "missing")
      )
  ) %>%
  tidyr::complete(concerned, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup() %>% 
  # Hacky need to fix later
  mutate(question = rep("", 8))

concerned_missing_data <-
  concerned_data %>% 
  filter(concerned == "missing") %>% 
  summarise(text = paste0(concerned, ": ", n))

concerned_data %>% 
  likert_plot_odd(.,
                  title = "Compared to what you think of as a typical researcher\nwho uses preexisting observational data in your field,\nhow concerned are you with research\n trustworthiness, bias, rigour, and reproducibility.",
                  questions = question,
                  levels = concerned,
                  n = n,
                  levels_to_drop = "missing",
                  positive_side = "lower",
                  text_push = 0.08,
                  limits = c(-0.5, 1),
                  legend_row = 3) +
  geom_text(concerned_missing_data, mapping = aes(x = 1, y = 1.1, label = text), size = 8) +
  theme(
    plot.margin = margin(t = 10, r = 220, b = 5, l = 50),
    legend.key.size = unit(0.8, 'cm'),
    legend.text = element_text(size = 19))
```

# Number of studies published

The median of the number of studies published is `r median(raw$n_studies, na.rm = T)` with an IQR of `r IQR(raw$n_studies, na.rm = T)`. The maximum number of studies published is `r max(raw$n_studies, na.rm = T)` while the minimum is `r min(raw$n_studies, na.rm = T)`. There were `r raw %>% filter(is.na(n_studies)) %>% nrow(.)` missing responses to this question.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
raw %>% 
  select(n_studies) %>% 
  filter(!is.na(n_studies)) %>%
  ggplot() +
  aes(
    x = n_studies
  ) +
  geom_histogram() +
  labs(
    x = "The number of studies published",
    y = "Count"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line()
  )
```

# Software used to analyse the data

There were `r raw %>% filter(is.na(programming_language)) %>% nrow(.)` missing responses to this question. The frequency of responses can be found below. Keep in mind that one respondent could choose multiple software languages.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
raw %>% 
  select(programming_language) %>% 
  filter(!is.na(programming_language)) %>% 
  separate_rows(programming_language, sep = ",") %>% 
  count(programming_language) %>% 
  rename(
    `Programming language` = programming_language,
    N = n) %>% 
  arrange(desc(N)) %>% 
  kableExtra::kbl()
```

### 2022-11-04 R/likert_plot_odd.R

@@ -1,0 +1,192 @@
+#' Horizontal stacked barplot for Likert-type data
+#' 
+#' The function creates a horizontal stacked barpolot for visualizing
+#' Likert-type data with an odd number of response levels. The function can
+#' take multiple scales and present them in one figure.
+#' 
+#' @section Source: The function is based on the code from \url{https://gist.github.com/trinker/0260a9dfdd9531f9b90d9fad2f7b4b12}
+#' 
+#' @param data dataframe containing the data for the plot
+#' @param title title of the plot
+#' @param levels name of the variable containing the levels. this variable should be a factor
+#' @param questions name of the variable that contains the questions to which the levels belong (multiple questions can be used)
+#' @param n name of the variable containing the number of responses corresponding to each level
+#' @param positive_side if `"lower"` lower levels than the middle are the positive side of the scale, while if `"upper"` levels bigger than the middle are the positive side
+#' @param levels_to_drop character vector containing the levels that should not be included e.g. "NA" or "I do not know" 
+#' @param limits the limits of the x axis
+#' @param text_push the distance between the x axis limits and the text along the y axis
+#' 
+#' @return The function returns a ggplot2 object.
+likert_plot_odd <- function(data, title, questions, levels, n, levels_to_drop, positive_side = c("lower", "upper"), limits = c(-0.5, 1), text_push = 0.1, legend_row = 1) {
+  .questions <- rlang::enquo(questions)
+  .levels <- rlang::enquo(levels)
+  
+  `%ni%` <- Negate(`%in%`)
+  
+  # Filter levels
+  likert_plot_data <-
+    data %>%
+    # Drop unwanted levels
+    filter({{levels}} %ni% levels_to_drop) %>% 
+    # Get levels as integer
+    mutate(levels_int = as.integer({{levels}}))
+  
+  # The number of levels should be odd
+  if (length(unique(likert_plot_data$levels_int)) %ni% c(3, 5, 7)) {
+    stop("The number of levels should 3, 5, or 7!")
+  }
+  
+  # Save color palette
+  if(length(unique(likert_plot_data$levels_int)) == 7) {
+    color_palette <- 
+      data.frame(
+        colors = c("#CD7F32", "#D5A16D", "#DDC3A9", "grey90", "#99D0CD", "#4EBCB5", "#03A89E"),
[…]

### 2022-11-04 initial_report.Rmd
updated analysis
added response rate and filtered for complete responses

@@
 library(viridis)
+library(scales) # fto display percentages
 
@@
 raw <- readr::read_tsv("data/ECAW_raw_data.tsv")
+# remove surveys where someone opened it, but didn't respond to any questions
+raw <- raw %>% filter(progress!=0)
+# number of respondents
+nResponse <- raw %>% nrow()
+
+# Filter for complete responses (I'm am maintaining the variable name "raw" only because it is used throughout the rest of the code)
+raw <- raw %>% filter(finished)
+#number of completed surveys
+nResponseComplete <- raw %>% nrow()
+
+invited <- 1148 # Number provided from Kyara at ALSPAC
+bounced <- 54 # Number provided from Kyara at ALSPAC
+received <- invited - bounced
 ```
+
+# Response rate
+
+We invited the ALSPAC mailing list to participate, which included `r invited` email addresses, of which `r bounced` bounced, leaving `r received` emails.
+The survey was completed `r nResponseComplete` times and partially completed `r nResponse - nResponseComplete` times.
+This gives a response rate for completed surveys of `r percent(nResponseComplete / received)` and a response rate for any questions `r percent(nResponse / received)`.
+
+The median time taken for complete survey responses was `r (median(raw$duration_in_seconds) / 60) %>% round(1)` minutes (IQR: `r (quantile(raw$duration_in_seconds, .25) / 60) %>% round(1)` - `r (quantile(raw$duration_in_seconds, .75) / 60) %>% round(1)`)
 
@@
 
-# ALSPAC
+# Current practices
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
+method_data <-
+  raw %>% 
+    select(starts_with("method_")) %>% 
+    pivot_longer(
+      cols = everything(),
+      names_to = "variable",
+      values_to = "scale",
+    ) %>%
+    separate(variable, into = c("name_prefix", "type")) %>% 
+    group_by(type) %>% 
+    count(scale) %>% 
+  ungroup() %>% 
+    replace_na(list(scale = "Missing")) %>% 
+    mutate(
+      scale = factor(scale, levels = c(
+          "Never or almost never",
+          "Sometimes",
+          "About half the time",
+          "Most of the time",
+          "Always or almost always",
+          "Missing",
+          "I don't understand the question")
+            ),
+      type = case_when(
+        type == "blind" ~ "Blind the data analyst",
+        type == "confirmatory" ~ "Contain confirmatory analysis",
+        type == "exploratory" ~ "Contain exploratory analysis",
+        type == "preregistered" ~ "Are preregistered",
+        type == "script" ~ "Share analysis scripts",
+      )
+      ) %>% 
+  group_by(type) %>%
+  tidyr::complete(scale, fill = list(n = 0)) %>% 
+  mutate(
+    percentage = n / sum(n)
+  ) %>% 
+  ungroup()
+
+# method_missing_data <-
+#   method_data %>% 
+#   filter(scale %in% c("Missing", "I don't understand the question")) %>% 
+#   mutate(
+#     scale = case_when(
+#       scale == "I don't understand the question" ~ "Don't understand",
+#       scale == "Missing" ~ "Missing"),
+#     text = paste0(scale, ": ", n)
+#     ) %>% 
+#   group_by(type) %>% 
+#   summarise(text = stringr::str_c(text, collapse = "\n"))
+
+method_data %>% 
+  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
+  ggplot() +
+  aes(
+    x = percentage,
+    y = type,
+    fill = scale
+  ) +
+  geom_bar(position = "stack", stat = "identity") +
+  scale_fill_viridis(option = "A", discrete = TRUE) +
+  scale_x_continuous(
+    limits = c(0, 1),
+    labels = scales::label_percent()
+  ) +
+  labs(
+    title = "The studies using preexisting observational data that I am involved in...",
+    x = "Percentage",
+    fill = "Response levels"
+  ) +
+  theme(
+    panel.background = element_blank(),
+    panel.grid = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.title.y = element_blank()
+  )
+```
+
+# ALSPAC and ECAWs
 
@@
     legend.text = element_text(size = 19))
-```
-
-# Current methods
-
-```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
-method_data <-
-  raw %>% 
-    select(starts_with("method_")) %>% 
-    pivot_longer(
-      cols = everything(),
-      names_to = "variable",
-      values_to = "scale",
-    ) %>%
-    separate(variable, into = c("name_prefix", "type")) %>% 
-    group_by(type) %>% 
-    count(scale) %>% 
-  ungroup() %>% 
-    replace_na(list(scale = "Missing")) %>% 
-    mutate(
-      scale = factor(scale, levels = c(
-          "Never or almost never",
-          "Sometimes",
-          "About half the time",
-          "Most of the time",
-          "Always or almost always",
-          "Missing",
-          "I don't understand the question")
-            ),
-      type = case_when(
-        type == "blind" ~ "Blind analysis",
-        type == "confirmatory" ~ "Confirmatory analysis",
-        type == "exploratory" ~ "Exploratory analysis",
-        type == "preregistered" ~ "Preregistration",
-        type == "script" ~ "Analysis script",
-      )
-      ) %>% 
-  group_by(type) %>%
-  tidyr::complete(scale, fill = list(n = 0)) %>% 
-  mutate(
-    percentage = n / sum(n)
-  ) %>% 
-  ungroup()
-
-# method_missing_data <-
-#   method_data %>% 
-#   filter(scale %in% c("Missing", "I don't understand the question")) %>% 
-#   mutate(
-#     scale = case_when(
-#       scale == "I don't understand the question" ~ "Don't understand",
-#       scale == "Missing" ~ "Missing"),
-#     text = paste0(scale, ": ", n)
-#     ) %>% 
-#   group_by(type) %>% 
-#   summarise(text = stringr::str_c(text, collapse = "\n"))
-
-method_data %>% 
-  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
-  ggplot() +
-  aes(
-    x = percentage,
-    y = type,
-    fill = scale
-  ) +
-  geom_bar(position = "stack", stat = "identity") +
-  scale_fill_viridis(option = "A", discrete = TRUE) +
-  scale_x_continuous(
-    limits = c(0, 1),
-    labels = scales::label_percent()
-  ) +
-  labs(
-    title = "The studies using preexisting observational data that I am involved in...",
-    x = "Percentage",
-    fill = "Response levels"
-  ) +
-  theme(
-    panel.background = element_blank(),
-    panel.grid = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.title.y = element_blank()
-  )
 ```

### 2022-12-07 manuscript/manuscript.Rmd
created basic project structure and draft analysis for manuscript

---
title             : "A survey of registration practices among observational researchers using preexisting datasets"
shorttitle        : "Explore and Confirm Analysis Workflow"
author:
  - name: Robert T. Thibault
    affiliation: '1,5'
    role:
      - Conceptualization
      - Data curation
      - Formal analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Supervision
      - Validation
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    corresponding: yes
    email: robert.thibault@stanford.edu
    address: Enter postal address here
  - name: Marton Kovacs
    affiliation: '2,6'
    role:
      - Data curation
      - Formal analysis
      - Software
      - Validation
      - Visualization
      - Writing - review & editing
    email: marton.balazs.kovacs@gmail.com
  - name: Tom E. Hardwicke
    affiliation: '3'
    role:
      - Methodology
      - Writing - review & editing
  - name: Alexandra Sarafoglou
    affiliation: '4'
    role:
      - Methodology
      - Writing - review & editing
  - name: John P. A. Ioannidis
    affiliation: '4'
    role:
      - Methodology
      - Writing - review & editing
  - name: Marcus R. Munafò
    affiliation: '1,7'
    role:
      - Conceptualization
      - Methodology
      - Supervision
      - Writing - review & editing
affiliation:
  - id: '1'
    institution: Meta-Research Innovation Center at Stanford (METRICS), Stanford University.
  - id: '2'
    institution: Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest,
      Hungary
  - id: '3'
    institution: Melbourne School of Psychological Sciences, University of Melbourne.
  - id: '4'
    institution: Department of Psychology, University of Amsterdam.
  - id: '5'
    institution: School of Psychological Science, University of Bristol.
  - id: '6'
    institution: Institute of Psychology, ELTE Eotvos Lorand University, Budapest,
      Hungary
  - id: '7'
    institution: Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for
      Transforming Biomedical Research, Berlin Institute of Health, Charité – Universitätsmedizin
      Berlin.
  - id: '8'
    institution: MRC Integrative Epidemiology Unit at the University of Bristol.
  - id: '9'
    institution: Departments of Medicine, Epidemiology and Population Health, Biomedical
      Data Science, and Statistics, Stanford University.
abstract: |
  placeholder for an abstract

keywords          : "keywords"
wordcount         : "X"
# bibliography      : "references.bib"
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, warning=FALSE, message=FALSE}
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
library(likert)
library(viridis)
# r_refs("references.bib")

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction
Many published research findings are non-reproducible and potentially false or misleading (Camerer et al., 2018; Errington et al., 2021; Ioannidis, 2005, 2008; Open Science Collaboration, 2015). Researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices can lead to research waste, and useless or even harmful healthcare and policy interventions (e.g., Prasad & Cifu, 2019). To mitigate these issues, clinical trials research often blinds both participants and experimenters, and prospectively registers outcome measures. Observational research using preexisting data, however, is rarely registered and the researchers analyzing the data generally have access to complete datasets without any blinding imposed on them.

In this survey study, we seeked to gather data on (1) the opinions and practices of observational researchers regarding rigour and reproducibility, and (2) their thoughts on the use of hold-out samples or cross validation for analyzing preexisting datasets. We hope to use this data to refine a potential intervention aimed to improve research quality and to determine whether observational researchers find such an intervention acceptable.

In this intervention we propose, a data management organization would first provide only a subset of the data a researcher requests. After the researcher prepares an analysis script based on the subset of data and uploads it to a registry—where it will be openly accessible and permanent—the data management organization then provides the full dataset and the researcher can proceed as they wish. The exact implementation (e.g., whether the data management organization runs quality checks on the analysis scripts) would depend on the preferences of the data management organization and the research community using their dataset.

We call this process an Explore and Confirm Analysis Workflow (ECAW). Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly. ECAWs could serve as an alternative to typical preregistrations for observational research on preexisting data and may be more precise and comprehensive. Whereas typical preregistrations are written descriptions of a planned analysis and obstacles may arise in their implementation, ECAWs provide an analysis script that has been shown to run. ECAWs would not solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration. A version of this workflow has been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al., 2020).

Moreover, a recent study recruited 120 teams to analyze a single observational dataset and had half the teams preregister an analysis plan and the other half prepare an analysis plan by writing an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded) (Sarafoglou et al., 2022). This blinded condition is comparable to ECAWs in that the researchers have access to enough data to develop an informed analysis, but without compromising the confirmatory nature of a final analysis on the complete dataset. The researchers found that the two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.

We are interested in exploring real-world applications of these types of blind data analysis and hold-out samples (e.g., ECAWs) to improve research rigour and reproducibility. One example of such an initiative is OpenSAFELY (www.opensafely.org). This platform provides a dataset of simulated health records from which researchers can develop an analysis script. When ready, the researchers submit their analysis script which is run in a Trusted Research Environment (TRE) and automatically logged and made public. Thus, this workflow keeps health records anonymous while also documenting all analyses that are run on the real data.

In the present study, we solicited structured feedback from observational researchers and found that respondents generally agreed that there could be a benefit to ECAWs and that they are worth testing.

# Methods
We ran a descriptive and exploratory survey study. We had no hypotheses, but we did have two specific objectives. (1) To gain insights on the opinions and practices of researchers who already use preexisting observational datasets, in regards to the trustworthiness and reproducibility of research. (2) To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices. This objective includes assessing and refining potential interventions—such as ECAWs—and assessing their acceptability.

We sent an email (supplementary material A) to invite researchers on the mailing list for the Avon Longitudinal Study of Parents and Children (ALSPAC) to participate in an online survey. ALSPAC is “a transgenerational prospective observational study investigating influences on health and development across the life course. It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.” (Boyd et al., 2013). Thus, our survey will reach researchers in diverse disciplines that use observational data. The survey was open from 10 Oct 2022 to 1 Nov 2022. We sent two reminder emails, exactly one week and two weeks after the original email invitation.

The survey (supplementary material B) contained 6 blocks. Block 1 assessed whether respondents agree that observational research using preexisting data is trustworthy and reproducible (2 questions). Block 2 asked respondents about their own research practices, including preregistration, blinding, and sharing analysis scripts (5 questions). Between Block 2 and Block 3, the survey described ECAWs. Block 3 assessed whether respondents agree that ECAWs would make observational research using preexisting data more trustworthy and reproducible (2 questions). Block 4 directly asked respondents whether ALSPAC should run a study on ECAWs and whether they would participate (5 questions). Block 5 contained open-ended questions to garner information that the survey could otherwise not capture (3 questions). Block 6 collected information about the respondents. (4 questions). Open ended questions were analyzed….[add explanation]

# Results
## Participants

```{r include=FALSE}
# List of used programming languages and their counts text
language_table <-
  processed %>% 
  rename(lang = programming_language) %>% 
  separate_rows(lang, sep = ",") %>% 
  count(lang) %>% 
  filter(!is.na(lang)) %>% 
  arrange(desc(n)) %>% 
  mutate(lang_n = glue::glue("{lang} (n = {n})")) %>% 
  pull(lang_n) %>% 
  glue::glue_collapse(., sep = ", ", last = ", and ")

# Calculate concerned percentage
concerned <- support_percentage(
  processed,
  concerned,
  c(
    "very much less concerned",
    "less concerned",
    "somewhat less concerned",
    "as concerned as a typical researcher in my field",
    "somewhat more concerned",
    "more concerned",
    "very much more concerned"
    )
  )
```

We invited the ALSPAC mailing list to participate, which included 1148 email addresses. 54 emails bounced, leaving 1094 emails that went through. The survey was completed `r nrow(processed)` times and partially completed `r nrow(filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for completed surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The results presented in this manuscript only include completed surveys.[^2] The median time taken for complete survey responses was `r median(processed$duration_in_mins)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .75), 1)` - `r round(quantile(processed$duration_in_mins, .25), 1)`).

Respondents published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .75), 1)` - `r round(quantile(processed$n_studies, .25), 1)`) studies using preexisting observational data (Figure S1). They reported using the programming languages `r language_text` (Table S1). `r pull(filter(concerned, support == "positive"), percentage)`% of participants reported being more concerned, versus `r pull(filter(concerned, support == "negative"), percentage)`% who reported being less concerned, with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure #).

```{r, include=FALSE}
# Concerned figure data preparation
concerned_plot_data <-
  processed %>% 
  select(concerned) %>% 
  mutate(
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned"
      )
    )
  )

# Check distinct values in data
distinct(concerned_plot_data, concerned)

# Create missing values text
concerned_missing_text <- 
  concerned_plot_data %>% 
  filter(is.na(concerned)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

# Create likert package data not including the missing values
concerned_plot_data <- likert(filter(concerned_plot_data, !is.na(concerned)))

# Create figure
concerned_plot <- 
  plot(concerned_plot_data) +
  theme(
    axis.text.y = element_blank()
  )
```

```{r echo=FALSE}
concerned_plot
```

[^1]: This mailing list has been around for 30 years and may contain email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years. Thus, the response rate may be greater if we are only including active researchers.
[^2]: Results that include incomplete surveys are available in the Supplementary Material.
[^3]: The survey defined trustworthy as: “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust.”
[^4]: The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.”

## Survey results

```{r include=FALSE}
# Calculate typical trustworthiness and reproducibility ratings
typically_trustworthy <- support_percentage(
  processed,
  typically_trustworthy,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  "I don't understand the question"
  )

typically_reproducible <- support_percentage(
  processed,
  typically_reproducible,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  "I don't understand the question"
  )

# Calculate ecaw trustworthiness and reproducibility ratings
ecaw_trustworthy <- support_percentage(
  processed,
  ecaw_trustworthy,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"),
  "I don't understand the question"
  )

ecaw_reproducible <- support_percentage(
  processed,
  ecaw_reproducible,
  c(
    "Much less",
    "Somewhat less",
    "About the same",
    "Somewhat more",
    "Much more"),
  "I don't understand the question"
  )
```

```{r include=FALSE}
# Create typically trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
processed %>% 
  select(contains("typically_")) %>% 
  dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>% 
  nrow()

# Prepare plot data
typically_plot_data <- 
  processed %>%
  select(contains("typically_")) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"))
    )) %>% 
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

# Create likert package data not including the missing values
typically_plot_data <- likert(typically_plot_data)

# Create figure
typically_plot <- 
  plot(typically_plot_data)

# Create ecaw trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
ecaw_missing_text <-
  processed %>% 
  select(contains("ecaw_")) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
    ) %>% 
  dplyr::filter(is.na(value) | value == "I don't understand the question") %>% 
  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
  replace_na(list(value = "Missing")) %>% 
  group_by(item , value) %>% 
  summarise(
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(text = paste(value, "=", n))

# Prepare plot data
ecaw_plot_data <- 
  processed %>%
  select(contains("ecaw_")) %>% 
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . == "I don't understand the question" ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
          "Much less",
          "Somewhat less",
          "About the same",
          "Somewhat more",
          "Much more"))
    )) %>% 
  rename(
    Reproducible = ecaw_reproducible,
    Trustworthy = ecaw_trustworthy
  )

# Create likert package data not including the missing values
ecaw_plot_data <- likert(ecaw_plot_data)

# Create figure
ecaw_plot <- 
  plot(ecaw_plot_data)
```

Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%) (Figure 1A). At the same time, many agreed that a study using an ECAW would be more trustworthy[^3] (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%) and more reproducible[^4] (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%) compared to a typical study using preexisting observational data (Figure 1B).

```{r}
typically_plot
```

```{r}
ecaw_plot
```

```{r include=FALSE}
confirmatory <- level_percentage(
  processed,
  method_confirmatory,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  ) %>% 
  filter(method_confirmatory != "Never or almost never") %>% 
  summarise(sum_percentage = sum(percentage)) %>% 
  pull(sum_percentage)

exploratory <- level_percentage(
  processed,
  method_exploratory,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  ) %>% 
  dplyr::filter(method_exploratory != "Never or almost never") %>% 
  summarise(sum_percentage = sum(percentage)) %>% 
  pull(sum_percentage)

preregistered <- level_percentage(
  processed,
  method_preregistered,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  )

blind <- level_percentage(
  processed,
  method_blind,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  )

script <- level_percentage(
  processed,
  method_script,
  c(
    "Never or almost never",
    "Sometimes",
    "About half the time",
    "Most of the time",
    "Always or almost always"),
  "I don't understand the question"
  )
```

Almost all respondents answered that they use both exploratory (`r exploratory`%) and confirmatory (`r confirmatory`%) analyses at least sometimes (Figure 2A-B). About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%) (Figure 2C). Over half reported preregistering their studies never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%) (Figure 2D). `r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% reported that they never or almost never blind the data analyst (Figure 2E).

```{r}
# Used methods figures
method_plot_data <-
  processed %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind analysis",
        type == "confirmatory" ~ "Confirmatory analysis",
        type == "exploratory" ~ "Exploratory analysis",
        type == "preregistered" ~ "Preregistration",
        type == "script" ~ "Analysis script",
      )
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

# method_missing_data <-
#   method_data %>% 
#   filter(scale %in% c("Missing", "I don't understand the question")) %>% 
#   mutate(
#     scale = case_when(
#       scale == "I don't understand the question" ~ "Don't understand",
#       scale == "Missing" ~ "Missing"),
#     text = paste0(scale, ": ", n)
#     ) %>% 
#   group_by(type) %>% 
#   summarise(text = stringr::str_c(text, collapse = "\n"))

method_plot <-
  method_plot_data %>% 
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  labs(
    x = "Percentage",
    fill = "Response levels"
  ) +
  papaja::theme_apa() +
  theme(
    axis.title.y = element_blank()
  ) 
```

```{r echo=FALSE}
method_plot
```

```{r include=FALSE}
willing <- support_percentage(
  processed,
  alspac_less_willing,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )

opt <- support_percentage(
  processed,
  alspac_opt_in,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )

study <- support_percentage(
  processed,
  alspac_study,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )

prefer <- support_percentage(
  processed,
  alspac_prefer_ecaw,
  c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"),
  c(
    "I don't understand the question",
    "Unsure"
    )
  )
```

`r pull(filter(willing, support == "positive"), percentage)`% of respondents agreed (versus `r pull(filter(willing, support == "negative"), percentage)`% who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3). `r pull(filter(opt, support == "positive"), percentage)`% agreed (`r pull(filter(opt, support == "negative"), percentage)`% disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. `r pull(filter(study, support == "positive"), percentage)`% agreed (`r pull(filter(study, support == "negative"), percentage)`% disagreed) that ALSPAC should run a study on ECAWs. `r pull(filter(prefer, support == "positive"), percentage)`% agreed (`r pull(filter(prefer, support == "negative"), percentage)`% disagreed) that they would prefer using an ECAW than using typical preregistration.

```{r}
# Check if there is any missing and don't understand responses separately
alspac_missing_text <-
  processed %>% 
  select(contains("alspac_")) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
    ) %>% 
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>% 
  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
  replace_na(list(value = "Missing")) %>% 
  group_by(item , value) %>% 
  summarise(
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(text = paste(value, "=", n))

# Prepare plot data
alspac_plot_data <- 
  processed %>%
  select(contains("alspac_")) %>% 
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree")
          )
      )
    ) %>% 
  rename(
    `ALSPAC should\nrun a study` = alspac_study,
    `I would\nprefer ECAW` = alspac_prefer_ecaw,
    `I would opt in` = alspac_opt_in,
    `I would be\nless willing to use` = alspac_less_willing
  )

# Create likert package data not including the missing values
alspac_plot_data <- likert(alspac_plot_data)

# Create figure
alspac_plot <- 
  plot(alspac_plot_data)
```

```{r}
alspac_plot
```

Address the concerns from the open-ended questions [transfer to right side of table 1]. For some research questions, the subset method may not work. In which case, a larger sample could be provided, or better yet, a synthetic dataset could be provided. The time burden is likely more than when not preregistering, but efforts can be made to minimize it. One study comparing ~120 analyses found that an ECAW-like workflow took similar effort to preregistration (although that was for a relatively clean dataset—whereas prereg may take longer if the dataset is a mess; or viceversa). The concern that analysis would need to change after writing an ECAW is not a big deal—having the ECAW written could provide evidence that the changes were small. This is often raised as an issue in preregistraiton as well, but is acceptable or encouraged when it makes sense to do so. Finally the issue that ECAW would not work for situations where researchers are reanalyzing a dataset they already have is true, and it simply would not be possible in this situation (in which case, the researcher would not be requesting that data again anyhow, so it’s beside the point).

## Exploratory analyses

# Discussion
Are ECAWs applicable 1? Yes. In short, ECAWs could be implemented in many cases. Many respondents use software that would allow for the ECAW format (R, ?Stata?), although some use software where this would be more difficult (SPSS, ?SAS?). Most have experience publishing relevant articles, so are familiar with the process. And most respondents also run studies that contain confirmatory analyses.

Are ECAWs applicable 2? Yes. Current methods to improve rigour and reproducibility are underused (sharing analysis scripts, blinding data analysts, preregisteration).

How biased are these results? Many that are earlier in their career based on # of publications. Also many are more concerned than others about reproducibility. So this may, unsurprisingly, be a somewhat biased sample. Although the 10% response rate was quite good.

Should we ALSPAC (or another org) run a study? Probably.

How would ECAWs be implemented? Would ALSPAC or the org check whether they run. Where would the be uploaded. Would the primary outcomes need to be clearly identified in the output. These would depend on the organization running such an experiment / workflow.

# Acknowledgements

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2022-12-07 preprocessing/initial_report.Rmd
created basic project structure and draft analysis for manuscript

--- 2022-11-04 initial_report.Rmd
@@
 ---
+TODO: subset by concerned number of studies and current methods
+```{r setup, include=FALSE}
+# General rmarkdown rendering setting
+knitr::opts_chunk$set(echo = TRUE)
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+# Loading used packages 
 library(tidyverse)
+library(likert)
 library(viridis)
@@
 ```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 20}
+# Prepearing data for Likert-scale plotting
 typically_likert_data <-
@@
 
+# Creating text labels for reporting the count of non-likert response levels on the Likert-scale plot 
 typically_missing_data <-
@@
 ```{r, echo = FALSE, message = FALSE, warning = FALSE}
+# Separate zero and bins of 5
 raw %>% 
@@
   ) +
-  geom_histogram() +
+  geom_histogram(binwidth = 3) +
   labs(
@@
   scale_y_continuous(expand = c(0, 0)) +
+  scale_x_continuous(
+    # breaks = seq(0, 100, by = 3)
+    )
   theme(

### 2023-03-14 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# A survey of registration practices among observational researchers using preexisting datasets

<!-- badges: start -->
<!-- badges: end -->

The repository contains the data and the analysis for the research paper titled "A survey of registration practices among observational researchers using preexisting datasets". Other materials belonging to the project can be found on the project's OSF page at [link](link).

## Folder structure

### 2023-03-14 manuscript/interactive_report.Rmd

---
title: "Results exploration"
author: "Marton Kovacs"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
library(likert)
library(viridis)
library(shiny)
# r_refs("references.bib")

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

what is this document...

# Subset

Subset data for all the results.

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
               min = min(processed$n_studies, na.rm = T), max = max(processed$n_studies, na.rm = T), value = max(processed$n_studies, na.rm = T), step = 1),
  radioButtons("methods", label = "Used the following methods:", choices = c("Blind", "Script", "Confirmatory", "Exploratory"), selected = NULL, inline = FALSE),
    radioButtons("concerned", label = "Concerned with rigour and reproducibility:", choices = c("Negative", "Neutral", "Support"), selected = NULL, inline = FALSE)
)

# Filter data
filtered_data <- reactive({
  processed %>% 
    filter(
      n_studies == input$n_studies
    )
  })
```

# Reproducibility and trustworthiness

The reproducible and trustworthiness ratings with the current ALSPAC system.

```{r}
# Prepare plot data
typically_plot_data <- 
  processed %>%
  select(contains("typically_")) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"))
    )) %>% 
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

# Create likert package data not including the missing values
typically_plot_data <- likert(typically_plot_data)

# Create figure
typically_plot <- 
  plot(typically_plot_data)
```

### 2023-03-14 manuscript/supplementary_materials.Rmd

---
title             : "Supplementary materials: A survey of registration practices among observational researchers using preexisting datasets"
shorttitle        : "Supplementary materials"

keywords          : "keywords"
wordcount         : "X"
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library("papaja")
library(tidyverse)

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Table

```{r}
# programming languages table
processed %>% 
  rename(lang = programming_language) %>% 
  separate_rows(lang, sep = ",") %>% 
  count(lang) %>% 
  replace_na(list(lang = "Missing")) %>% 
  arrange(desc(n)) %>% 
  rename(
    `Programming language` = lang,
    `Number of occurance` = n
  ) %>% 
  kbl(
  .,
  format = "latex",
  col.names = c("Programming language", "Number of occurance"),
  booktabs = T,
  escape = F,
  caption = "Programming Languages Used to Analyse Preexisting Observational Data",
  centering = T,
  position = "H") %>%
  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
  row_spec(row = 0, bold = TRUE) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
```

# Figure

```{r}
# Create text for missing responses
n_studies_missing_text <- 
  processed %>% 
  filter(is.na(n_studies)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

# Create histogram of the number of observational studies done
processed %>% 
  # Exclude missing responses
  filter(!is.na(n_studies)) %>% 
  ggplot() +
  aes(x = n_studies) +
  geom_histogram() +
  annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
  labs(
    x = "Number of observational studies done prior using ALSPAC",
    y = "Count"
  ) +
  papaja::theme_apa()
```

### 2023-03-14 preprocessing/ecaw_raw_processed_preprocessing.Rmd

---
title: "Data preprocessing: raw data to processed data"
author: "Marton Kovacs"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load used packages
library(tidyverse)
library(here)
library(writexl)

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)
```

# Load raw data

```{r}
raw <- readr::read_csv(here("data/raw/ecaw_raw_data.csv"))
```

Before any exclusions we had `r nrow(raw)` responses in the dataset.

# Exclusions
## Test data

The survey was sent out on 2022/10/10 to the participants. Thus, any responses in the dataset before that data are coming from the testing of the survey. These responses will be excluded from further analysis. We excluded `r raw %>% dplyr::filter(lubridate::as_date(start_date) < lubridate::as_date("2022-10-10")) %>% nrow()` test responses.

```{r}
raw <-
  raw %>% 
  dplyr::filter(lubridate::as_date(start_date) >= lubridate::as_date("2022-10-10"))
```

## Partial responses

We only include full responses in the final analysis as preregistered (see the full text of the preregistration [HERE](link)). There were `r dplyr::filter(raw, !finished)` respondents who did not finish the survey.

```{r}
raw <- 
  raw %>% 
  dplyr::filter(finished)
```

At the end of the exclusion we had `r nrow(raw)` responses remaining for the analysis.

It was optional to answer all of the questions in our survey, so it is possible to have missing responses even if the respondent finished the survey. We will keep these responses.

```{r}
raw <-
  raw %>% 
    # dropping those questions that were open ended or conditional on other responses or irrelevant
    # flag those responses that has any missing values
    mutate(
      has_missing = case_when(
        if_any(
          c(
            -start_date,
            -end_date,
            -status,
            -progress,
            -duration_in_seconds,
            -finished,
            -recorded_date,
            -distribution_channel,
            -user_language,
            -programming_language_other,
            -unsure_explain,
            -drawback_ecaw,
            -suggestions,
            -comments
            ),
          .fns = is.na) ~ 1L,
        TRUE ~ 0L
        )
      )
```

There were `r filter(raw, has_missing) %>% nrow()` respondents who had at least one missing value in the main questions of interests.

# Distribution of completition time

On median it took `r median(raw$duration_in_seconds) / 60` to finish the survey to the participants in hour sample.

```{r}
# We will only need the duration in minutes for the main analysis so we calculate it here
raw <-
  raw %>% 
  mutate(duration_in_mins = duration_in_seconds / 60)

raw %>% 
  ggplot() +
  aes(x = duration_in_mins) +
  geom_histogram() +
  labs(
    x = "Response duration in minutes",
    y = "Count"
  )
```

# Delete not used variables

```{r}
raw <-
  raw %>% 
  select(
    - start_date,
    - end_date,
    - status,
    - progress,
    - finished,
    - duration_in_seconds,
    - recorded_date,
    - distribution_channel,
    - user_language
  )
```

# Exploring the number of responses to open-ended questions

The survey contained 4 open-ended questions (*unsure_explain*, *drawback_ecaw*, *suggestions*, *comments*) that we analyzed qualitatively as preregistered (see page X line Y of the preregistration).

Here, we calculate the number of responses for each open ended question.

```{r}
raw %>% 
  select(unsure_explain, drawback_ecaw, suggestions, comments) %>% 
  # convert to long format
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "response",
  ) %>% 
  # flag missing responses
  mutate(is_na = is.na(response)) %>%
  # count missing responses per variables
  count(variable, is_na) %>% 
  # keep only complete responses
  filter(!is_na)
```

# Categorizing free-text responses

In some cases respondents had the option to provide free-text responses in case they choose the _Other_ option in a multiple-choice question. We will categorize these responses manually.

```{r}
# prepearing free-text responses for manual rating
# this part of the code is commented out not to overwrite the manual grouping
# raw %>%
#   select(response_id, programming_language_other) %>%
#   dplyr::filter(!is.na(programming_language_other)) %>%
#   mutate(group_mk = NA_character_) %>%
#   rename(programming_language_to_group = programming_language_other) %>%
#   write_xlsx(., here("data/raw/ecaw_free_text_grouping_data.xlsx"))

# processing grouped responses
raw <-
  # joining free-text responses to the data table
  left_join(
    raw,
    # reading manually grouped free-text responses
    readxl::read_xlsx(here("data/raw/ecaw_free_text_grouping_data.xlsx")),
    by = "response_id"
  ) %>% 
  # replacing "Other (please describe)" text in response with grouped response
  mutate(
    programming_language  = stringr::str_replace(programming_language, "Other \\(please describe\\)", group_mk)
  ) %>% 
  # drop not needed variables
  select(
    - programming_language_other,
    - programming_language_to_group,
    - group_mk)
```

# Creating a simple codebook

For creating the codebook we will use the codebook created for the source data.

```{r}
codebook <-
  read_csv(here("data/source/ecaw_221102_source_codebook.csv")) %>% 
  select(-original_var_name) %>% 
  rename(var_name = new_var_name) %>% 
  filter(var_name %in% colnames(raw))

# Save the codebook
readr::write_csv(codebook, here("data/raw/ecaw_raw_codebook.csv"))
```

# Save processed dataset

```{r}
readr::write_csv(raw, here("data/processed/ecaw_processed_data.csv"))
```

### 2023-03-14 preprocessing/ecaw_source_raw_preprocessing.Rmd

---
title: "Data preprocessing: source data to raw data"
author: "Marton Kovacs"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

# Load source data

```{r}
source <- readr::read_csv(here("data/source/ecaw_221102_source_data.csv"))
```

# Clean variable names

```{r}
source <- janitor::clean_names(source)
```

# Removing Qualtrics labels

```{r}
label <-
  source %>% 
  slice(1:2)

source <-
  source %>% 
  slice(-(1:2))
```

# Creating a simple codebook

```{r}
codebook <-
  label %>% 
  slice(1) %>% 
  pivot_longer(cols = everything(), names_to = "original_var_name", values_to = "description") %>% 
  mutate(
    new_var_name = case_when(
      original_var_name == "q3_1_1"       ~ "typically_trustworthy",
      original_var_name == "q3_1_2"       ~ "typically_reproducible",
      original_var_name == "q4_1_1"       ~ "method_preregistered",
      original_var_name == "q4_1_2"       ~ "method_blind",
      original_var_name == "q4_1_3"       ~ "method_script",
      original_var_name == "q4_1_4"       ~ "method_confirmatory",
      original_var_name == "q4_1_5"       ~ "method_exploratory",
      original_var_name == "q8_1_1"       ~ "ecaw_trustworthy",
      original_var_name == "q8_1_2"       ~ "ecaw_reproducible",
      original_var_name == "q9_1_1"       ~ "alspac_less_willing",
      original_var_name == "q9_1_2"       ~ "alspac_opt_in",
      original_var_name == "q9_1_3"       ~ "alspac_study",
      original_var_name == "q9_1_4"       ~ "alspac_prefer_ecaw",
      original_var_name == "q9_2"         ~ "unsure_explain",
      original_var_name == "q10_1"        ~ "drawback_ecaw",
      original_var_name == "q10_2"        ~ "suggestions",
      original_var_name == "q10_3"        ~ "contact",
      original_var_name == "q11_1_1"      ~ "concerned",
      original_var_name == "q11_2"        ~ "n_studies",
      original_var_name == "q11_3"        ~ "programming_language",
      original_var_name == "q11_3_5_text" ~ "programming_language_other",
      original_var_name == "q11_4"        ~ "comments",
      TRUE ~ original_var_name
      )
    )

# Save the codebook
readr::write_csv(codebook, here("data/source/ecaw_221102_source_codebook.csv"))
```

# Variable renaming

We are assigning the new variable names to the source data table using the codebook.

```{r}
colnames(source) <- dplyr::recode(
  colnames(source), 
  !!!setNames(as.character(codebook$new_var_name), codebook$original_var_name)
)
```

# Save raw dataset

```{r}
readr::write_csv(source, here("data/raw/ecaw_raw_data.csv"))
```

### 2023-03-14 manuscript/manuscript.Rmd
initial draft of the manuscript with the results

@@
 ```{r setup, include = FALSE, warning=FALSE, message=FALSE}
+# Load pacakges
 library("papaja")
@@
 # List of used programming languages and their counts text
-language_table <-
+language_text <-
   processed %>% 
@@
 54 emails bounced, leaving 1094 emails that went through.
-The survey was completed `r nrow(processed)` times and partially completed `r nrow(filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for completed surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The results presented in this manuscript only include completed surveys.[^2] The median time taken for complete survey responses was `r median(processed$duration_in_mins)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .75), 1)` - `r round(quantile(processed$duration_in_mins, .25), 1)`).
+The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for completed surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The results presented in this manuscript only include completed surveys.[^2] The median time taken for complete survey responses was `r median(processed$duration_in_mins)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .75, na.rm = TRUE), 1)` - `r round(quantile(processed$duration_in_mins, .25, na.rm = T), 1)`).
 
-Respondents published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .75), 1)` - `r round(quantile(processed$n_studies, .25), 1)`) studies using preexisting observational data (Figure S1).
+Respondents published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)` - `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)`) studies using preexisting observational data (Figure S1).
 They reported using the programming languages `r language_text` (Table S1).
@@
   mutate(
+    concerned = tolower(concerned),
     concerned = factor(concerned, levels = c(
@@
 
-```{r include=FALSE}
+```{r include=FALSE, warning=FALSE, message=FALSE}
 willing <- support_percentage(
@@
 
-```{r}
+```{r warning=FALSE, message=FALSE}
 # Check if there is any missing and don't understand responses separately

### 2023-03-14 preprocessing/ecaw_raw_processed_preprocessing.Rmd
initial draft of the manuscript with the results

@@
             ),
-          .fns = is.na) ~ 1L,
-        TRUE ~ 0L
+          .fns = is.na) ~ TRUE,
+        TRUE ~ FALSE
         )

### 2023-03-14 manuscript/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A survey of registration practices among observational researchers using preexisting datasets},
  pdfauthor={Robert T. Thibault1,5, Marton Kovacs2,6, Tom E. Hardwicke3, Alexandra Sarafoglou4, John P. A. Ioannidis4, \& Marcus R. Munafò1,7},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{A survey of registration practices among observational researchers using preexisting datasets}
\author{Robert T. Thibault\textsuperscript{1,5}, Marton Kovacs\textsuperscript{2,6}, Tom E. Hardwicke\textsuperscript{3}, Alexandra Sarafoglou\textsuperscript{4}, John P. A. Ioannidis\textsuperscript{4}, \& Marcus R. Munafò\textsuperscript{1,7}}
\date{}

\shorttitle{Explore and Confirm Analysis Workflow}

\authornote{

The authors made the following contributions. Robert T. Thibault: Conceptualization, Data curation, Formal analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Supervision, Validation, Visualization, Writing - original draft, Writing - review \& editing; Marton Kovacs: Data curation, Formal analysis, Software, Validation, Visualization, Writing - review \& editing; Tom E. Hardwicke: Methodology, Writing - review \& editing; Alexandra Sarafoglou: Methodology, Writing - review \& editing; John P. A. Ioannidis: Methodology, Writing - review \& editing; Marcus R. Munafò: Conceptualization, Methodology, Supervision, Writing - review \& editing.

Correspondence concerning this article should be addressed to Robert T. Thibault, Enter postal address here. E-mail: \href{mailto:robert.thibault@stanford.edu}{\nolinkurl{robert.thibault@stanford.edu}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Meta-Research Innovation Center at Stanford (METRICS), Stanford University.\\\textsuperscript{2} Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{3} Melbourne School of Psychological Sciences, University of Melbourne.\\\textsuperscript{4} Department of Psychology, University of Amsterdam.\\\textsuperscript{5} School of Psychological Science, University of Bristol.\\\textsuperscript{6} Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{7} Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Berlin Institute of Health, Charité -- Universitätsmedizin Berlin.\\\textsuperscript{8} MRC Integrative Epidemiology Unit at the University of Bristol.\\\textsuperscript{9} Departments of Medicine, Epidemiology and Population Health, Biomedical Data Science, and Statistics, Stanford University.}

\abstract{%
placeholder for an abstract
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Many published research findings are non-reproducible and potentially false or misleading (Camerer et al., 2018; Errington et al., 2021; Ioannidis, 2005, 2008; Open Science Collaboration, 2015). Researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices can lead to research waste, and useless or even harmful healthcare and policy interventions (e.g., Prasad \& Cifu, 2019). To mitigate these issues, clinical trials research often blinds both participants and experimenters, and prospectively registers outcome measures. Observational research using preexisting data, however, is rarely registered and the researchers analyzing the data generally have access to complete datasets without any blinding imposed on them.

In this survey study, we seeked to gather data on (1) the opinions and practices of observational researchers regarding rigour and reproducibility, and (2) their thoughts on the use of hold-out samples or cross validation for analyzing preexisting datasets. We hope to use this data to refine a potential intervention aimed to improve research quality and to determine whether observational researchers find such an intervention acceptable.

In this intervention we propose, a data management organization would first provide only a subset of the data a researcher requests. After the researcher prepares an analysis script based on the subset of data and uploads it to a registry---where it will be openly accessible and permanent---the data management organization then provides the full dataset and the researcher can proceed as they wish. The exact implementation (e.g., whether the data management organization runs quality checks on the analysis scripts) would depend on the preferences of the data management organization and the research community using their dataset.

We call this process an Explore and Confirm Analysis Workflow (ECAW). Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly. ECAWs could serve as an alternative to typical preregistrations for observational research on preexisting data and may be more precise and comprehensive. Whereas typical preregistrations are written descriptions of a planned analysis and obstacles may arise in their implementation, ECAWs provide an analysis script that has been shown to run. ECAWs would not solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration. A version of this workflow has been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al., 2020).

Moreover, a recent study recruited 120 teams to analyze a single observational dataset and had half the teams preregister an analysis plan and the other half prepare an analysis plan by writing an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded) (Sarafoglou et al., 2022). This blinded condition is comparable to ECAWs in that the researchers have access to enough data to develop an informed analysis, but without compromising the confirmatory nature of a final analysis on the complete dataset. The researchers found that the two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.

We are interested in exploring real-world applications of these types of blind data analysis and hold-out samples (e.g., ECAWs) to improve research rigour and reproducibility. One example of such an initiative is OpenSAFELY (www.opensafely.org). This platform provides a dataset of simulated health records from which researchers can develop an analysis script. When ready, the researchers submit their analysis script which is run in a Trusted Research Environment (TRE) and automatically logged and made public. Thus, this workflow keeps health records anonymous while also documenting all analyses that are run on the real data.

In the present study, we solicited structured feedback from observational researchers and found that respondents generally agreed that there could be a benefit to ECAWs and that they are worth testing.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We ran a descriptive and exploratory survey study. We had no hypotheses, but we did have two specific objectives. (1) To gain insights on the opinions and practices of researchers who already use preexisting observational datasets, in regards to the trustworthiness and reproducibility of research. (2) To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices. This objective includes assessing and refining potential interventions---such as ECAWs---and assessing their acceptability.

We sent an email (supplementary material A) to invite researchers on the mailing list for the Avon Longitudinal Study of Parents and Children (ALSPAC) to participate in an online survey. ALSPAC is ``a transgenerational prospective observational study investigating influences on health and development across the life course. It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.'' (Boyd et al., 2013). Thus, our survey will reach researchers in diverse disciplines that use observational data. The survey was open from 10 Oct 2022 to 1 Nov 2022. We sent two reminder emails, exactly one week and two weeks after the original email invitation.

The survey (supplementary material B) contained 6 blocks. Block 1 assessed whether respondents agree that observational research using preexisting data is trustworthy and reproducible (2 questions). Block 2 asked respondents about their own research practices, including preregistration, blinding, and sharing analysis scripts (5 questions). Between Block 2 and Block 3, the survey described ECAWs. Block 3 assessed whether respondents agree that ECAWs would make observational research using preexisting data more trustworthy and reproducible (2 questions). Block 4 directly asked respondents whether ALSPAC should run a study on ECAWs and whether they would participate (5 questions). Block 5 contained open-ended questions to garner information that the survey could otherwise not capture (3 questions). Block 6 collected information about the respondents. (4 questions). Open ended questions were analyzed\ldots.{[}add explanation{]}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We invited the ALSPAC mailing list to participate, which included 1148 email addresses. 54 emails bounced, leaving 1094 emails that went through. The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for completed surveys and 2\% for incomplete surveys.\footnote{This mailing list has been around for 30 years and may contain email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn't been active in research for 30 years. Thus, the response rate may be greater if we are only including active researchers.} The results presented in this manuscript only include completed surveys.\footnote{Results that include incomplete surveys are available in the Supplementary Material.} The median time taken for complete survey responses was 7.22 minutes (IQR: 13.10 - 4.30).

Respondents published a median of NA (IQR 25 - 2) studies using preexisting observational data (Figure S1). They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), irrelevant (n = 1), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1). 52\% of participants reported being more concerned, versus 6\% who reported being less concerned, with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure \#).

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}

\hypertarget{survey-results}{%
\subsection{Survey results}\label{survey-results}}

Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy (72\%) and reproducible (79\%) (Figure 1A). At the same time, many agreed that a study using an ECAW would be more trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust.''} (70\%) and more reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (69\%) compared to a typical study using preexisting observational data (Figure 1B).

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}

Almost all respondents answered that they use both exploratory (95\%) and confirmatory (92\%) analyses at least sometimes (Figure 2A-B). About half reported sharing their analysis scripts never or almost never (20\%), or sometimes (32\%) (Figure 2C). Over half reported preregistering their studies never or almost never (38\%), or sometimes (27\%) (Figure 2D). 79\% reported that they never or almost never blind the data analyst (Figure 2E).

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}

25\% of respondents agreed (versus 44\% who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3). 53\% agreed (20\% disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. 55\% agreed (10\% disagreed) that ALSPAC should run a study on ECAWs. 45\% agreed (22\% disagreed) that they would prefer using an ECAW than using typical preregistration.

\begin{verbatim}
## Warning in likert(alspac_plot_data): Variable I would be
## less willing to use is not a factor.
\end{verbatim}

\begin{verbatim}
## Warning in likert(alspac_plot_data): Variable I would opt in is not a factor.
\end{verbatim}

\begin{verbatim}
## Warning in likert(alspac_plot_data): Variable ALSPAC should
## run a study is not a factor.
\end{verbatim}

\begin{verbatim}
## Warning in likert(alspac_plot_data): Variable I would
## prefer ECAW is not a factor.
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}

Address the concerns from the open-ended questions {[}transfer to right side of table 1{]}. For some research questions, the subset method may not work. In which case, a larger sample could be provided, or better yet, a synthetic dataset could be provided. The time burden is likely more than when not preregistering, but efforts can be made to minimize it. One study comparing \textasciitilde120 analyses found that an ECAW-like workflow took similar effort to preregistration (although that was for a relatively clean dataset---whereas prereg may take longer if the dataset is a mess; or viceversa). The concern that analysis would need to change after writing an ECAW is not a big deal---having the ECAW written could provide evidence that the changes were small. This is often raised as an issue in preregistraiton as well, but is acceptable or encouraged when it makes sense to do so. Finally the issue that ECAW would not work for situations where researchers are reanalyzing a dataset they already have is true, and it simply would not be possible in this situation (in which case, the researcher would not be requesting that data again anyhow, so it's beside the point).

\hypertarget{exploratory-analyses}{%
\subsection{Exploratory analyses}\label{exploratory-analyses}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Are ECAWs applicable 1? Yes. In short, ECAWs could be implemented in many cases. Many respondents use software that would allow for the ECAW format (R, ?Stata?), although some use software where this would be more difficult (SPSS, ?SAS?). Most have experience publishing relevant articles, so are familiar with the process. And most respondents also run studies that contain confirmatory analyses.

Are ECAWs applicable 2? Yes. Current methods to improve rigour and reproducibility are underused (sharing analysis scripts, blinding data analysts, preregisteration).

How biased are these results? Many that are earlier in their career based on \# of publications. Also many are more concerned than others about reproducibility. So this may, unsurprisingly, be a somewhat biased sample. Although the 10\% response rate was quite good.

Should we ALSPAC (or another org) run a study? Probably.

How would ECAWs be implemented? Would ALSPAC or the org check whether they run. Where would the be uploaded. Would the primary outcomes need to be clearly identified in the output. These would depend on the organization running such an experiment / workflow.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}

\end{document}

### 2023-03-14 manuscript/supplementary_materials.Rmd

@@
   aes(x = n_studies) +
-  geom_histogram() +
-  annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
+  geom_histogram(binwidth = 3) +
+  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
+  labs(
+    x = "Number of observational studies done prior using ALSPAC",
+    y = "Count"
+  ) +
+  papaja::theme_apa()
+
+valami <- 
+  processed %>% 
+  # Exclude missing responses
+  filter(!is.na(n_studies))
+
+ggplot() +
+  geom_histogram(
+    data = dplyr::filter(valami, n_studies == 0),
+    mapping = aes(x = n_studies), binwidth = 1, color = 'red') +
+  geom_histogram(
+    data = dplyr::filter(valami, n_studies != 0),
+    mapping = aes(x = n_studies), binwidth = 5) +
+  # annotate(geom = "text", label = n_studies_missing_text, y = 30) +
   scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +

### 2023-03-17 README.Rmd
updated figures, analysis, and readme based on Robbie's comments

@@
 
-<!-- badges: start -->
-<!-- badges: end -->
-
 The repository contains the data and the analysis for the research paper titled "A survey of registration practices among observational researchers using preexisting datasets".
-Other materials belonging to the project can be found on the project's OSF page at [link](link).
+Other materials belonging to the project can be found on the project's OSF page at [https://osf.io/x2dpn/](https://osf.io/x2dpn/).
 
 ## Folder structure
+
+The `data/` folder contains all the datafiles and the corresponding codebooks that are needed to reproduce the results of the project.
+
+The `preprocessing/` folder contains all the data preprocessing files in rmarkdowns. Within this folder you can find the following files:
+
+  * `ecaw_source_raw_preprocessing.Rmd` file contains the code necessary for the transformation of the source data (the datafile downloaded directly from Qualtrics as is) to the raw datafile (datafile with standard naming).
+We also made sure that the raw datafile does not contain any information that can be used to indetify any of the respondents.
+  * `ecaw_raw_processed_preprocessing.Rmd` file contains the code that cleans the dataset and transforms is into tidy format ready for the analysis.
+
+The `manuscript/` folder contains the manuscript and supplementary materials documents with the analysis code embedded in them.
+
+The `R/` folder contains any custom R functions and their documentation.

### 2023-03-17 manuscript/interactive_report.Rmd
updated figures, analysis, and readme based on Robbie's comments

@@
 library(shiny)
+library(plotly)
 # r_refs("references.bib")
@@
   sliderInput("n_studies", label = "Number of studies using ALSPAC:",
-               min = min(processed$n_studies, na.rm = T), max = max(processed$n_studies, na.rm = T), value = max(processed$n_studies, na.rm = T), step = 1),
-  radioButtons("methods", label = "Used the following methods:", choices = c("Blind", "Script", "Confirmatory", "Exploratory"), selected = NULL, inline = FALSE),
-    radioButtons("concerned", label = "Concerned with rigour and reproducibility:", choices = c("Negative", "Neutral", "Support"), selected = NULL, inline = FALSE)
+              min = min(processed$n_studies, na.rm = T),
+              max = max(processed$n_studies, na.rm = T), value = c(0, max(processed$n_studies, na.rm = T)), step = 1),
+  selectInput("methods", label = "Used the following methods:",
+              choices = c(
+                "Blind" = "method_blind",
+                "Script" = "method_script",
+                "Preregistration" = "method_preregistered",
+                "Confirmatory" = "method_confirmatory", 
+                "Exploratory" = "method_exploratory"),
+              selected = NULL, multiple = TRUE),
+  selectInput("concerned", label = "Level of concern with rigour and reproducibility:",
+              choices = c(
+                "Less concerned",
+                "Neutral",
+                "More concerned"),
+              selected = c(
+                "Less concerned",
+                "Neutral",
+                "More concerned"), multiple = TRUE),
+  selectInput("language", label = "Programming language used for the analysis:",
+              choices = c(
+                "R",
+                "Stata",
+                "SPSS",
+                "SAS",
+                "Python",
+                "Mplus",
+                "Bash",
+                "MATLAB",
+                "Nextflow",
+                "plink2",
+                "irrelevant"
+              ),
+              selected = c(
+                "R",
+                "Stata",
+                "SPSS",
+                "SAS",
+                "Python",
+                "Mplus",
+                "Bash",
+                "MATLAB",
+                "Nextflow",
+                "plink2",
+                "irrelevant"
+              ), multiple = TRUE)
   )
@@
 filtered_data <- reactive({
+  concerned_options <-list(
+    "Less concerned" =  c("very much less concerned", "less concerned", "somewhat less concerned"),
+    "Neutral" =  c("as concerned as a typical researcher in my field"),
+    "More concerned" = c("somewhat more concerned", "more concerned", "Very much more concerned")
+  )
+
+  concerned_filter <- unlist(concerned_options[input$concerned], use.names = FALSE)
+
   processed %>% 
-    filter(
-      n_studies == input$n_studies
-    )
+    filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
+    filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
+    filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
+    filter(str_detect(programming_language, paste(input$language, collapse="|")) | is.na(programming_language))
   })
+# select(filtered_data(), concerned)
+# shiny::renderTable(filtered_data())
+shiny::renderText({nrow(filtered_data())})
 ```
 
-# Reproducibility and trustworthiness
-
-The reproducible and trustworthiness ratings with the current ALSPAC system.
-
-```{r}
+```{r echo=FALSE, warning=FALSE}
+# Create typically trustworthy and reproducible plot
 # Prepare plot data
-typically_plot_data <- 
-  processed %>%
+typically_plot_data <- reactive({
+  filtered_data() %>%
     select(contains("typically_")) %>%
@@
       Trustworthy = typically_trustworthy
-  )
-
+      ) %>%
     # Create likert package data not including the missing values
-typically_plot_data <- likert(typically_plot_data)
+    likert(.)
+})
 
 # Create figure
-typically_plot <- 
-  plot(typically_plot_data)
+typically_plot <- reactive({
+  plot(typically_plot_data(), digits = 1, ordered = FALSE, legend.position = "right") +
+  labs(title = "Typically, studies that analyze preexisting observational datasets (such as the ALSPAC dataset) are...")
+})
+
+# Create ecaw trustworthy and reproducible plot
+# Prepare plot data
+ecaw_plot_data <- reactive({
+  filtered_data() %>%
+  select(contains("ecaw_")) %>%
+  # Transform not wanted response values to NA
+  # likert::likert drops NA values silently
+  # When var transformed to factor these values would be transformed to NA
+  # Automatically but I try to be explicit
+  mutate(
+    across(
+      everything(),
+      ~ case_when(
+        . == "I don't understand the question" ~ NA_character_,
+        TRUE ~ .
+      )
+    )
+  ) %>%
+  mutate(
+    across(
+      .fns = ~ factor(., levels =   c(
+          "Much less",
+          "Somewhat less",
+          "About the same",
+          "Somewhat more",
+          "Much more"))
+    )) %>%
+  rename(
+    Reproducible = ecaw_reproducible,
+    Trustworthy = ecaw_trustworthy
+  ) %>%
+    # Create likert package data not including the missing values
+    likert(.)
+})
+
+# Create figure
+ecaw_plot <- reactive({
+  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right") +
+  labs(title = "Compared to a typical study using preexisting observational data, a study using an ECAW would be...")
+})
+
+# Join the two plots together to make one figure
+renderPlot({ typically_plot + ecaw_plot + plot_layout(ncol = 1, nrow = 2) })
 ```
+
+```{r}
+renderPlot({
+# Create histogram of the number of observational studies done
+filtered_data() %>% 
+  # Exclude missing responses
+  filter(!is.na(n_studies)) %>% 
+  ggplot() +
+  aes(x = n_studies) +
+  geom_histogram(binwidth = 3) +
+  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
+  labs(
+    x = "Number of studies",
+    y = "Count",
+    title = "Approximately how many studies have you published using a preexisting observational dataset (e.g., the ALSPAC dataset)?"
+  ) +
+  papaja::theme_apa()
+})
+```

### 2023-03-17 manuscript/manuscript.Rmd
updated figures, analysis, and readme based on Robbie's comments

@@
 # bibliography      : "references.bib"
-floatsintext      : no
+floatsintext      : yes
 linenumbers       : yes
@@
 ```{r setup, include = FALSE, warning=FALSE, message=FALSE}
-# Load pacakges
 library("papaja")
@@
 library(patchwork)
+# The developer version is needed
 library(likert)
@@
 # r_refs("references.bib")
+```
 
+```{r custom_scripts}
 # Source R scripts
@@
 purrr::walk(r_scripts, source)
+```
 
-# Loading data
+```{r preprocessing}
+# loads source data, performs preprocessing, saves raw datafiles
+# source(here("preprocessing", "ecaw_source_raw_preprocessing.R"))
+
+# loads raw data, performs preprocessing, saves processed datafiles
+# source(here("preprocessing", "ecaw_raw_processed_preprocessing.R"))
+```
+
+```{r loading_data}
 processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
@@
 
-```{r analysis-preferences}
+```{r analysis_preferences}
 # Seed for random number generation
-set.seed(42)
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
@@
 # Introduction
-Many published research findings are non-reproducible and potentially false or misleading (Camerer et al., 2018; Errington et al., 2021; Ioannidis, 2005, 2008; Open Science Collaboration, 2015).
-Researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices can lead to research waste, and useless or even harmful healthcare and policy interventions (e.g., Prasad & Cifu, 2019).
-To mitigate these issues, clinical trials research often blinds both participants and experimenters, and prospectively registers outcome measures.
-Observational research using preexisting data, however, is rarely registered and the researchers analyzing the data generally have access to complete datasets without any blinding imposed on them.
-
-In this survey study, we seeked to gather data on (1) the opinions and practices of observational researchers regarding rigour and reproducibility, and (2) their thoughts on the use of hold-out samples or cross validation for analyzing preexisting datasets.
-We hope to use this data to refine a potential intervention aimed to improve research quality and to determine whether observational researchers find such an intervention acceptable.
-
-In this intervention we propose, a data management organization would first provide only a subset of the data a researcher requests.
-After the researcher prepares an analysis script based on the subset of data and uploads it to a registry—where it will be openly accessible and permanent—the data management organization then provides the full dataset and the researcher can proceed as they wish.
-The exact implementation (e.g., whether the data management organization runs quality checks on the analysis scripts) would depend on the preferences of the data management organization and the research community using their dataset.
-
-We call this process an Explore and Confirm Analysis Workflow (ECAW).
-Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly.
-ECAWs could serve as an alternative to typical preregistrations for observational research on preexisting data and may be more precise and comprehensive.
-Whereas typical preregistrations are written descriptions of a planned analysis and obstacles may arise in their implementation, ECAWs provide an analysis script that has been shown to run.
-ECAWs would not solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration.
-A version of this workflow has been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al., 2020).
-
-Moreover, a recent study recruited 120 teams to analyze a single observational dataset and had half the teams preregister an analysis plan and the other half prepare an analysis plan by writing an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded) (Sarafoglou et al., 2022).
-This blinded condition is comparable to ECAWs in that the researchers have access to enough data to develop an informed analysis, but without compromising the confirmatory nature of a final analysis on the complete dataset.
-The researchers found that the two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.
-
-We are interested in exploring real-world applications of these types of blind data analysis and hold-out samples (e.g., ECAWs) to improve research rigour and reproducibility.
-One example of such an initiative is OpenSAFELY (www.opensafely.org).
-This platform provides a dataset of simulated health records from which researchers can develop an analysis script.
-When ready, the researchers submit their analysis script which is run in a Trusted Research Environment (TRE) and automatically logged and made public.
-Thus, this workflow keeps health records anonymous while also documenting all analyses that are run on the real data.
-
-In the present study, we solicited structured feedback from observational researchers and found that respondents generally agreed that there could be a benefit to ECAWs and that they are worth testing.
 
 # Methods
-We ran a descriptive and exploratory survey study.
-We had no hypotheses, but we did have two specific objectives.
-(1) To gain insights on the opinions and practices of researchers who already use preexisting observational datasets, in regards to the trustworthiness and reproducibility of research.
-(2) To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices.
-This objective includes assessing and refining potential interventions—such as ECAWs—and assessing their acceptability.
-
-We sent an email (supplementary material A) to invite researchers on the mailing list for the Avon Longitudinal Study of Parents and Children (ALSPAC) to participate in an online survey.
-ALSPAC is “a transgenerational prospective observational study investigating influences on health and development across the life course.
-It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.” (Boyd et al., 2013).
-Thus, our survey will reach researchers in diverse disciplines that use observational data.
-The survey was open from 10 Oct 2022 to 1 Nov 2022.
-We sent two reminder emails, exactly one week and two weeks after the original email invitation.
-
-The survey (supplementary material B) contained 6 blocks.
-Block 1 assessed whether respondents agree that observational research using preexisting data is trustworthy and reproducible (2 questions).
-Block 2 asked respondents about their own research practices, including preregistration, blinding, and sharing analysis scripts (5 questions).
-Between Block 2 and Block 3, the survey described ECAWs.
-Block 3 assessed whether respondents agree that ECAWs would make observational research using preexisting data more trustworthy and reproducible (2 questions).
-Block 4 directly asked respondents whether ALSPAC should run a study on ECAWs and whether they would participate (5 questions).
-Block 5 contained open-ended questions to garner information that the survey could otherwise not capture (3 questions).
-Block 6 collected information about the respondents.
-(4 questions).
-Open ended questions were analyzed….[add explanation]
 
@@
   arrange(desc(n)) %>% 
+  # Removing irrelevant
+  filter(lang != "irrelevant") %>% 
   mutate(lang_n = glue::glue("{lang} (n = {n})")) %>% 
@@
     "more concerned",
-    "very much more concerned"
+    # This level label starts with a capital letter
+    "Very much more concerned"
     )
@@
 54 emails bounced, leaving 1094 emails that went through.
-The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for completed surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The results presented in this manuscript only include completed surveys.[^2] The median time taken for complete survey responses was `r median(processed$duration_in_mins)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .75, na.rm = TRUE), 1)` - `r round(quantile(processed$duration_in_mins, .25, na.rm = T), 1)`).
+The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r median(processed$duration_in_mins)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
 
-Respondents published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)` - `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)`) studies using preexisting observational data (Figure S1).
-They reported using the programming languages `r language_text` (Table S1).
-`r pull(filter(concerned, support == "positive"), percentage)`% of participants reported being more concerned, versus `r pull(filter(concerned, support == "negative"), percentage)`% who reported being less concerned, with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure #).
+Respondents published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)`) studies using preexisting observational data (Figure S1).
+They reported using the programming languages `r language_text` (Table S1)[^2].
+`r pull(filter(concerned, support == "positive"), percentage)`% of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); `r pull(filter(concerned, support == "negative"), percentage)`% reported being less concerned.
 
-```{r, include=FALSE}
-# Concerned figure data preparation
-concerned_plot_data <-
-  processed %>% 
-  select(concerned) %>% 
-  mutate(
-    concerned = tolower(concerned),
-    concerned = factor(concerned, levels = c(
-      "very much less concerned",
-      "less concerned",
-      "somewhat less concerned",
-      "as concerned as a typical researcher in my field",
-      "somewhat more concerned",
-      "more concerned",
-      "very much more concerned"
-      )
-    )
-  )
-
-# Check distinct values in data
-distinct(concerned_plot_data, concerned)
-
-# Create missing values text
-concerned_missing_text <- 
-  concerned_plot_data %>% 
-  filter(is.na(concerned)) %>% 
-  summarise(
-    n = n(),
-    missing_text = paste0("missing = ", n)
-  ) %>% 
-  pull(missing_text)
-
-# Create likert package data not including the missing values
-concerned_plot_data <- likert(filter(concerned_plot_data, !is.na(concerned)))
-
-# Create figure
-concerned_plot <- 
-  plot(concerned_plot_data) +
-  theme(
-    axis.text.y = element_blank()
-  )
-```
-
-```{r echo=FALSE}
-concerned_plot
-```
-
-[^1]: This mailing list has been around for 30 years and may contain email addresses that are no longer monitored.
+[^1]: The ALSPAC mailing list has been active for >30 years and may contain email addresses that are no longer monitored.
 For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years.
-Thus, the response rate may be greater if we are only including active researchers.
-[^2]: Results that include incomplete surveys are available in the Supplementary Material.
+Excluding these email addresses would increase the response rate, but we do not know by how much.
+[^2]: Participants could select multiple responses to this survey question.
 [^3]: The survey defined trustworthy as: “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust.”
@@
 
-```{r include=FALSE}
+```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=6}
 # Create typically trustworthy and reproducible plot
 # Check if there is any missing and don't understand responses separately
-processed %>% 
-  select(contains("typically_")) %>% 
-  dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
-  nrow()
+# processed %>% 
+#   select(contains("typically_")) %>% 
+#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
+#   nrow()
 
@@
 typically_plot <- 
-  plot(typically_plot_data)
+  plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.7) +
+  labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    legend.text = element_text(size = 12),
+    legend.title = element_blank()
+  )
 
 # Create ecaw trustworthy and reproducible plot
-# Check if there is any missing and don't understand responses separately
+# Calculate missing responses
 ecaw_missing_text <-
@@
 ecaw_plot <- 
-  plot(ecaw_plot_data)
+  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.7) +
+  labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    legend.text = element_text(size = 12),
+    legend.title = element_blank()
+  )
+
+# Join the two plots together to make one figure
+typically_plot + ecaw_plot + plot_layout(ncol = 1, nrow = 2)
 ```
 
-Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%) (Figure 1A).
-At the same time, many agreed that a study using an ECAW would be more trustworthy[^3] (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%) and more reproducible[^4] (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%) compared to a typical study using preexisting observational data (Figure 1B).
-
-```{r}
-typically_plot
-```
-
-```{r}
-ecaw_plot
-```
+Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy[^3] (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%) and reproducible[^4] (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%) (Figure 1A).
+At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%) compared to a typical study using preexisting observational data (Figure 1B).
 
@@
 
-Almost all respondents answered that they use both exploratory (`r exploratory`%) and confirmatory (`r confirmatory`%) analyses at least sometimes (Figure 2A-B).
-About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%) (Figure 2C).
-Over half reported preregistering their studies never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%) (Figure 2D).
-`r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% reported that they never or almost never blind the data analyst (Figure 2E).
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%) (Figure 2A).
+About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%) (Figure 2B).
+`r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% reported that they never or almost never blind the data analyst (Figure 2C).
+Almost all respondents answered that they use both exploratory (`r exploratory`%) and confirmatory (`r confirmatory`%) analyses at least sometimes (Figure 2D-E).
 
-```{r}
+```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4.5}
 # Used methods figures
@@
       type = case_when(
-        type == "blind" ~ "Blind analysis",
-        type == "confirmatory" ~ "Confirmatory analysis",
-        type == "exploratory" ~ "Exploratory analysis",
-        type == "preregistered" ~ "Preregistration",
-        type == "script" ~ "Analysis script",
-      )
+        type == "blind" ~ "Blind the data analyst",
+        type == "confirmatory" ~ "Contain confirmatory analysis",
+        type == "exploratory" ~ "Contain exploratory analysis",
+        type == "preregistered" ~ "Are preregistered",
+        type == "script" ~ "Share analysis scripts",
+      ),
+      type = factor(type, levels = c(
+        "Contain exploratory analysis",
+        "Contain confirmatory analysis",
+        "Blind the data analyst",
+        "Share analysis scripts",
+        "Are preregistered"
+      ))
       ) %>% 
@@
 
-method_plot <-
 method_plot_data %>% 
@@
     x = "Percentage",
-    fill = "Response levels"
+    fill = "Response",
+    title = "The studies using preexisting observational data that I am involved in..."
   ) +
@@
   theme(
-    axis.title.y = element_blank()
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 12),
+    legend.title = element_blank()
   ) 
-```
-
-```{r echo=FALSE}
-method_plot
 ```
@@
 
-```{r warning=FALSE, message=FALSE}
+```{r alspacPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4, message=FALSE}
 # Check if there is any missing and don't understand responses separately
@@
   rename(
-    `ALSPAC should\nrun a study` = alspac_study,
-    `I would\nprefer ECAW` = alspac_prefer_ecaw,
-    `I would opt in` = alspac_opt_in,
-    `I would be\nless willing to use` = alspac_less_willing
+    `If ALSPAC required that I\nuse an ECAW, I would be less willing to\nuse their data in my research` = alspac_less_willing,
+    `If ALSPAC ran a study\non ECAWs, I would opt-in.` = alspac_opt_in,
+    `ALSPAC should run\na study on ECAWs.` = alspac_study,
+    `I would prefer using\nan ECAW than using typical preregistration` = alspac_prefer_ecaw
   )
@@
 # Create likert package data not including the missing values
-alspac_plot_data <- likert(alspac_plot_data)
+alspac_plot_likert_data <- likert(alspac_plot_data)
 
@@
 alspac_plot <- 
-  plot(alspac_plot_data)
-```
+  plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, group.order = names(alspac_plot_data), text.size = 6) +
+  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    legend.text = element_text(size = 12),
+    legend.title = element_blank()
+  )
 
-```{r}
 alspac_plot
@@
 
-Address the concerns from the open-ended questions [transfer to right side of table 1].
-For some research questions, the subset method may not work.
-In which case, a larger sample could be provided, or better yet, a synthetic dataset could be provided.
-The time burden is likely more than when not preregistering, but efforts can be made to minimize it.
-One study comparing ~120 analyses found that an ECAW-like workflow took similar effort to preregistration (although that was for a relatively clean dataset—whereas prereg may take longer if the dataset is a mess; or viceversa).
-The concern that analysis would need to change after writing an ECAW is not a big deal—having the ECAW written could provide evidence that the changes were small.
-This is often raised as an issue in preregistraiton as well, but is acceptable or encouraged when it makes sense to do so.
-Finally the issue that ECAW would not work for situations where researchers are reanalyzing a dataset they already have is true, and it simply would not be possible in this situation (in which case, the researcher would not be requesting that data again anyhow, so it’s beside the point).
+```{r include=FALSE}
+# Open ended questions
+open_ended <-
+  processed %>% 
+  select(response_id, unsure_explain, drawback_ecaw, suggestions, comments) %>% 
+  pivot_longer(cols = c(unsure_explain, drawback_ecaw, suggestions, comments), names_to = "variable", values_to = "value") %>% 
+  filter(!is.na(value))
+```
+
+Table 1.
+Recurring topics in responses to the open-ended survey questions.
+The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
+These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents.
+A complete list of responses are viewable in the open data [LINK].
+We synthesized the response to open-ended questions into the 9 topics on the left side of this table.
+We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as typical preregistration and ECAWs.
+On the right side of the table, we provide a reflection on each topic.
 
@@
 # Discussion
-Are ECAWs applicable 1?
-Yes.
-In short, ECAWs could be implemented in many cases.
-Many respondents use software that would allow for the ECAW format (R, ?Stata?), although some use software where this would be more difficult (SPSS, ?SAS?).
-Most have experience publishing relevant articles, so are familiar with the process.
-And most respondents also run studies that contain confirmatory analyses.
-
-Are ECAWs applicable 2? Yes. Current methods to improve rigour and reproducibility are underused (sharing analysis scripts, blinding data analysts, preregisteration).
-
-How biased are these results?
-Many that are earlier in their career based on # of publications.
-Also many are more concerned than others about reproducibility.
-So this may, unsurprisingly, be a somewhat biased sample.
-Although the 10% response rate was quite good.
-
-Should we ALSPAC (or another org) run a study? Probably.
-
-How would ECAWs be implemented?
-Would ALSPAC or the org check whether they run.
-Where would the be uploaded.
-Would the primary outcomes need to be clearly identified in the output.
-These would depend on the organization running such an experiment / workflow.
 

### 2023-03-17 manuscript/manuscript.tex
updated figures, analysis, and readme based on Robbie's comments

@@
 \documentclass[
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={A survey of registration practices among observational researchers using preexisting datasets},
-  pdfauthor={Robert T. Thibault1,5, Marton Kovacs2,6, Tom E. Hardwicke3, Alexandra Sarafoglou4, John P. A. Ioannidis4, \& Marcus R. Munafò1,7},
-  pdflang={en-EN},
-  pdfkeywords={keywords},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 \keywords{keywords\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
 \usepackage{lineno}
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={A survey of registration practices among observational researchers using preexisting datasets},
+  pdfauthor={Robert T. Thibault1,5, Marton Kovacs2,6, Tom E. Hardwicke3, Alexandra Sarafoglou4, John P. A. Ioannidis4, \& Marcus R. Munafò1,7},
+  pdflang={en-EN},
+  pdfkeywords={keywords},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 
-Many published research findings are non-reproducible and potentially false or misleading (Camerer et al., 2018; Errington et al., 2021; Ioannidis, 2005, 2008; Open Science Collaboration, 2015).
-Researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices can lead to research waste, and useless or even harmful healthcare and policy interventions (e.g., Prasad \& Cifu, 2019).
-To mitigate these issues, clinical trials research often blinds both participants and experimenters, and prospectively registers outcome measures.
-Observational research using preexisting data, however, is rarely registered and the researchers analyzing the data generally have access to complete datasets without any blinding imposed on them.
-
-In this survey study, we seeked to gather data on (1) the opinions and practices of observational researchers regarding rigour and reproducibility, and (2) their thoughts on the use of hold-out samples or cross validation for analyzing preexisting datasets.
-We hope to use this data to refine a potential intervention aimed to improve research quality and to determine whether observational researchers find such an intervention acceptable.
-
-In this intervention we propose, a data management organization would first provide only a subset of the data a researcher requests.
-After the researcher prepares an analysis script based on the subset of data and uploads it to a registry---where it will be openly accessible and permanent---the data management organization then provides the full dataset and the researcher can proceed as they wish.
-The exact implementation (e.g., whether the data management organization runs quality checks on the analysis scripts) would depend on the preferences of the data management organization and the research community using their dataset.
-
-We call this process an Explore and Confirm Analysis Workflow (ECAW).
-Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly.
-ECAWs could serve as an alternative to typical preregistrations for observational research on preexisting data and may be more precise and comprehensive.
-Whereas typical preregistrations are written descriptions of a planned analysis and obstacles may arise in their implementation, ECAWs provide an analysis script that has been shown to run.
-ECAWs would not solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration.
-A version of this workflow has been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al., 2020).
-
-Moreover, a recent study recruited 120 teams to analyze a single observational dataset and had half the teams preregister an analysis plan and the other half prepare an analysis plan by writing an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded) (Sarafoglou et al., 2022).
-This blinded condition is comparable to ECAWs in that the researchers have access to enough data to develop an informed analysis, but without compromising the confirmatory nature of a final analysis on the complete dataset.
-The researchers found that the two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.
-
-We are interested in exploring real-world applications of these types of blind data analysis and hold-out samples (e.g., ECAWs) to improve research rigour and reproducibility.
-One example of such an initiative is OpenSAFELY (www.opensafely.org).
-This platform provides a dataset of simulated health records from which researchers can develop an analysis script.
-When ready, the researchers submit their analysis script which is run in a Trusted Research Environment (TRE) and automatically logged and made public.
-Thus, this workflow keeps health records anonymous while also documenting all analyses that are run on the real data.
-
-In the present study, we solicited structured feedback from observational researchers and found that respondents generally agreed that there could be a benefit to ECAWs and that they are worth testing.
-
 \hypertarget{methods}{%
 \section{Methods}\label{methods}}
-
-We ran a descriptive and exploratory survey study.
-We had no hypotheses, but we did have two specific objectives.
-(1) To gain insights on the opinions and practices of researchers who already use preexisting observational datasets, in regards to the trustworthiness and reproducibility of research.
-(2) To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices.
-This objective includes assessing and refining potential interventions---such as ECAWs---and assessing their acceptability.
-
-We sent an email (supplementary material A) to invite researchers on the mailing list for the Avon Longitudinal Study of Parents and Children (ALSPAC) to participate in an online survey.
-ALSPAC is ``a transgenerational prospective observational study investigating influences on health and development across the life course.
-It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.'' (Boyd et al., 2013).
-Thus, our survey will reach researchers in diverse disciplines that use observational data.
-The survey was open from 10 Oct 2022 to 1 Nov 2022.
-We sent two reminder emails, exactly one week and two weeks after the original email invitation.
-
-The survey (supplementary material B) contained 6 blocks.
-Block 1 assessed whether respondents agree that observational research using preexisting data is trustworthy and reproducible (2 questions).
-Block 2 asked respondents about their own research practices, including preregistration, blinding, and sharing analysis scripts (5 questions).
-Between Block 2 and Block 3, the survey described ECAWs.
-Block 3 assessed whether respondents agree that ECAWs would make observational research using preexisting data more trustworthy and reproducible (2 questions).
-Block 4 directly asked respondents whether ALSPAC should run a study on ECAWs and whether they would participate (5 questions).
-Block 5 contained open-ended questions to garner information that the survey could otherwise not capture (3 questions).
-Block 6 collected information about the respondents.
-(4 questions).
-Open ended questions were analyzed\ldots.{[}add explanation{]}
 
@@
 54 emails bounced, leaving 1094 emails that went through.
-The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for completed surveys and 2\% for incomplete surveys.\footnote{This mailing list has been around for 30 years and may contain email addresses that are no longer monitored.
+The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for complete surveys and 2\% for incomplete surveys.\footnote{The ALSPAC mailing list has been active for \textgreater30 years and may contain email addresses that are no longer monitored.
 For example, we received one email reply stating that the recipient hasn't been active in research for 30 years.
-Thus, the response rate may be greater if we are only including active researchers.} The results presented in this manuscript only include completed surveys.\footnote{Results that include incomplete surveys are available in the Supplementary Material.} The median time taken for complete survey responses was 7.22 minutes (IQR: 13.10 - 4.30).
+Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.22 minutes (IQR: 4.30 to 13.10).
 
-Respondents published a median of NA (IQR 25 - 2) studies using preexisting observational data (Figure S1).
-They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), irrelevant (n = 1), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1).
-52\% of participants reported being more concerned, versus 6\% who reported being less concerned, with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure \#).
-
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
+Respondents published a median of NA (IQR 2 to 25) studies using preexisting observational data (Figure S1).
+They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1)\footnote{Participants could select multiple responses to this survey question.}.
+61\% of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% reported being less concerned.
 
@@
 
-Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy (72\%) and reproducible (79\%) (Figure 1A).
-At the same time, many agreed that a study using an ECAW would be more trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
-That they merit trust.''} (70\%) and more reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (69\%) compared to a typical study using preexisting observational data (Figure 1B).
+\begin{figure}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
+{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/typicallyEcawPlot-1}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
+}
 
-Almost all respondents answered that they use both exploratory (95\%) and confirmatory (92\%) analyses at least sometimes (Figure 2A-B).
-About half reported sharing their analysis scripts never or almost never (20\%), or sometimes (32\%) (Figure 2C).
-Over half reported preregistering their studies never or almost never (38\%), or sometimes (27\%) (Figure 2D).
-79\% reported that they never or almost never blind the data analyst (Figure 2E).
+\caption{Caption goes here...}\label{fig:typicallyEcawPlot}
+\end{figure}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
+Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+That they merit trust.''} (72\%) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%) (Figure 1A).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%) and \emph{more} reproducible (69\%) compared to a typical study using preexisting observational data (Figure 1B).
+
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (38\%), or sometimes (27\%) (Figure 2A).
+About half reported sharing their analysis scripts never or almost never (20\%), or sometimes (32\%) (Figure 2B).
+79\% reported that they never or almost never blind the data analyst (Figure 2C).
+Almost all respondents answered that they use both exploratory (95\%) and confirmatory (92\%) analyses at least sometimes (Figure 2D-E).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/methodPlot-1}
+
+}
+
+\caption{Caption goes here...}\label{fig:methodPlot}
+\end{figure}
 
@@
 
-\begin{verbatim}
-## Warning in likert(alspac_plot_data): Variable I would be
-## less willing to use is not a factor.
-\end{verbatim}
+\begin{figure}
 
-\begin{verbatim}
-## Warning in likert(alspac_plot_data): Variable I would opt in is not a factor.
-\end{verbatim}
+{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/alspacPlot-1}
 
-\begin{verbatim}
-## Warning in likert(alspac_plot_data): Variable ALSPAC should
-## run a study is not a factor.
-\end{verbatim}
+}
 
-\begin{verbatim}
-## Warning in likert(alspac_plot_data): Variable I would
-## prefer ECAW is not a factor.
-\end{verbatim}
+\caption{Caption goes here...}\label{fig:alspacPlot}
+\end{figure}
 
-\includegraphics{manuscript_files/figure-latex/unnamed-chunk-13-1.pdf}
-
-Address the concerns from the open-ended questions {[}transfer to right side of table 1{]}.
-For some research questions, the subset method may not work.
-In which case, a larger sample could be provided, or better yet, a synthetic dataset could be provided.
-The time burden is likely more than when not preregistering, but efforts can be made to minimize it.
-One study comparing \textasciitilde120 analyses found that an ECAW-like workflow took similar effort to preregistration (although that was for a relatively clean dataset---whereas prereg may take longer if the dataset is a mess; or viceversa).
-The concern that analysis would need to change after writing an ECAW is not a big deal---having the ECAW written could provide evidence that the changes were small.
-This is often raised as an issue in preregistraiton as well, but is acceptable or encouraged when it makes sense to do so.
-Finally the issue that ECAW would not work for situations where researchers are reanalyzing a dataset they already have is true, and it simply would not be possible in this situation (in which case, the researcher would not be requesting that data again anyhow, so it's beside the point).
+Table 1.
+Recurring topics in responses to the open-ended survey questions.
+The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
+These questions received a total of (92) responses from (55) unique respondents.
+A complete list of responses are viewable in the open data {[}LINK{]}.
+We synthesized the response to open-ended questions into the 9 topics on the left side of this table.
+We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as typical preregistration and ECAWs.
+On the right side of the table, we provide a reflection on each topic.
 
@@
 \section{Discussion}\label{discussion}}
-
-Are ECAWs applicable 1?
-Yes.
-In short, ECAWs could be implemented in many cases.
-Many respondents use software that would allow for the ECAW format (R, ?Stata?), although some use software where this would be more difficult (SPSS, ?SAS?).
-Most have experience publishing relevant articles, so are familiar with the process.
-And most respondents also run studies that contain confirmatory analyses.
-
-Are ECAWs applicable 2? Yes. Current methods to improve rigour and reproducibility are underused (sharing analysis scripts, blinding data analysts, preregisteration).
-
-How biased are these results?
-Many that are earlier in their career based on \# of publications.
-Also many are more concerned than others about reproducibility.
-So this may, unsurprisingly, be a somewhat biased sample.
-Although the 10\% response rate was quite good.
-
-Should we ALSPAC (or another org) run a study? Probably.
-
-How would ECAWs be implemented?
-Would ALSPAC or the org check whether they run.
-Where would the be uploaded.
-Would the primary outcomes need to be clearly identified in the output.
-These would depend on the organization running such an experiment / workflow.
 

### 2023-03-17 manuscript/supplementary_materials_c.Rmd
updated figures, analysis, and readme based on Robbie's comments

---
title             : "Supplementary materials C: A survey of registration practices among observational researchers using preexisting datasets"
shorttitle        : "Supplementary materials C"

keywords          : "keywords"
wordcount         : "X"
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library("papaja")
library(tidyverse)
library(here)
library(likert)
library(kableExtra)

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

```{r analysis-preferences}
# Seed for random number generation
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r numStudies, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=6, fig.height=3.2}
# Create text for missing responses
n_studies_missing_text <- 
  processed %>% 
  filter(is.na(n_studies)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

# Create histogram of the number of observational studies done
processed %>% 
  # Exclude missing responses
  filter(!is.na(n_studies)) %>% 
  ggplot() +
  aes(x = n_studies) +
  geom_histogram(binwidth = 3) +
  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  labs(
    x = "Number of studies",
    y = "Count",
    title = "Approximately how many studies have you published using a preexisting observational dataset (e.g., the ALSPAC dataset)?"
  ) +
  papaja::theme_apa()

# ggplot() +
#   geom_histogram(
#     data = dplyr::filter(valami, n_studies == 0),
#     mapping = aes(x = n_studies), binwidth = 1, color = 'red') +
#   geom_histogram(
#     data = dplyr::filter(valami, n_studies != 0),
#     mapping = aes(x = n_studies), binwidth = 5) +
#   # annotate(geom = "text", label = n_studies_missing_text, y = 30) +
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
#   labs(
#     x = "Number of observational studies done prior using ALSPAC",
#     y = "Count"
#   ) +
#   papaja::theme_apa()
```

```{r}
# Programming languages table
processed %>% 
  rename(lang = programming_language) %>% 
  separate_rows(lang, sep = ",") %>% 
  count(lang) %>% 
  replace_na(list(lang = "Missing")) %>% 
  # Removing irrelevant
  filter(lang != "irrelevant") %>% 
  ungroup() %>% 
  mutate(
    `Percentage of respondents` = round(n / sum(n) * 100, 0)
  ) %>% 
  arrange(desc(n)) %>% 
  rename(
    `Programming language` = lang,
    `N` = n
  ) %>% 
  kbl(
  .,
  format = "latex",
  col.names = c("Programming language", "N", "Percentage of respondents"),
  booktabs = T,
  escape = F,
  caption = "(ref:table1c-caption)",
  centering = T,
  position = "H") %>%
  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
  row_spec(row = 0, bold = TRUE) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity") %>% 
  add_footnote("One response was removed because it was irrelevant to the question. ", notation = "symbol")
```

(ref:table1c-caption) What programming language or software do you use for your analyses of preexisting observational data?

```{r concerned, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4}
# Concerned figure data preparation
concerned_plot_data <-
  processed %>% 
  select(concerned) %>% 
  mutate(
    concerned = tolower(concerned),
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned"
      )
    )
  )

# Check distinct values in data
# distinct(concerned_plot_data, concerned)

# Create missing values text
concerned_missing_text <- 
  concerned_plot_data %>% 
  filter(is.na(concerned)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

# Create likert package data not including the missing values
concerned_plot_data <- likert(filter(concerned_plot_data, !is.na(concerned)))

# Create figure
plot(concerned_plot_data, digits = 1, text.size = 6) +
  labs(title = "Compared to what you think of as a typical researcher who uses preexisting observational data in your field,\nhow concerned are you with research trustworthiness, bias, rigour, and reproducibility ...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 10),
    legend.title = element_blank()
  )
```

### 2023-03-17 R/utils.R

@@ -8,14 +8,23 @@ n_levels <- function(var) length(levels(var))
 center_level <- function(var) (n_levels(var) - 1) / 2 + 1
 
 #' Calculate the percentage of responses of each factor level
-level_percentage <- function(data, var, levels, exclude = NULL) {
-  data %>% 
+level_percentage <- function(data, var, levels, exclude = NULL, exclude_missing = TRUE) {
+  res <-
+    data %>% 
     filter(
       # exclude specified answer option(s)
-      {{var}} %ni% exclude & 
-        # exclude missing values by default
-        !is.na({{var}})
-    ) %>% 
+      {{var}} %ni% exclude
+    )
+  
+  # exclude missing responses
+  if(exclude_missing) {
+    res <-
+      res %>% 
+      filter(exclude_missing & !is.na({{var}}))
+  }
+
+  res <-
+    res %>% 
     mutate(
       # Transform to factor with level order specified in levels arg
       {{var}} := factor({{var}}, levels = levels)
@@ -37,9 +46,9 @@ level_percentage <- function(data, var, levels, exclude = NULL) {
 }
 
 #' Calculate percentage for each support category
-support_percentage <- function(data, var, levels, exclude = NULL) {
+support_percentage <- function(data, var, levels, exclude = NULL, exclude_missing = TRUE) {
   # Calculate the percentage of responses of each factor level
-  level_percentage(data, {{var}}, levels, exclude = exclude) %>% 
+  level_percentage(data, {{var}}, levels, exclude = exclude, exclude_missing = exclude_missing) %>% 
     # Drop these variables
     select(-n_sum, -percentage) %>% 
     mutate(

### 2023-03-17 manuscript/interactive_report.Rmd
fixed interactive document

@@
 
-what is this document...
-
-# Subset
+# Settings
 
@@
 
+# Results from the manuscript
+
 ```{r echo=FALSE, warning=FALSE}
@@
 ecaw_plot <- reactive({
-  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right") +
+  plot(ecaw_plot_data(), digits = 1, ordered = FALSE, legend.position = "right") +
   labs(title = "Compared to a typical study using preexisting observational data, a study using an ECAW would be...")
@@
 # Join the two plots together to make one figure
-renderPlot({ typically_plot + ecaw_plot + plot_layout(ncol = 1, nrow = 2) })
+renderPlot({ typically_plot() + ecaw_plot() + plot_layout(ncol = 1, nrow = 2) })
 ```
 
-```{r}
+---
+
+```{r echo=FALSE}
+# Used methods figures
+method_plot_data <- reactive({
+  filtered_data() %>% 
+    select(starts_with("method_")) %>% 
+    pivot_longer(
+      cols = everything(),
+      names_to = "variable",
+      values_to = "scale",
+    ) %>%
+    separate(variable, into = c("name_prefix", "type")) %>% 
+    group_by(type) %>% 
+    count(scale) %>% 
+  ungroup() %>% 
+    replace_na(list(scale = "Missing")) %>% 
+    mutate(
+      scale = factor(scale, levels = c(
+          "Never or almost never",
+          "Sometimes",
+          "About half the time",
+          "Most of the time",
+          "Always or almost always",
+          "Missing",
+          "I don't understand the question")
+            ),
+      type = case_when(
+        type == "blind" ~ "Blind the data analyst",
+        type == "confirmatory" ~ "Contain confirmatory analysis",
+        type == "exploratory" ~ "Contain exploratory analysis",
+        type == "preregistered" ~ "Are preregistered",
+        type == "script" ~ "Share analysis scripts",
+      ),
+      type = factor(type, levels = c(
+        "Contain exploratory analysis",
+        "Contain confirmatory analysis",
+        "Blind the data analyst",
+        "Share analysis scripts",
+        "Are preregistered"
+      ))
+      ) %>% 
+  group_by(type) %>%
+  tidyr::complete(scale, fill = list(n = 0)) %>% 
+  mutate(
+    percentage = n / sum(n)
+  ) %>% 
+  ungroup()
+})
+
+renderPlot({
+method_plot_data() %>% 
+  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
+  ggplot() +
+  aes(
+    x = percentage,
+    y = type,
+    fill = scale
+  ) +
+  geom_bar(position = "stack", stat = "identity") +
+  scale_x_continuous(
+    limits = c(0, 1),
+    labels = scales::label_percent()
+  ) +
+  scale_fill_viridis(option = "A", discrete = TRUE) +
+  labs(
+    x = "Percentage",
+    fill = "Response",
+    title = "The studies using preexisting observational data that I am involved in..."
+  ) +
+  papaja::theme_apa() +
+  theme(
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 12),
+    legend.title = element_blank()
+  ) 
+})
+```
+
+---
+
+```{r echo=FALSE}
+# Prepare plot data
+alspac_plot_data <- reactive({
+  filtered_data() %>%
+  select(contains("alspac_")) %>% 
+  # Transform not wanted response values to NA
+  # likert::likert drops NA values silently
+  # When var transformed to factor these values would be transformed to NA
+  # Automatically but I try to be explicit
+  mutate(
+    across(
+      everything(),
+      ~ case_when(
+        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
+        TRUE ~ .
+      )
+    )
+  ) %>% 
+  mutate(
+    across(
+      .fns = ~ factor(., levels = c(
+          "Strongly disagree",
+          "Somewhat disagree",
+          "Neither agree nor disagree",
+          "Somewhat agree",
+          "Strongly agree")
+          )
+      )
+    ) %>% 
+  rename(
+    `If ALSPAC required that I\nuse an ECAW, I would be less willing to\nuse their data in my research` = alspac_less_willing,
+    `If ALSPAC ran a study\non ECAWs, I would opt-in.` = alspac_opt_in,
+    `ALSPAC should run\na study on ECAWs.` = alspac_study,
+    `I would prefer using\nan ECAW than using typical preregistration` = alspac_prefer_ecaw
+  )
+})
+
+# Create likert package data not including the missing values
+alspac_plot_likert_data <- reactive({likert(alspac_plot_data())})
+
+# Create figure
+renderPlot({
+  plot(alspac_plot_likert_data(), digits = 1, ordered = TRUE, group.order = names(alspac_plot_data()), text.size = 6) +
+  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    legend.text = element_text(size = 12),
+    legend.title = element_blank()
+  )
+})
+```
+
+# Results from supplementary materials C
+
+```{r echo=FALSE}
 renderPlot({
@@
 ```
+
+---
+
+```{r echo=FALSE}
+# Programming languages table
+renderTable({
+filtered_data() %>% 
+  rename(lang = programming_language) %>% 
+  separate_rows(lang, sep = ",") %>% 
+  count(lang) %>% 
+  replace_na(list(lang = "Missing")) %>% 
+  # Removing irrelevant
+  filter(lang != "irrelevant") %>% 
+  ungroup() %>% 
+  mutate(
+    `Percentage of respondents` = round(n / sum(n) * 100, 0)
+  ) %>% 
+  arrange(desc(n)) %>% 
+  rename(
+    `Programming language` = lang,
+    `N` = n
+  )
+})
+```
+
+---
+
+```{r echo=FALSE}
+# Concerned figure data preparation
+concerned_plot_data <- reactive({
+  processed %>% 
+  select(concerned) %>% 
+  mutate(
+    concerned = tolower(concerned),
+    concerned = factor(concerned, levels = c(
+      "very much less concerned",
+      "less concerned",
+      "somewhat less concerned",
+      "as concerned as a typical researcher in my field",
+      "somewhat more concerned",
+      "more concerned",
+      "very much more concerned"
+      )
+    )
+  ) %>% 
+    filter(!is.na(concerned))
+})
+
+# Create likert package data not including the missing values
+concerned_likert_plot_data <- reactive({likert(concerned_plot_data())})
+
+# Create figure
+renderPlot({
+  plot(concerned_likert_plot_data(), digits = 1, text.size = 6, group.order = names(concerned_plot_data())) +
+  labs(title = "Compared to what you think of as a typical researcher who uses preexisting observational data in your field,\nhow concerned are you with research trustworthiness, bias, rigour, and reproducibility ...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    legend.text = element_text(size = 10),
+    legend.title = element_blank()
+  )
+})
+```

### 2023-03-17 README.Rmd
manuscript.Rmd runs the preprocessing files in a new R session safely, readme updated with instructions on how to run interactive document and replicate the restuls

@@
 
+__In order to replicate our results including the whole data preprocessing procedure with one click, please knit the `manuscript/manuscript.Rmd` file.__
+
 ## Folder structure
@@
 The `manuscript/` folder contains the manuscript and supplementary materials documents with the analysis code embedded in them.
+The folder also includes the `interactive_report.Rmd` file, which includes an interactive document for reanalysing our results with different subsets of the data.
+To run the interactive document locally on your computer, click the _Run document_ button in Rstudio or run the `rmarkdown::run(here::here("manuscript/interactive_report.Rmd"))` function in your R console.
 

### 2023-03-17 manuscript/manuscript.Rmd
manuscript.Rmd runs the preprocessing files in a new R session safely, readme updated with instructions on how to run interactive document and replicate the restuls

@@
 library(likert)
+library(knitr)
 library(viridis)
@@
 
-```{r preprocessing}
+```{r preprocessing, include=FALSE}
 # loads source data, performs preprocessing, saves raw datafiles
-# source(here("preprocessing", "ecaw_source_raw_preprocessing.R"))
+xfun::Rscript_call(
+  rmarkdown::render,
+  list(
+    input = here::here("preprocessing", "ecaw_source_raw_preprocessing.Rmd"),
+    output_format = 'html_document')
+  )
 
 # loads raw data, performs preprocessing, saves processed datafiles
-# source(here("preprocessing", "ecaw_raw_processed_preprocessing.R"))
+xfun::Rscript_call(
+  rmarkdown::render,
+  list(
+    input = here::here("preprocessing", "ecaw_raw_processed_preprocessing.Rmd"),
+    output_format = 'html_document')
+  )
 ```
@@
 ```{r loading_data}
-processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
+processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
 ```

### 2023-03-17 manuscript/supplementary_materials_c.Rmd
manuscript.Rmd runs the preprocessing files in a new R session safely, readme updated with instructions on how to run interactive document and replicate the restuls

@@
 
+concerned_plot_data <- filter(concerned_plot_data, !is.na(concerned))
+
 # Create likert package data not including the missing values
-concerned_plot_data <- likert(filter(concerned_plot_data, !is.na(concerned)))
+concerned_likert_plot_data <- likert(concerned_plot_data)
 
 # Create figure
-plot(concerned_plot_data, digits = 1, text.size = 6) +
+plot(concerned_likert_plot_data, digits = 1, text.size = 6, group.order = names(concerned_plot_data)) +
   labs(title = "Compared to what you think of as a typical researcher who uses preexisting observational data in your field,\nhow concerned are you with research trustworthiness, bias, rigour, and reproducibility ...") +

### 2023-03-17 manuscript/manuscript.Rmd
small fixes to the plots and tables

@@
 footnotelist      : no
+appendix: "supplementary_materials.Rmd"
+
 classoption       : "man"
@@
 
-```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=6}
+```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=12.2, fig.height=6}
 # Create typically trustworthy and reproducible plot
@@
 typically_plot <- 
-  plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.7) +
+  plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
   labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
@@
     title = element_text(size = 14),
-    axis.text.y = element_blank(),
+    axis.text.y = element_text(size = 14),
     axis.text.x = element_text(size = 15),
-    axis.title.x = element_text(size = 15),
-    legend.text = element_text(size = 12),
+    axis.title.x = element_blank(),
+    legend.text = element_text(size = 13.5),
     legend.title = element_blank()
@@
 ecaw_plot <- 
-  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.7) +
+  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
   labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
@@
     title = element_text(size = 14),
-    axis.text.y = element_blank(),
+    axis.text.y = element_text(size = 14),
     axis.text.x = element_text(size = 15),
     axis.title.x = element_text(size = 15),
-    legend.text = element_text(size = 12),
-    legend.title = element_blank()
+    legend.text = element_text(size = 13.5),
+    legend.title = element_blank(),
+    # legend.box.spacing = unit(0, "cm")
+    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2, unit = "cm")
   )
@@
 
-```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4.5}
+```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5}
 # Used methods figures
@@
       type = case_when(
-        type == "blind" ~ "Blind the data analyst",
-        type == "confirmatory" ~ "Contain confirmatory analysis",
-        type == "exploratory" ~ "Contain exploratory analysis",
+        type == "blind" ~ "Blind the\ndata analyst",
+        type == "confirmatory" ~ "Contain\nconfirmatory analysis",
+        type == "exploratory" ~ "Contain\nexploratory analysis",
         type == "preregistered" ~ "Are preregistered",
@@
       type = factor(type, levels = c(
-        "Contain exploratory analysis",
-        "Contain confirmatory analysis",
-        "Blind the data analyst",
+        "Contain\nexploratory analysis",
+        "Contain\nconfirmatory analysis",
+        "Blind the\ndata analyst",
         "Share analysis scripts",
@@
   theme(
+    title = element_text(size = 17),
     axis.title.y = element_blank(),
-    legend.text = element_text(size = 12),
-    legend.title = element_blank()
+    axis.text = element_text(size = 15),
+    axis.title.x = element_text(size = 17),
+    legend.text = element_text(size = 14),
+    legend.title = element_blank(),
+    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
   ) 
@@
 
-```{r alspacPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4, message=FALSE}
+```{r alspacPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=14, fig.height=4, message=FALSE}
 # Check if there is any missing and don't understand responses separately
@@
   rename(
-    `If ALSPAC required that I\nuse an ECAW, I would be less willing to\nuse their data in my research` = alspac_less_willing,
-    `If ALSPAC ran a study\non ECAWs, I would opt-in.` = alspac_opt_in,
-    `ALSPAC should run\na study on ECAWs.` = alspac_study,
-    `I would prefer using\nan ECAW than using typical preregistration` = alspac_prefer_ecaw
+    `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
+    `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
+    `ALSPAC should run a study on ECAWs.` = alspac_study,
+    `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
   )
@@
 # Create figure
-alspac_plot <- 
 plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, group.order = names(alspac_plot_data), text.size = 6) +
@@
     title = element_text(size = 14),
-    axis.text.y = element_blank(),
+    axis.text.y = element_text(size = 13),
     axis.text.x = element_text(size = 15),
     axis.title.x = element_text(size = 15),
-    legend.text = element_text(size = 12),
+    legend.text = element_text(size = 13),
     legend.title = element_blank()
   )
-
-alspac_plot
 ```
@@
 :::
+
+\newpage
+
+# (APPENDIX) Appendix {-}
+
+```{r child = "supplementary_materials.Rmd"}
+```

### 2023-03-17 manuscript/manuscript.tex
small fixes to the plots and tables

@@
 
+\newpage
+
+\hypertarget{appendix-appendix}{%
+\appendix}
+
+\hypertarget{supplementary-materials-c}{%
+\section{Supplementary materials C}\label{supplementary-materials-c}}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/numStudies-1}
+
+}
+
+\caption{Caption goes here...}\label{fig:numStudies}
+\end{figure}
+
+\begin{table}[H]
+
+\caption{\label{tab:unnamed-chunk-7}What programming language or software do you use for your analyses of preexisting observational data?}
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lrr}
+\toprule
+\textbf{Programming language} & \textbf{N} & \textbf{Percentage of respondents}\\
+\midrule
+R & 65 & 61\\
+\cmidrule{1-3}
+Stata & 48 & 45\\
+\cmidrule{1-3}
+SPSS & 17 & 16\\
+\cmidrule{1-3}
+SAS & 15 & 14\\
+\cmidrule{1-3}
+Python & 6 & 6\\
+\cmidrule{1-3}
+Missing & 4 & 4\\
+\cmidrule{1-3}
+Mplus & 3 & 3\\
+\cmidrule{1-3}
+Bash & 2 & 2\\
+\cmidrule{1-3}
+MATLAB & 1 & 1\\
+\cmidrule{1-3}
+Nextflow & 1 & 1\\
+\cmidrule{1-3}
+plink2 & 1 & 1\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/concerned-1}
+
+}
+
+\caption{Caption goes here...}\label{fig:concerned}
+\end{figure}
+
 \end{document}

### 2023-03-17 manuscript/supplementary_materials.Rmd
small fixes to the plots and tables

--- 2023-03-14 manuscript/supplementary_materials.Rmd
@@
----
-title             : "Supplementary materials: A survey of registration practices among observational researchers using preexisting datasets"
-shorttitle        : "Supplementary materials"
+# Supplementary materials C
 
-keywords          : "keywords"
-wordcount         : "X"
-floatsintext      : no
-linenumbers       : yes
-draft             : no
-mask              : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-classoption       : "man"
-output            : papaja::apa6_pdf
-editor_options: 
-  chunk_output_type: console
----
-
-```{r setup, include = FALSE}
-library("papaja")
-library(tidyverse)
-
-# Source R scripts
-r_scripts <- list.files(here::here("R/"), full.names = TRUE)
-purrr::walk(r_scripts, source)
-
-# Loading data
-processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
-```
-
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
-```
-
-# Table
-
-```{r}
-# programming languages table
-processed %>% 
-  rename(lang = programming_language) %>% 
-  separate_rows(lang, sep = ",") %>% 
-  count(lang) %>% 
-  replace_na(list(lang = "Missing")) %>% 
-  arrange(desc(n)) %>% 
-  rename(
-    `Programming language` = lang,
-    `Number of occurance` = n
-  ) %>% 
-  kbl(
-  .,
-  format = "latex",
-  col.names = c("Programming language", "Number of occurance"),
-  booktabs = T,
-  escape = F,
-  caption = "Programming Languages Used to Analyse Preexisting Observational Data",
-  centering = T,
-  position = "H") %>%
-  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
-  row_spec(row = 0, bold = TRUE) %>%
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
-```
-
-# Figure
-
-```{r}
+```{r numStudies, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=6, fig.height=3.2}
 # Create text for missing responses
@@
   aes(x = n_studies) +
-  geom_histogram() +
-  annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
+  geom_histogram(binwidth = 3) +
+  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
   labs(
-    x = "Number of observational studies done prior using ALSPAC",
-    y = "Count"
+    x = "Number of publications",
+    y = "Number of respondents",
+    title = "Approximately how many studies have you published using\na preexisting observational dataset (e.g., the ALSPAC dataset)?"
   ) +
   papaja::theme_apa()
+
+# ggplot() +
+#   geom_histogram(
+#     data = dplyr::filter(valami, n_studies == 0),
+#     mapping = aes(x = n_studies), binwidth = 1, color = 'red') +
+#   geom_histogram(
+#     data = dplyr::filter(valami, n_studies != 0),
+#     mapping = aes(x = n_studies), binwidth = 5) +
+#   # annotate(geom = "text", label = n_studies_missing_text, y = 30) +
+#   scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
+#   labs(
+#     x = "Number of observational studies done prior using ALSPAC",
+#     y = "Count"
+#   ) +
+#   papaja::theme_apa()
 ```
+
+```{r}
+# Programming languages table
+# We use the number of all participants who responded to the questionnaire as the denominator
+denominator <- nrow(processed)
+
+processed %>% 
+  rename(lang = programming_language) %>% 
+  separate_rows(lang, sep = ",") %>% 
+  count(lang) %>% 
+  replace_na(list(lang = "Missing")) %>% 
+  # Removing irrelevant
+  filter(lang != "irrelevant") %>% 
+  ungroup() %>% 
+  mutate(
+    `Percentage of respondents` = round(n / denominator * 100, 0)
+  ) %>% 
+  arrange(desc(n)) %>% 
+  rename(
+    `Programming language` = lang,
+    `N` = n
+  ) %>% 
+  kbl(
+  .,
+  format = "latex",
+  col.names = c("Programming language", "N", "Percentage of respondents"),
+  booktabs = T,
+  escape = F,
+  caption = "(ref:table1c-caption)",
+  centering = T,
+  position = "H") %>%
+  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
+  row_spec(row = 0, bold = TRUE) %>%
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
+```
+
+(ref:table1c-caption) What programming language or software do you use for your analyses of preexisting observational data?
+
+```{r concerned, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4}
+# Concerned figure data preparation
+concerned_plot_data <-
+  processed %>% 
+  select(concerned) %>% 
+  mutate(
+    concerned = tolower(concerned),
+    concerned = factor(concerned, levels = c(
+      "very much less concerned",
+      "less concerned",
+      "somewhat less concerned",
+      "as concerned as a typical researcher in my field",
+      "somewhat more concerned",
+      "more concerned",
+      "very much more concerned"
+      )
+    )
+  )
+
+# Check distinct values in data
+# distinct(concerned_plot_data, concerned)
+
+# Create missing values text
+concerned_missing_text <- 
+  concerned_plot_data %>% 
+  filter(is.na(concerned)) %>% 
+  summarise(
+    n = n(),
+    missing_text = paste0("missing = ", n)
+  ) %>% 
+  pull(missing_text)
+
+concerned_plot_data <- filter(concerned_plot_data, !is.na(concerned))
+
+# Create likert package data not including the missing values
+concerned_likert_plot_data <- likert(concerned_plot_data)
+
+# Create figure
+plot(concerned_likert_plot_data, digits = 1, text.size = 6, group.order = names(concerned_plot_data)) +
+  labs(title = "Compared to what you think of as a typical researcher who uses preexisting observational\ndata in your field, how concerned are you with\nresearch trustworthiness, bias, rigour, and reproducibility ...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_blank(),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_text(size = 15),
+    legend.text = element_text(size = 10),
+    legend.title = element_blank()
+  )
+```

### 2023-03-18 manuscript/interactive_report.Rmd
cosmetic changes to plot, small bug fixes with the code, styling

@@
   selectInput("methods", label = "Used the following methods:",
-              choices = c(
-                "Blind" = "method_blind",
-                "Script" = "method_script",
-                "Preregistration" = "method_preregistered",
-                "Confirmatory" = "method_confirmatory", 
-                "Exploratory" = "method_exploratory"),
+              choices = list(
+                "Blind" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+                "Script" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+                "Preregistration" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+                "Confirmatory" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"), 
+                "Exploratory" = list("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question")),
               selected = NULL, multiple = TRUE),
@@
 filtered_data <- reactive({
+  # list(
+  #               "Blind" = "method_blind",
+  #               "Script" = "method_script",
+  #               "Preregistration" = "method_preregistered",
+  #               "Confirmatory" = "method_confirmatory", 
+  #               "Exploratory" = "method_exploratory")
+  # 
   concerned_options <-list(
@@
     filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
-    filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
+    # filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
     filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
@@
 # shiny::renderTable(filtered_data())
-shiny::renderText({nrow(filtered_data())})
+shiny::renderText({input$methods})
 ```
@@
 concerned_plot_data <- reactive({
-  processed %>% 
+  filtered_data() %>% 
   select(concerned) %>% 

### 2023-03-18 manuscript/manuscript.Rmd
cosmetic changes to plot, small bug fixes with the code, styling

@@
 
-```{r setup, include = FALSE, warning=FALSE, message=FALSE}
+```{r setup, include=FALSE, warning=FALSE, message=FALSE}
 library("papaja")
@@
 
-```{r custom_scripts}
+```{r customScripts}
 # Source R scripts
@@
     input = here::here("preprocessing", "ecaw_source_raw_preprocessing.Rmd"),
-    output_format = 'html_document')
+    output_format = "html_document"
+  )
 )
@@
     input = here::here("preprocessing", "ecaw_raw_processed_preprocessing.Rmd"),
-    output_format = 'html_document')
+    output_format = "html_document"
+  )
 )
@@
 
-```{r loading_data}
+```{r loadingData}
 processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
@@
 
-```{r analysis_preferences}
+```{r analysisPreferences}
 # Seed for random number generation
@@
 
-```{r include=FALSE}
+```{r languageDescriptives, include=FALSE}
 # List of used programming languages and their counts text
@@
   glue::glue_collapse(., sep = ", ", last = ", and ")
+```
 
+```{r concernedDescriptives, include=FALSE}
 # Calculate concerned percentage
@@
 
-```{r include=FALSE}
+```{r typicallyEcawDescriptives, include=FALSE}
 # Calculate typical trustworthiness and reproducibility ratings
@@
     "Somewhat agree",
-    "Strongly agree"),
+    "Strongly agree"
+  ),
   "I don't understand the question"
@@
     "Somewhat more",
-    "Much more"),
+    "Much more"
+  ),
   "I don't understand the question"
@@
 
-```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=12.2, fig.height=6}
+```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
 # Create typically trustworthy and reproducible plot
@@
         "Somewhat agree",
-    "Strongly agree"))
-    )) %>% 
+        "Strongly agree"
+      ))
+    )
+  ) %>%
   rename(
@@
   plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
   labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
@@
   dplyr::filter(is.na(value) | value == "I don't understand the question") %>%
-  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
+  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
   replace_na(list(value = "Missing")) %>%
-  group_by(item , value) %>% 
+  group_by(item, value) %>%
   summarise(
@@
         "Somewhat more",
-          "Much more"))
-    )) %>% 
+        "Much more"
+      ))
+    )
+  ) %>%
   rename(
@@
   plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
   labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
@@
     axis.text.x = element_text(size = 15),
-    axis.title.x = element_text(size = 15),
+    axis.title.x = element_blank(),
     legend.text = element_text(size = 13.5),
     legend.title = element_blank(),
-    # legend.box.spacing = unit(0, "cm")
-    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2, unit = "cm")
+    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2.3, unit = "cm")
   )
@@
 
-```{r include=FALSE}
+```{r methodDescriptives, include=FALSE}
 confirmatory <- level_percentage(
@@
     "Most of the time",
-    "Always or almost always"),
+    "Always or almost always"
+  ),
   "I don't understand the question"
@@
 
-```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5}
+```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
 # Used methods figures
@@
       "Missing",
-          "I don't understand the question")
-            ),
+      "I don't understand the question"
+    )),
     type = case_when(
-        type == "blind" ~ "Blind the\ndata analyst",
-        type == "confirmatory" ~ "Contain\nconfirmatory analysis",
-        type == "exploratory" ~ "Contain\nexploratory analysis",
+      type == "blind" ~ "Blind the data analyst",
+      type == "confirmatory" ~ "Contain confirmatory analysis",
+      type == "exploratory" ~ "Contain exploratory analysis",
       type == "preregistered" ~ "Are preregistered",
@@
     type = factor(type, levels = c(
-        "Contain\nexploratory analysis",
-        "Contain\nconfirmatory analysis",
-        "Blind the\ndata analyst",
+      "Contain exploratory analysis",
+      "Contain confirmatory analysis",
+      "Blind the data analyst",
       "Share analysis scripts",
@@
   scale_fill_viridis(option = "A", discrete = TRUE) +
+  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) +
   labs(
-    x = "Percentage",
-    fill = "Response",
     title = "The studies using preexisting observational data that I am involved in..."
@@
     title = element_text(size = 17),
+    axis.text = element_text(size = 15),
+    axis.title.x = element_blank(),
     axis.title.y = element_blank(),
-    axis.text = element_text(size = 15),
-    axis.title.x = element_text(size = 17),
     legend.text = element_text(size = 14),
@@
 
-```{r include=FALSE, warning=FALSE, message=FALSE}
+```{r aslpacDescriptives, include=FALSE, warning=FALSE, message=FALSE}
 willing <- support_percentage(
@@
     "Somewhat agree",
-    "Strongly agree"),
+    "Strongly agree"
+  ),
   c(
@@
 
-```{r alspacPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=14, fig.height=4, message=FALSE}
+```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
 # Check if there is any missing and don't understand responses separately
@@
   dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>%
-  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
+  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
   replace_na(list(value = "Missing")) %>%
-  group_by(item , value) %>% 
+  group_by(item, value) %>%
   summarise(
@@
         "Somewhat agree",
-          "Strongly agree")
-          )
+        "Strongly agree"
+      ))
     )
@@
 # Create figure
-plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, group.order = names(alspac_plot_data), text.size = 6) +
+plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, text.size = 6) +
+  # TODO: have to check if this is not messing with the results
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 30), limits = rev(names(alspac_plot_data))) +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
   labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
   theme(
-    title = element_text(size = 14),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 15),
-    axis.title.x = element_text(size = 15),
-    legend.text = element_text(size = 13),
+    axis.title.x = element_blank(),
+    title = element_text(size = 15.5),
+    axis.text.y = element_text(size = 14),
+    axis.text.x = element_text(size = 16),
+    legend.text = element_text(size = 15),
     legend.title = element_blank()

### 2023-03-18 manuscript/manuscript.tex
cosmetic changes to plot, small bug fixes with the code, styling

@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/typicallyEcawPlot-1}
+{\centering \includegraphics[width=1\linewidth]{figs/typicallyEcawPlot-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/methodPlot-1}
+{\centering \includegraphics[width=1\linewidth]{figs/methodPlot-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/alspacPlot-1}
+{\centering \includegraphics[width=1\linewidth]{figs/alspacPlot-1}
 
@@
 
-\caption{\label{tab:unnamed-chunk-7}What programming language or software do you use for your analyses of preexisting observational data?}
+\caption{\label{tab:unnamed-chunk-3}What programming language or software do you use for your analyses of preexisting observational data?}
 \resizebox{\linewidth}{!}{

### 2023-03-18 manuscript/supplementary_materials.Rmd
cosmetic changes to plot, small bug fixes with the code, styling

@@
 plot(concerned_likert_plot_data, digits = 1, text.size = 6, group.order = names(concerned_plot_data)) +
-  labs(title = "Compared to what you think of as a typical researcher who uses preexisting observational\ndata in your field, how concerned are you with\nresearch trustworthiness, bias, rigour, and reproducibility ...") +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+  labs(title = "Compared to what you think of as a typical researcher who uses\npreexisting observational data in your field, how concerned\nare you with research trustworthiness, bias, rigour, and reproducibility ...") +
   theme(
@@
     axis.text.x = element_text(size = 15),
-    axis.title.x = element_text(size = 15),
+    axis.title.x = element_blank(),
     legend.text = element_text(size = 10),

### 2023-03-18 R/utils.R

@@ -20,10 +20,10 @@ level_percentage <- function(data, var, levels, exclude = NULL, exclude_missing 
   if(exclude_missing) {
     res <-
       res %>% 
-      filter(exclude_missing & !is.na({{var}}))
+      filter(!is.na({{var}}))
   }
 
-  res <-
+  # res <-
     res %>% 
     mutate(
       # Transform to factor with level order specified in levels arg

### 2023-03-18 manuscript/interactive_report.Rmd
made interactive report prettier, added more setting options

--- 2023-03-17 manuscript/interactive_report.Rmd
@@
 title: "Results exploration"
-author: "Marton Kovacs"
-date: "`r Sys.Date()`"
 output: html_document
@@
 library(plotly)
+library(DT)
 # r_refs("references.bib")
@@
 
-what is this document...
+The aim of this document is to allow users to interactively explore the descriptive results of the study titled _"A survey of registration practices among observational researchers using preexisting datasets"_.
 
-# Subset
+# Settings
 
-Subset data for all the results.
+By using these filters you can subset the dataset for all the results presented in this report.
+The full dataset only contains the responses of those participants who finished the survey (n = `r nrow(processed)`).
 
 ```{r, echo=FALSE}
+# Create a panel for the settings
 inputPanel(
+  # Filter by the number of studies
   sliderInput("n_studies", label = "Number of studies using ALSPAC:",
@@
               max = max(processed$n_studies, na.rm = T), value = c(0, max(processed$n_studies, na.rm = T)), step = 1),
-  selectInput("methods", label = "Used the following methods:",
-              choices = c(
-                "Blind" = "method_blind",
-                "Script" = "method_script",
-                "Preregistration" = "method_preregistered",
-                "Confirmatory" = "method_confirmatory", 
-                "Exploratory" = "method_exploratory"),
-              selected = NULL, multiple = TRUE),
+  # Filter by how concerned the respondent is with rigour and reproducibility
+  # TODO: The last level was recorded with a capital letter, needs to be changed because it is a major source of error
   selectInput("concerned", label = "Level of concern with rigour and reproducibility:",
-              choices = c(
-                "Less concerned",
-                "Neutral",
-                "More concerned"),
-              selected = c(
-                "Less concerned",
-                "Neutral",
-                "More concerned"), multiple = TRUE),
+              choices = c("very much less concerned", "less concerned", "somewhat less concerned", "as concerned as a typical researcher in my field", "somewhat more concerned", "more concerned", "Very much more concerned", "missing"),
+              selected = c("very much less concerned", "less concerned", "somewhat less concerned", "as concerned as a typical researcher in my field", "somewhat more concerned", "more concerned", "Very much more concerned", "missing"),
+              multiple = TRUE),
+  # Filter by programming language used for the analysis
   selectInput("language", label = "Programming language used for the analysis:",
-              choices = c(
-                "R",
-                "Stata",
-                "SPSS",
-                "SAS",
-                "Python",
-                "Mplus",
-                "Bash",
-                "MATLAB",
-                "Nextflow",
-                "plink2",
-                "irrelevant"
-              ),
-              selected = c(
-                "R",
-                "Stata",
-                "SPSS",
-                "SAS",
-                "Python",
-                "Mplus",
-                "Bash",
-                "MATLAB",
-                "Nextflow",
-                "plink2",
-                "irrelevant"
-              ), multiple = TRUE)
+              choices = c("R", "Stata", "SPSS", "SAS", "Python", "Mplus", "Bash", "MATLAB", "Nextflow", "plink2", "irrelevant"),
+              selected = c("R", "Stata", "SPSS", "SAS", "Python", "Mplus", "Bash", "MATLAB", "Nextflow", "plink2", "irrelevant"),
+              multiple = TRUE),
+  br(),
+  # Filtered by the methods used by the respondents
+  ## BLinded analysis
+  selectInput("blinded", label = "Blinded data analyst",
+              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              multiple = TRUE),
+  ## Preregistration
+  selectInput("prereg", label = "Preregistration",
+              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              multiple = TRUE),
+  ## Script
+  selectInput("script", label = "Analysis script",
+              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              multiple = TRUE),
+  ## Confirmatory analysis
+  selectInput("confirmatory", label = "Confirmatory analysis",
+              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              multiple = TRUE),
+  ## Exploratory analysis
+  selectInput("exploratory", label = "Exploratory analysis",
+              choices = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              selected = c("Never or almost never", "Sometimes", "About half the time", "Most of the time", "Always or almost always", "Missing", "I don't understand the question"),
+              multiple = TRUE)
   )
 
-# Filter data
+# Filter data based on the settings
 filtered_data <- reactive({
-  concerned_options <-list(
-    "Less concerned" =  c("very much less concerned", "less concerned", "somewhat less concerned"),
-    "Neutral" =  c("as concerned as a typical researcher in my field"),
-    "More concerned" = c("somewhat more concerned", "more concerned", "Very much more concerned")
-  )
+  # Create a filter for concerned and methods that only includes response levels
+  # TODO: Not the most elegant solution...
+  concerned_filter <- input$concerned[input$concerned != "missing"]
+  blinded_filter <- input$blinded[input$blinded != "Missing"]
+  prereg_filter <- input$prereg[input$prereg != "Missing"]
+  script_filter <- input$script[input$script != "Missing"]
+  confirmatory_filter <- input$confirmatory[input$confirmatory != "Missing"]
+  exploratory_filter <- input$exploratory[input$exploratory != "Missing"]
 
-  concerned_filter <- unlist(concerned_options[input$concerned], use.names = FALSE)
-
+  res <-
     processed %>% 
     filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
-    filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
+    # Filter concerned by response levels and keep all missing responses for now
     filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
+    # Filter methods by levels and keep all missing responses for now
+    filter(method_blind %in% blinded_filter | is.na(method_blind)) %>% 
+    filter(method_preregistered %in% prereg_filter | is.na(method_preregistered)) %>% 
+    filter(method_script %in% script_filter | is.na(method_script)) %>% 
+    filter(method_confirmatory %in% confirmatory_filter | is.na(method_confirmatory)) %>% 
+    filter(method_exploratory %in% exploratory_filter | is.na(method_exploratory)) %>% 
     filter(str_detect(programming_language, paste(input$language, collapse="|")) | is.na(programming_language))
+
+  # Filter missing responses from concerned and method if specified
+  # TODO: Not the most elegant solution...
+  if ("missing" %ni% input$concerned) res <- dplyr::filter(res, !is.na(concerned))
+  if ("Missing" %ni% input$blinded) res <- dplyr::filter(res, !is.na(method_blind))
+  if ("Missing" %ni% input$prereg) res <- dplyr::filter(res, !is.na(method_preregistered))
+  if ("Missing" %ni% input$script) res <- dplyr::filter(res, !is.na(method_script))
+  if ("Missing" %ni% input$confirmatory) res <- dplyr::filter(res, !is.na(method_confirmatory))
+  if ("Missing" %ni% input$exploratory) res <- dplyr::filter(res, !is.na(method_exploratory))
+
+  # Return filtered data
+  return(res)
   })
-# select(filtered_data(), concerned)
-# shiny::renderTable(filtered_data())
-shiny::renderText({nrow(filtered_data())})
 ```
+
+# Dataset
+
+```{r echo=FALSE}
+# Get the index of open ended question for condensing the long text
+condensed_cols <- which(colnames(processed) %in% c("unsure_explain", "drawback_ecaw", "suggestions", "comments", "contact")) - 1
+
+# Render the dataset as a pretty table
+DT::renderDataTable(filtered_data(), rownames = FALSE, options = list(
+  pageLength = 5,
+  lengthMenu = c(5, 10, 15, 20),
+  scrollX = TRUE,
+  autoWidth = TRUE,
+  columnDefs = list(
+    list(
+      className = "dt-center", targets = 0
+    ),
+    list(
+      targets = condensed_cols,
+      render = htmlwidgets::JS(
+        "function(data, type, row, meta) {",
+        "return type === 'display' && data != null && data.length > 25 ?",
+        "'<span title=\"' + data + '\">' + data.substr(0, 25) + '...</span>' : data;",
+        "}"
+      )
+    )
+  )
+))
+```
+
+# Results from the manuscript
 
@@
           "Somewhat agree",
-    "Strongly agree"))
-    )) %>%
+          "Strongly agree"
+        ))
+      )
+    ) %>%
     rename(
@@
 typically_plot <- reactive({
-  plot(typically_plot_data(), digits = 1, ordered = FALSE, legend.position = "right") +
-  labs(title = "Typically, studies that analyze preexisting observational datasets (such as the ALSPAC dataset) are...")
+  plot(typically_plot_data(), digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.5) +
+    scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+    labs(title = str_wrap("Typically, studies that analyze preexisting observational datasets (such as the ALSPAC dataset) are...", width = 50)) +
+    theme(
+      title = element_text(size = 14),
+      axis.text.y = element_text(size = 14),
+      axis.text.x = element_text(size = 15),
+      axis.title.x = element_blank(),
+      legend.text = element_text(size = 13.5),
+      legend.title = element_blank()
+    )
 })
@@
           "Somewhat more",
-          "Much more"))
-    )) %>%
+          "Much more"
+        ))
+      )
+    ) %>%
     rename(
@@
 ecaw_plot <- reactive({
-  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right") +
-  labs(title = "Compared to a typical study using preexisting observational data, a study using an ECAW would be...")
+  plot(ecaw_plot_data(), digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.5) +
+    scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+    labs(title = str_wrap("Compared to a typical study using preexisting observational data, a study using an ECAW would be...", width = 50)) +
+    theme(
+      title = element_text(size = 14),
+      axis.text.y = element_text(size = 14),
+      axis.text.x = element_text(size = 15),
+      axis.title.x = element_blank(),
+      legend.text = element_text(size = 13.5),
+      legend.title = element_blank(),
+      legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2.3, unit = "cm")
+    )
 })
@@
 # Join the two plots together to make one figure
-renderPlot({ typically_plot + ecaw_plot + plot_layout(ncol = 1, nrow = 2) })
+renderPlot({
+  typically_plot() + ecaw_plot() + plot_layout(ncol = 1, nrow = 2)
+})
 ```
 
-```{r}
+---
+
+```{r echo=FALSE}
+# Used methods figures
+method_plot_data <- reactive({
+  filtered_data() %>%
+    select(starts_with("method_")) %>%
+    pivot_longer(
+      cols = everything(),
+      names_to = "variable",
+      values_to = "scale",
+    ) %>%
+    separate(variable, into = c("name_prefix", "type")) %>%
+    group_by(type) %>%
+    count(scale) %>%
+    ungroup() %>%
+    replace_na(list(scale = "Missing")) %>%
+    mutate(
+      scale = factor(scale, levels = c(
+        "Never or almost never",
+        "Sometimes",
+        "About half the time",
+        "Most of the time",
+        "Always or almost always",
+        "Missing",
+        "I don't understand the question"
+      )),
+      type = case_when(
+        type == "blind" ~ "Blind the data analyst",
+        type == "confirmatory" ~ "Contain confirmatory analysis",
+        type == "exploratory" ~ "Contain exploratory analysis",
+        type == "preregistered" ~ "Are preregistered",
+        type == "script" ~ "Share analysis scripts",
+      ),
+      type = factor(type, levels = c(
+        "Contain exploratory analysis",
+        "Contain confirmatory analysis",
+        "Blind the data analyst",
+        "Share analysis scripts",
+        "Are preregistered"
+      ))
+    ) %>%
+    group_by(type) %>%
+    tidyr::complete(scale, fill = list(n = 0)) %>%
+    mutate(
+      percentage = n / sum(n)
+    ) %>%
+    ungroup()
+})
+
+renderPlot({
+  method_plot_data() %>%
+    # filter(scale %ni% c("Missing", "I don't understand the question")) %>%
+    ggplot() +
+    aes(
+      x = percentage,
+      y = type,
+      fill = scale
+    ) +
+    geom_bar(position = "stack", stat = "identity") +
+    scale_x_continuous(
+      limits = c(0, 1),
+      labels = scales::label_percent()
+    ) +
+    scale_fill_viridis(option = "A", discrete = TRUE) +
+    scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) +
+    labs(
+      x = "Percentage",
+      fill = "Response",
+      title = "The studies using preexisting observational data that I am involved in..."
+    ) +
+    papaja::theme_apa() +
+    theme(
+      title = element_text(size = 17),
+      axis.text = element_text(size = 15),
+      axis.title.x = element_blank(),
+      axis.title.y = element_blank(),
+      legend.text = element_text(size = 14),
+      legend.title = element_blank(),
+      legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
+    )
+})
+```
+
+---
+
+```{r echo=FALSE}
+# Prepare plot data
+alspac_plot_data <- reactive({
+  filtered_data() %>%
+    select(contains("alspac_")) %>%
+    # Transform not wanted response values to NA
+    # likert::likert drops NA values silently
+    # When var transformed to factor these values would be transformed to NA
+    # Automatically but I try to be explicit
+    mutate(
+      across(
+        everything(),
+        ~ case_when(
+          . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
+          TRUE ~ .
+        )
+      )
+    ) %>%
+    mutate(
+      across(
+        .fns = ~ factor(., levels = c(
+          "Strongly disagree",
+          "Somewhat disagree",
+          "Neither agree nor disagree",
+          "Somewhat agree",
+          "Strongly agree"
+        ))
+      )
+    ) %>%
+    rename(
+      `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
+      `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
+      `ALSPAC should run a study on ECAWs.` = alspac_study,
+      `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
+    )
+})
+
+# Create likert package data not including the missing values
+alspac_plot_likert_data <- reactive({
+  likert(alspac_plot_data())
+})
+
+# Create figure
+renderPlot({
+  plot(alspac_plot_likert_data(), digits = 1, ordered = TRUE, group.order = names(alspac_plot_data()), text.size = 5.5) +
+    scale_x_discrete(labels = function(x) str_wrap(x, width = 25), limits = rev(names(alspac_plot_data))) +
+    labs(title = str_wrap("Thinking about a study you may run with ALSPAC data (or one that you have recently run)..."), width = 25) +
+    theme(
+      axis.title.x = element_blank(),
+      title = element_text(size = 14),
+      axis.text.y = element_text(size = 14),
+      axis.text.x = element_text(size = 16),
+      legend.text = element_text(size = 15),
+      legend.title = element_blank(),
+      plot.margin = margin(t = 0, r = 0, b = 3.5, l = 0, unit = "lines"),
+      legend.position = c(.35,-.2),
+      legend.direction = "horizontal"
+    )
+})
+```
+
+# Results from supplementary materials C
+
+```{r echo=FALSE}
 renderPlot({
@@
       y = "Count",
-    title = "Approximately how many studies have you published using a preexisting observational dataset (e.g., the ALSPAC dataset)?"
+      title = str_wrap("Approximately how many studies have you published using a preexisting observational dataset (e.g., the ALSPAC dataset)?", width = 70)
     ) +
-  papaja::theme_apa()
+    papaja::theme_apa() +
+    theme(
+      title = element_text(size = 15),
+      axis.text = element_text(size = 15),
+      axis.title = element_text(size = 17)
+    )
 })
 ```
+
+---
+
+```{r echo=FALSE}
+# Programming languages table
+renderTable({
+  denominator <- nrow(filtered_data())
+
+  filtered_data() %>%
+    rename(lang = programming_language) %>%
+    separate_rows(lang, sep = ",") %>%
+    count(lang) %>%
+    replace_na(list(lang = "Missing")) %>%
+    # Removing irrelevant
+    filter(lang != "irrelevant") %>%
+    ungroup() %>%
+    mutate(
+      `Percentage of respondents` = round(n / denominator * 100, 2)
+    ) %>%
+    arrange(desc(n)) %>%
+    rename(
+      `Programming language` = lang,
+      `N` = n
+    )
+})
+```
+
+---
+
+```{r echo=FALSE}
+# Concerned figure data preparation
+concerned_plot_data <- reactive({
+  filtered_data() %>%
+    select(concerned) %>%
+    mutate(
+      concerned = tolower(concerned),
+      concerned = factor(concerned, levels = c(
+        "very much less concerned",
+        "less concerned",
+        "somewhat less concerned",
+        "as concerned as a typical researcher in my field",
+        "somewhat more concerned",
+        "more concerned",
+        "very much more concerned"
+      ))
+    ) %>%
+    filter(!is.na(concerned))
+})
+
+# Create likert package data not including the missing values
+concerned_likert_plot_data <- reactive({
+  likert(concerned_plot_data())
+})
+
+# Create figure
+renderPlot({
+  plot(concerned_likert_plot_data(), digits = 1, text.size = 6, group.order = names(concerned_plot_data())) +
+    scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+    labs(title = str_wrap("Compared to what you think of as a typical researcher who uses preexisting observational data in your field, how concerned are you with research trustworthiness, bias, rigour, and reproducibility ...", width = 100)) +
+    theme(
+      title = element_text(size = 14),
+      axis.text.y = element_blank(),
+      axis.text.x = element_text(size = 15),
+      axis.title.x = element_blank(),
+      legend.text = element_text(size = 11),
+      legend.title = element_blank()
+    )
+})
+```

### 2023-03-18 manuscript/manuscript.tex
made interactive report prettier, added more setting options

@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/numStudies-1}
+{\centering \includegraphics[width=1\linewidth]{figs/numStudiesPlot-1}
 
@@
 
-\caption{Caption goes here...}\label{fig:numStudies}
+\caption{Caption goes here...}\label{fig:numStudiesPlot}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-3}What programming language or software do you use for your analyses of preexisting observational data?}
+\caption{\label{tab:languageTable}What programming language or software do you use for your analyses of preexisting observational data?}
 \resizebox{\linewidth}{!}{
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/concerned-1}
+{\centering \includegraphics[width=1\linewidth]{figs/concernedPlot-1}
 
@@
 
-\caption{Caption goes here...}\label{fig:concerned}
+\caption{Caption goes here...}\label{fig:concernedPlot}
 \end{figure}

### 2023-03-18 manuscript/supplementary_materials.Rmd
made interactive report prettier, added more setting options

--- 2023-03-17 manuscript/supplementary_materials.Rmd
@@
 
-```{r numStudies, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=6, fig.height=3.2}
+```{r numStudiesPlot, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=6, fig.height=3.2, fig.path='figs/', dev=c('png', 'pdf')}
 # Create text for missing responses
@@
 
-```{r}
+```{r languageTable}
 # Programming languages table
@@
 
-```{r concerned, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4}
+```{r concernedPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4, fig.path='figs/', dev=c('png', 'pdf')}
 # Concerned figure data preparation
@@
 plot(concerned_likert_plot_data, digits = 1, text.size = 6, group.order = names(concerned_plot_data)) +
-  labs(title = "Compared to what you think of as a typical researcher who uses preexisting observational\ndata in your field, how concerned are you with\nresearch trustworthiness, bias, rigour, and reproducibility ...") +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+  labs(title = str_wrap("Compared to what you think of as a typical researcher who uses preexisting observational data in your field, how concerned are you with research trustworthiness, bias, rigour, and reproducibility ...", width = 100)) +
   theme(

### 2023-03-18 README.Rmd
updated readme with fig folder description

@@
   * `ecaw_source_raw_preprocessing.Rmd` file contains the code necessary for the transformation of the source data (the datafile downloaded directly from Qualtrics as is) to the raw datafile (datafile with standard naming).
-We also made sure that the raw datafile does not contain any information that can be used to indetify any of the respondents.
+We also made sure that the raw datafile does not contain any information that can be used to identify any of the respondents.
   * `ecaw_raw_processed_preprocessing.Rmd` file contains the code that cleans the dataset and transforms is into tidy format ready for the analysis.
@@
 The `manuscript/` folder contains the manuscript and supplementary materials documents with the analysis code embedded in them.
-The folder also includes the `interactive_report.Rmd` file, which includes an interactive document for reanalysing our results with different subsets of the data.
+The folder also includes the `interactive_report.Rmd` file, which includes an interactive document for reanalyzing our results with different subsets of the data.
 To run the interactive document locally on your computer, click the _Run document_ button in Rstudio or run the `rmarkdown::run(here::here("manuscript/interactive_report.Rmd"))` function in your R console.
+The `figs/` subfolder contains the rendered figures from the manuscript and supplementary materials in .pdf and .png formats.
 

### 2023-03-23 README.Rmd
updated manuscript based on contributor feedback and restructured the folders to deploy the interactive plot

@@
 
+The `root` folder includes the `interactive_report.Rmd` file, which includes an interactive document for reanalyzing our results with different subsets of the data.
+To run the interactive document locally on your computer, click the _Run document_ button in Rstudio or run the `rmarkdown::run("interactive_report.Rmd")` function in your R console.
+
 The `data/` folder contains all the datafiles and the corresponding codebooks that are needed to reproduce the results of the project.
@@
 The `manuscript/` folder contains the manuscript and supplementary materials documents with the analysis code embedded in them.
-The folder also includes the `interactive_report.Rmd` file, which includes an interactive document for reanalyzing our results with different subsets of the data.
-To run the interactive document locally on your computer, click the _Run document_ button in Rstudio or run the `rmarkdown::run(here::here("manuscript/interactive_report.Rmd"))` function in your R console.
 The `figs/` subfolder contains the rendered figures from the manuscript and supplementary materials in .pdf and .png formats.

### 2023-03-23 interactive_report.Rmd
updated manuscript based on contributor feedback and restructured the folders to deploy the interactive plot

--- 2023-03-18 manuscript/interactive_report.Rmd
@@
 runtime: shiny
+resource_files:
+- R/utils.R
 ---
@@
 renderPlot({
-  plot(alspac_plot_likert_data(), digits = 1, ordered = TRUE, group.order = names(alspac_plot_data()), text.size = 5.5) +
-    scale_x_discrete(labels = function(x) str_wrap(x, width = 25), limits = rev(names(alspac_plot_data))) +
+  plot(alspac_plot_likert_data(), digits = 1, text.size = 5.5) +
+    scale_x_discrete(labels = function(x) str_wrap(x, width = 25), limits = rev(names(alspac_plot_data()))) +
     labs(title = str_wrap("Thinking about a study you may run with ALSPAC data (or one that you have recently run)..."), width = 25) +

### 2023-03-23 manuscript/manuscript.Rmd
updated manuscript based on contributor feedback and restructured the folders to deploy the interactive plot

@@
 
-```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy[^3] (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible[^4] (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 2, top panel).
+At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 2, bottom panel).
+
+```{r typicallyEcawPlot, warning=FALSE, fig.cap="(ref:typicallyEcawPlotCaption)", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
 # Create typically trustworthy and reproducible plot
@@
 # Calculate missing responses
-ecaw_missing_text <-
+ecaw_missing <-
   processed %>%
@@
     n = n()
-  ) %>%
-  ungroup() %>%
-  mutate(text = paste(value, "=", n))
+  )
 
@@
 
-Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy[^3] (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%) and reproducible[^4] (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%) (Figure 1A).
-At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%) compared to a typical study using preexisting observational data (Figure 1B).
+(ref:typicallyEcawPlotCaption) __Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+That they merit trust”.
+The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.” For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange.
+The number in the center of the data bar (gray) indicates the percentage of neutral responses.
+The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
+For the bottom panel we excluded the missing responses (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "Missing"), n)`) and responses of “I don't understand the question” (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "I don't understand the question"), n)`).
 
 ```{r methodDescriptives, include=FALSE}
-confirmatory <- level_percentage(
-  processed,
-  method_confirmatory,
-  c(
+# Define the response level order for these variables
+method_levels <- c(
   "Never or almost never",
@@
   "Most of the time",
-    "Always or almost always"
-  ),
-  "I don't understand the question"
+  "Always or almost always",
+  "I don't understand the question",
+  "Missing"
+)
+
+# Calculate the fequency and proportion of responses for each response level
+confirmatory <- level_percentage(
+  processed,
+  method_confirmatory,
+  method_levels,
+  exclude_missing = FALSE
 ) %>%
-  filter(method_confirmatory != "Never or almost never") %>%
-  summarise(sum_percentage = sum(percentage)) %>%
-  pull(sum_percentage)
+  filter(method_confirmatory %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
+  summarise(
+    n = sum(n),
+    # Since n_sum should be the same for the whole dataset we can use unique to return one value
+    n_sum = unique(n_sum),
+    percentage = sum(percentage))
 
@@
   method_exploratory,
-  c(
-    "Never or almost never",
-    "Sometimes",
-    "About half the time",
-    "Most of the time",
-    "Always or almost always"
-  ),
-  "I don't understand the question"
+  method_levels,
+  exclude_missing = FALSE
 ) %>%
-  dplyr::filter(method_exploratory != "Never or almost never") %>%
-  summarise(sum_percentage = sum(percentage)) %>%
-  pull(sum_percentage)
+  dplyr::filter(method_exploratory  %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
+  summarise(
+    n = sum(n),
+    # Since n_sum should be the same for the whole dataset we can use unique to return one value
+    n_sum = unique(n_sum),
+    percentage = sum(percentage))
 
@@
   method_preregistered,
-  c(
-    "Never or almost never",
-    "Sometimes",
-    "About half the time",
-    "Most of the time",
-    "Always or almost always"
-  ),
-  "I don't understand the question"
+  method_levels,
+  exclude_missing = FALSE
 )
@@
   method_blind,
-  c(
-    "Never or almost never",
-    "Sometimes",
-    "About half the time",
-    "Most of the time",
-    "Always or almost always"
-  ),
-  "I don't understand the question"
+  method_levels,
+  exclude_missing = FALSE
 )
@@
   method_script,
-  c(
-    "Never or almost never",
-    "Sometimes",
-    "About half the time",
-    "Most of the time",
-    "Always or almost always"
-  ),
-  "I don't understand the question"
+  method_levels,
+  exclude_missing = FALSE
 )
@@
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%) (Figure 2A).
-About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%) (Figure 2B).
-`r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% reported that they never or almost never blind the data analyst (Figure 2C).
-Almost all respondents answered that they use both exploratory (`r exploratory`%) and confirmatory (`r confirmatory`%) analyses at least sometimes (Figure 2D-E).
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2A).
+About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script, method_script == "Never or almost never") %>% pull(n)`/`r filter(script, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script, method_script == "Sometimes") %>% pull(n)`/`r filter(script, method_script == "Sometimes") %>% pull(n_sum)`) (Figure 3).
+`r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst (Figure 3).
+Almost all respondents answered that they use both exploratory (`r pull(exploratory, percentage)`%; `r pull(exploratory, n)`/`r pull(exploratory, n_sum)`) and confirmatory (`r pull(confirmatory, percentage)`%; `r pull(confirmatory, n)`/`r pull(confirmatory, n_sum)`) analyses at least sometimes (Figure 3).
 
-```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
+```{r methodPlot, warning=FALSE, fig.cap="(ref:methodPlotCaption)", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
 # Used methods figures
@@
 
+(ref:methodPlotCaption) __Responses to survey questions about the research practices of participants.__
+
 ```{r aslpacDescriptives, include=FALSE, warning=FALSE, message=FALSE}
@@
 
-`r pull(filter(willing, support == "positive"), percentage)`% of respondents agreed (versus `r pull(filter(willing, support == "negative"), percentage)`% who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3).
-`r pull(filter(opt, support == "positive"), percentage)`% agreed (`r pull(filter(opt, support == "negative"), percentage)`% disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
-`r pull(filter(study, support == "positive"), percentage)`% agreed (`r pull(filter(study, support == "negative"), percentage)`% disagreed) that ALSPAC should run a study on ECAWs.
-`r pull(filter(prefer, support == "positive"), percentage)`% agreed (`r pull(filter(prefer, support == "negative"), percentage)`% disagreed) that they would prefer using an ECAW than using typical preregistration.
+`r pull(filter(willing, support == "positive"), percentage)`% (`r pull(filter(willing, support == "positive"), n)`/`r pull(filter(willing, support == "positive"), n_sum)`) of respondents agreed (versus `r pull(filter(willing, support == "negative"), percentage)`%; `r pull(filter(willing, support == "negative"), n)`/`r pull(filter(willing, support == "negative"), n_sum)` who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3).
+`r pull(filter(opt, support == "positive"), percentage)`% (`r pull(filter(opt, support == "positive"), n)`/`r pull(filter(opt, support == "positive"), n_sum)`) agreed (`r pull(filter(opt, support == "negative"), percentage)`%; `r pull(filter(opt, support == "negative"), n)`/`r pull(filter(opt, support == "negative"), n_sum)` disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
+`r pull(filter(study, support == "positive"), percentage)`% (`r pull(filter(study, support == "positive"), n)`/`r pull(filter(study, support == "positive"), n_sum)`) agreed (`r pull(filter(study, support == "negative"), percentage)`%; `r pull(filter(study, support == "negative"), n)`/`r pull(filter(study, support == "negative"), n_sum)` disagreed) that ALSPAC should run a study on ECAWs.
+`r pull(filter(prefer, support == "positive"), percentage)`% (`r pull(filter(prefer, support == "positive"), n)`/`r pull(filter(prefer, support == "positive"), n_sum)`) agreed (`r pull(filter(prefer, support == "negative"), percentage)`%; `r pull(filter(prefer, support == "negative"), n)`/`r pull(filter(prefer, support == "negative"), n_sum)` disagreed) that they would prefer using an ECAW than using typical preregistration.
 
-```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
+```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotCaption)", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
 # Check if there is any missing and don't understand responses separately
-alspac_missing_text <-
+alspac_missing <-
   processed %>%
@@
   dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>%
-  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
+  mutate(
+    item = str_to_title(str_remove(item, "^[^_]*_")),
+    # We reorder it by making it a factor
+    item = factor(item, levels = c(
+      "Less_willing",
+      "Opt_in",
+      "Study",
+      "Prefer_ecaw"
+    )),
+    # By making it a factor we can complete missing levels in the data later
+    value = factor(value, levels = c(
+      "Missing",
+      "I don't understand the question",
+      "Unsure"
+    ))
+  ) %>%
   replace_na(list(value = "Missing")) %>%
@@
   ) %>% 
-  ungroup() %>%
-  mutate(text = paste(value, "=", n))
+  # Add n = 0 if no one choose a specific factor level
+  tidyr::complete(value, fill = list(n = 0))
 
@@
 # Create figure
-plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, text.size = 6) +
+plot(alspac_plot_likert_data, digits = 1, text.size = 6) +
   # TODO: have to check if this is not messing with the results
@@
 ```
+
+(ref:alspacPlotCaption) __Responses to survey questions about using ECAWs.__ The 4 questions we excluded missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`), responses of "I don't understand the question" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of "Unsure" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
 

### 2023-03-23 manuscript/manuscript.tex
updated manuscript based on contributor feedback and restructured the folders to deploy the interactive plot

@@
 
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+That they merit trust.''} (72\%; 77/107) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%; 84/107) (Figure 2, top panel).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 71/101) and \emph{more} reproducible (69\%; 70/102) compared to a typical study using preexisting observational data (Figure 2, bottom panel).
+
 \begin{figure}
@@
 
-\caption{Caption goes here...}\label{fig:typicallyEcawPlot}
+\caption{\textbf{Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+That they merit trust''.
+The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange.
+The number in the center of the data bar (gray) indicates the percentage of neutral responses.
+The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
+For the bottom panel we excluded the missing responses (n = 2; 2) and responses of ``I don't understand the question'' (n = 4; 3).}\label{fig:typicallyEcawPlot}
 \end{figure}
 
-Respondents generally agreed that studies that analyze preexisting observational datasets are trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
-That they merit trust.''} (72\%) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%) (Figure 1A).
-At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%) and \emph{more} reproducible (69\%) compared to a typical study using preexisting observational data (Figure 1B).
-
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (38\%), or sometimes (27\%) (Figure 2A).
-About half reported sharing their analysis scripts never or almost never (20\%), or sometimes (32\%) (Figure 2B).
-79\% reported that they never or almost never blind the data analyst (Figure 2C).
-Almost all respondents answered that they use both exploratory (95\%) and confirmatory (92\%) analyses at least sometimes (Figure 2D-E).
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (35\%; 37/107), or sometimes (24\%; 26/107) (Figure 2A).
+About half reported sharing their analysis scripts never or almost never (20\%; 21/107), or sometimes (31\%; 33/107) (Figure 3).
+75\% (80/107) reported that they never or almost never blind the data analyst (Figure 3).
+Almost all respondents answered that they use both exploratory (91\%; 98/107) and confirmatory (85\%; 92/107) analyses at least sometimes (Figure 3).
 
@@
 
-\caption{Caption goes here...}\label{fig:methodPlot}
+\caption{\textbf{Responses to survey questions about the research practices of participants.}}\label{fig:methodPlot}
 \end{figure}
 
-25\% of respondents agreed (versus 44\% who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3).
-53\% agreed (20\% disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
-55\% agreed (10\% disagreed) that ALSPAC should run a study on ECAWs.
-45\% agreed (22\% disagreed) that they would prefer using an ECAW than using typical preregistration.
+25\% (26/102) of respondents agreed (versus 44\%; 45/102 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3).
+53\% (50/95) agreed (20\%; 19/95 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
+55\% (53/97) agreed (10\%; 10/97 disagreed) that ALSPAC should run a study on ECAWs.
+45\% (43/95) agreed (22\%; 21/95 disagreed) that they would prefer using an ECAW than using typical preregistration.
 
@@
 
-\caption{Caption goes here...}\label{fig:alspacPlot}
+\caption{\textbf{Responses to survey questions about using ECAWs.} The 4 questions we excluded missing values (\emph{n} = 3; 3; 3; 3), responses of ``I don't understand the question'' (\emph{n} = 0; 4; 1; 1), and responses of ``Unsure'' (\emph{n} = 2; 5; 6; 8).}\label{fig:alspacPlot}
 \end{figure}

### 2023-03-23 preprocessing/ecaw_raw_processed_preprocessing.Rmd
updated manuscript based on contributor feedback and restructured the folders to deploy the interactive plot

@@
 
-```{r}
+```{r message=FALSE}
 raw <- readr::read_csv(here("data/raw/ecaw_raw_data.csv"))
@@
 
-We only include full responses in the final analysis as preregistered (see the full text of the preregistration [HERE](link)).
-There were `r dplyr::filter(raw, !finished)` respondents who did not finish the survey.
+We only include full responses in the final analysis as preregistered (see the full text of the preregistration [https://osf.io/g2fw5](https://osf.io/g2fw5)).
+There were `r nrow(dplyr::filter(raw, !finished))` respondents who did not finish the survey.
 
@@
 ```
+
+These responses will be included in an exploratory analysis that will be published in supplementary materials D.
 
@@
 
-```{r}
+```{r message=FALSE}
 codebook <-

### 2023-03-23 R/likert_plot_odd.R

@@ -1,192 +1,0 @@
-#' Horizontal stacked barplot for Likert-type data
-#' 
-#' The function creates a horizontal stacked barpolot for visualizing
-#' Likert-type data with an odd number of response levels. The function can
-#' take multiple scales and present them in one figure.
-#' 
-#' @section Source: The function is based on the code from \url{https://gist.github.com/trinker/0260a9dfdd9531f9b90d9fad2f7b4b12}
-#' 
-#' @param data dataframe containing the data for the plot
-#' @param title title of the plot
-#' @param levels name of the variable containing the levels. this variable should be a factor
-#' @param questions name of the variable that contains the questions to which the levels belong (multiple questions can be used)
-#' @param n name of the variable containing the number of responses corresponding to each level
-#' @param positive_side if `"lower"` lower levels than the middle are the positive side of the scale, while if `"upper"` levels bigger than the middle are the positive side
-#' @param levels_to_drop character vector containing the levels that should not be included e.g. "NA" or "I do not know" 
-#' @param limits the limits of the x axis
-#' @param text_push the distance between the x axis limits and the text along the y axis
-#' 
-#' @return The function returns a ggplot2 object.
-likert_plot_odd <- function(data, title, questions, levels, n, levels_to_drop, positive_side = c("lower", "upper"), limits = c(-0.5, 1), text_push = 0.1, legend_row = 1) {
-  .questions <- rlang::enquo(questions)
-  .levels <- rlang::enquo(levels)
-  
-  `%ni%` <- Negate(`%in%`)
-  
-  # Filter levels
-  likert_plot_data <-
-    data %>%
-    # Drop unwanted levels
-    filter({{levels}} %ni% levels_to_drop) %>% 
-    # Get levels as integer
-    mutate(levels_int = as.integer({{levels}}))
-  
-  # The number of levels should be odd
-  if (length(unique(likert_plot_data$levels_int)) %ni% c(3, 5, 7)) {
-    stop("The number of levels should 3, 5, or 7!")
-  }
-  
-  # Save color palette
-  if(length(unique(likert_plot_data$levels_int)) == 7) {
-    color_palette <- 
-      data.frame(
-        colors = c("#CD7F32", "#D5A16D", "#DDC3A9", "grey90", "#99D0CD", "#4EBCB5", "#03A89E"),
[…]

### 2023-03-23 R/utils.R

@@ -10,39 +10,42 @@ center_level <- function(var) (n_levels(var) - 1) / 2 + 1
 #' Calculate the percentage of responses of each factor level
 level_percentage <- function(data, var, levels, exclude = NULL, exclude_missing = TRUE) {
   res <-
-    data %>% 
+    data %>%
     filter(
-      # exclude specified answer option(s)
-      {{var}} %ni% exclude
+      # Exclude specified answer option(s)
+      {{ var }} %ni% exclude
     )
-  
-  # exclude missing responses
-  if(exclude_missing) {
+
+  # Exclude missing responses
+  if (exclude_missing) {
+    message(paste(nrow(filter(data, is.na({{ var }}))), "missing values were excluded."))
     res <-
-      res %>% 
-      filter(!is.na({{var}}))
+      res %>%
+      filter(!is.na({{ var }}))
   }
 
-  # res <-
-    res %>% 
+  res %>%
+    # Replace missing values with explicit text
     mutate(
+      # Replace missing values with explicit text
+      {{ var }} := replace_na({{ var }}, "Missing"),
       # Transform to factor with level order specified in levels arg
-      {{var}} := factor({{var}}, levels = levels)
-    ) %>% 
+      {{ var }} := factor({{ var }}, levels = levels)
+    ) %>%
     # Count the number of response on each factor level
-    count({{var}}) %>% 
+    count({{ var }}) %>%
     # Add n = 0 if no one choose a specific factor level
-    tidyr::complete({{var}}, fill = list(n = 0)) %>%
+    tidyr::complete({{ var }}, fill = list(n = 0)) %>%
     mutate(
       # Assign level id to each factor level in increasing order
-      levels_int = as.integer({{var}}),
+      levels_int = as.integer({{ var }}),
       # Calculate the number of all responses
       n_sum = sum(n),
       # Calculate the percentage for all factor level and round
       percentage = round(n / n_sum * 100)
-    ) %>% 
+    ) %>%
     # Reorder variables for cleaner output
-    select(levels_int, {{var}}, n, n_sum, percentage)
+    select(levels_int, {{ var }}, n, n_sum, percentage)
 }
 
 #' Calculate percentage for each support category

### 2023-03-23 preprocessing/ecaw_raw_processed_preprocessing.Rmd
updated data preprocessing files so the codebooks are done and the full dataset is saved for supplementary materials d

@@
 
-## Partial responses
-
-We only include full responses in the final analysis as preregistered (see the full text of the preregistration [https://osf.io/g2fw5](https://osf.io/g2fw5)).
-There were `r nrow(dplyr::filter(raw, !finished))` respondents who did not finish the survey.
-
-```{r}
-raw <- 
-  raw %>% 
-  dplyr::filter(finished)
-```
-
-These responses will be included in an exploratory analysis that will be published in supplementary materials D.
-
-At the end of the exclusion we had `r nrow(raw)` responses remaining for the analysis.
-
 It was optional to answer all of the questions in our survey, so it is possible to have missing responses even if the respondent finished the survey. We will keep these responses.
@@
 
+## Partial responses
+
+We only include full responses in the final analysis as preregistered (see the full text of the preregistration [https://osf.io/g2fw5](https://osf.io/g2fw5)).
+There were `r nrow(dplyr::filter(raw, !finished))` respondents who did not finish the survey.
+
+However, these responses will be included in an exploratory analysis that will be published in supplementary materials D. So we save them separately before exclusion.
+
+```{r}
+processed_all <- 
+  raw %>%
+  mutate(duration_in_mins = duration_in_seconds / 60) %>%
+  select(
+    -start_date,
+    -end_date,
+    -status,
+    -progress,
+    -finished,
+    -duration_in_seconds,
+    -recorded_date,
+    -distribution_channel,
+    -user_language
+  )
+
+readr::write_csv(processed_all, here("data/processed/ecaw_processed_all_data.csv"))
+```
+
+The full dataset that we are going to use to calculate the results in Supplement D includes `r nrow(processed_all)` responses.
+
+We create a simple codebook for this dataset.
+
+```{r}
+processed_all_codebook <-
+  read_csv(here("data/source/ecaw_221102_source_codebook.csv")) %>% 
+  select(-original_var_name) %>% 
+  rename(var_name = new_var_name) %>% 
+  filter(var_name %in% colnames(processed_all)) %>% 
+  # Adding new column descriptions
+  mutate(
+    has_missing = "True if respondent has any missing responses.",
+    duration_in_mins = "Time spent from opening the survey to closing it in minutes."
+  )
+
+readr::write_csv(processed_all_codebook, here("data/processed/ecaw_processed_all_codebook.csv"))
+```
+
+Excluding the respondents who did not finish the questionnaire.
+
+```{r}
+raw <- 
+  raw %>% 
+  dplyr::filter(finished)
+```
+
+At the end of the exclusion we had `r nrow(raw)` responses remaining for the analysis.
 There were `r filter(raw, has_missing) %>% nrow()` respondents who had at least one missing value in the main questions of interests.
@@
 
-For creating the codebook we will use the codebook created for the source data.
+For creating the codebook for the processed data we will use the codebook created for the source data.
 
@@
   rename(var_name = new_var_name) %>%
-  filter(var_name %in% colnames(raw))
+  filter(var_name %in% colnames(raw)) %>%
+  # Adding new column descriptions
+  mutate(
+    has_missing = "True if respondent has any missing responses.",
+    duration_in_mins = "Time spent from opening the survey to closing it in minutes."
+  )
 
 # Save the codebook
-readr::write_csv(codebook, here("data/raw/ecaw_raw_codebook.csv"))
+readr::write_csv(codebook, here("data/processed/ecaw_processed_codebook.csv"))
 ```

### 2023-03-27 manuscript/manuscript.Rmd
small fixes in the manuscript calculations and rerender

@@
     "Very much more concerned"
-  )
+  ),
+  exclude_missing = TRUE
 )
@@
 54 emails bounced, leaving 1094 emails that went through.
-The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r median(processed$duration_in_mins)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
+The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
 
@@
 They reported using the programming languages `r language_text` (Table S1)[^2].
-`r pull(filter(concerned, support == "positive"), percentage)`% of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); `r pull(filter(concerned, support == "negative"), percentage)`% reported being less concerned.
+`r pull(filter(concerned, support == "positive"), percentage)`% (`r pull(filter(concerned, support == "positive"), n)`/`r pull(filter(concerned, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); `r pull(filter(concerned, support == "negative"), percentage)`% (`r pull(filter(concerned, support == "negative"), n)`/`r pull(filter(concerned, support == "negative"), n_sum)`) reported being less concerned.
 
@@
 
-(ref:alspacPlotCaption) __Responses to survey questions about using ECAWs.__ The 4 questions we excluded missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`), responses of "I don't understand the question" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of "Unsure" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
+(ref:alspacPlotCaption) __Responses to survey questions about using ECAWs.__ For the 4 questions we excluded missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`), responses of "I don't understand the question" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of "Unsure" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
 

### 2023-03-27 manuscript/manuscript.tex
small fixes in the manuscript calculations and rerender

@@
 For example, we received one email reply stating that the recipient hasn't been active in research for 30 years.
-Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.22 minutes (IQR: 4.30 to 13.10).
+Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.20 minutes (IQR: 4.30 to 13.10).
 
@@
 They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1)\footnote{Participants could select multiple responses to this survey question.}.
-61\% of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% reported being less concerned.
+61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% (6/101) reported being less concerned.
 

### 2023-04-04 manuscript/manuscript.Rmd
Code check (Robby)
removed participants, edited codebook generation, minor tweaks to comments.

@@
 processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
+processed_all <- read_csv(here::here("data/processed/ecaw_processed_all_data.csv"))
 ```
@@
 54 emails bounced, leaving 1094 emails that went through.
-The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
+The survey was completed `r nrow(processed)` times and partially completed `r nrow(processed_all) - nrow(processed) ` times, leading to a response rate of `r round(nrow(processed_all) / 1094 * 100)`% for complete surveys and `r round((nrow(processed_all) - nrow(processed)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
 
@@
 
-# Calculate the fequency and proportion of responses for each response level
+# Calculate the frequency and proportion of responses for each response level
 confirmatory <- level_percentage(

### 2023-04-04 manuscript/manuscript.tex
Code check (Robby)
removed participants, edited codebook generation, minor tweaks to comments.

--- 2023-03-23 manuscript/manuscript.tex
@@
 54 emails bounced, leaving 1094 emails that went through.
-The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for complete surveys and 2\% for incomplete surveys.\footnote{The ALSPAC mailing list has been active for \textgreater30 years and may contain email addresses that are no longer monitored.
+The survey was completed 103 times and partially completed 20 times, leading to a response rate of 11\% for complete surveys and 2\% for incomplete surveys.\footnote{The ALSPAC mailing list has been active for \textgreater30 years and may contain email addresses that are no longer monitored.
 For example, we received one email reply stating that the recipient hasn't been active in research for 30 years.
-Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.22 minutes (IQR: 4.30 to 13.10).
+Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.40 minutes (IQR: 4.60 to 13.10).
 
-Respondents published a median of NA (IQR 2 to 25) studies using preexisting observational data (Figure S1).
+Respondents published a median of NA (IQR 2 to 26.20) studies using preexisting observational data (Figure S1).
 They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1)\footnote{Participants could select multiple responses to this survey question.}.
-61\% of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% reported being less concerned.
+62\% (62/100) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% (6/100) reported being less concerned.
 
@@
 Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
-That they merit trust.''} (72\%; 77/107) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%; 84/107) (Figure 2, top panel).
-At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 71/101) and \emph{more} reproducible (69\%; 70/102) compared to a typical study using preexisting observational data (Figure 2, bottom panel).
+That they merit trust.''} (72\%; 74/103) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%; 81/103) (Figure 2, top panel).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using preexisting observational data (Figure 2, bottom panel).
 
@@
 The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
-For the bottom panel we excluded the missing responses (n = 2; 2) and responses of ``I don't understand the question'' (n = 4; 3).}\label{fig:typicallyEcawPlot}
+For the bottom panel we excluded the missing responses (n = ; ) and responses of ``I don't understand the question'' (n = 3; 2).}\label{fig:typicallyEcawPlot}
 \end{figure}
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (35\%; 37/107), or sometimes (24\%; 26/107) (Figure 2A).
-About half reported sharing their analysis scripts never or almost never (20\%; 21/107), or sometimes (31\%; 33/107) (Figure 3).
-75\% (80/107) reported that they never or almost never blind the data analyst (Figure 3).
-Almost all respondents answered that they use both exploratory (91\%; 98/107) and confirmatory (85\%; 92/107) analyses at least sometimes (Figure 3).
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2A).
+About half reported sharing their analysis scripts never or almost never (20\%; 21/103), or sometimes (32\%; 33/103) (Figure 3).
+77\% (79/103) reported that they never or almost never blind the data analyst (Figure 3).
+Almost all respondents answered that they use both exploratory (93\%; 96/103) and confirmatory (87\%; 90/103) analyses at least sometimes (Figure 3).
 
@@
 
-25\% (26/102) of respondents agreed (versus 44\%; 45/102 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3).
-53\% (50/95) agreed (20\%; 19/95 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
-55\% (53/97) agreed (10\%; 10/97 disagreed) that ALSPAC should run a study on ECAWs.
-45\% (43/95) agreed (22\%; 21/95 disagreed) that they would prefer using an ECAW than using typical preregistration.
+26\% (26/101) of respondents agreed (versus 45\%; 45/101 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3).
+53\% (50/94) agreed (20\%; 19/94 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
+55\% (53/96) agreed (10\%; 10/96 disagreed) that ALSPAC should run a study on ECAWs.
+46\% (43/94) agreed (22\%; 21/94 disagreed) that they would prefer using an ECAW than using typical preregistration.
 
@@
 
-\caption{\textbf{Responses to survey questions about using ECAWs.} The 4 questions we excluded missing values (\emph{n} = 3; 3; 3; 3), responses of ``I don't understand the question'' (\emph{n} = 0; 4; 1; 1), and responses of ``Unsure'' (\emph{n} = 2; 5; 6; 8).}\label{fig:alspacPlot}
+\caption{\textbf{Responses to survey questions about using ECAWs.} For the 4 questions we excluded missing values (\emph{n} = 0; 0; 0; 0), responses of ``I don't understand the question'' (\emph{n} = 0; 4; 1; 1), and responses of ``Unsure'' (\emph{n} = 2; 5; 6; 8).}\label{fig:alspacPlot}
 \end{figure}
@@
 
-\hypertarget{supplementary-materials-c}{%
-\section{Supplementary materials C}\label{supplementary-materials-c}}
-
 \begin{figure}
@@
 \midrule
-R & 65 & 61\\
+R & 65 & 63\\
 \cmidrule{1-3}
-Stata & 48 & 45\\
+Stata & 48 & 47\\
 \cmidrule{1-3}
-SPSS & 17 & 16\\
+SPSS & 17 & 17\\
 \cmidrule{1-3}
-SAS & 15 & 14\\
+SAS & 15 & 15\\
 \cmidrule{1-3}
 Python & 6 & 6\\
-\cmidrule{1-3}
-Missing & 4 & 4\\
 \cmidrule{1-3}

### 2023-04-25 manuscript/manuscript.Rmd
updated rmarkdown manuscript according to the review

--- 2023-03-27 manuscript/manuscript.Rmd
@@
 ---
-title             : "A survey of registration practices among observational researchers using preexisting datasets"
+title             : "Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers"
 shorttitle        : "Explore and Confirm Analysis Workflow"
@@
   - name: Robert T. Thibault
-    affiliation: '1,5'
-    role:
-      - Conceptualization
-      - Data curation
-      - Formal analysis
-      - Funding acquisition
-      - Investigation
-      - Methodology
-      - Project administration
-      - Resources
-      - Supervision
-      - Validation
-      - Visualization
-      - Writing - original draft
-      - Writing - review & editing
+    affiliation: '1,2'
     corresponding: yes
@@
   - name: Marton Kovacs
-    affiliation: '2,6'
-    role:
-      - Data curation
-      - Formal analysis
-      - Software
-      - Validation
-      - Visualization
-      - Writing - review & editing
+    affiliation: '3,4'
     email: marton.balazs.kovacs@gmail.com
   - name: Tom E. Hardwicke
-    affiliation: '3'
-    role:
-      - Methodology
-      - Writing - review & editing
+    affiliation: '5'
   - name: Alexandra Sarafoglou
-    affiliation: '4'
-    role:
-      - Methodology
-      - Writing - review & editing
+    affiliation: '6'
   - name: John P. A. Ioannidis
-    affiliation: '4'
-    role:
-      - Methodology
-      - Writing - review & editing
+    affiliation: '1,7'
   - name: Marcus R. Munafò
-    affiliation: '1,7'
-    role:
-      - Conceptualization
-      - Methodology
-      - Supervision
-      - Writing - review & editing
+    affiliation: '2,8'
 affiliation:
@@
   - id: '2'
-    institution: Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest,
-      Hungary
+    institution: School of Psychological Science, University of Bristol.
   - id: '3'
+    institution: Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
+  - id: '4'
+    institution: Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
+  - id: '5'
     institution: Melbourne School of Psychological Sciences, University of Melbourne.
-  - id: '4'
+  - id: '6'
     institution: Department of Psychology, University of Amsterdam.
-  - id: '5'
-    institution: School of Psychological Science, University of Bristol.
-  - id: '6'
-    institution: Institute of Psychology, ELTE Eotvos Lorand University, Budapest,
-      Hungary
   - id: '7'
@@
     institution: MRC Integrative Epidemiology Unit at the University of Bristol.
-  - id: '9'
-    institution: Departments of Medicine, Epidemiology and Population Health, Biomedical
-      Data Science, and Statistics, Stanford University.
 abstract: |
-  placeholder for an abstract
+  \textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis.
+An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW).
+In this workflow, a data management organization initially provides access to only a subset of their dataset to researchers who request it.
+The researchers then prepare an analysis script based on the subset of data, upload the analysis script to a registry, and then receive access to the full dataset.
+ECAWs aim to achieve similar goals as preregistration, but make access to the full dataset contingent on compliance.
+The present survey aimed to garner information from the research community where ECAWs could be applied—employing the Avon Longitudinal Study of Parents and Children (ALSPAC) as a case example.
+\hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC’s transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}.
+We received 107 responses, for a 10% response rate.
+The results suggest that—at least among our sample of respondents—ECAWs hold the potential to serve their intended purpose and appear relatively acceptable.
+For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55% who agreed).
+However, as many as 25% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions.
+\hfill \break \hfill \break \textbf{Preregistration:} https://osf.io/g2fw5 Deviations from the preregistration are outlined in Supplementary Material A.
 
-keywords          : "keywords"
-wordcount         : "X"
-# bibliography      : "references.bib"
+keywords          : "Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)"
+wordcount         : "3546"
+bibliography      : "ecaw-ref.bib"
 floatsintext      : yes
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            :
+  papaja::apa6_pdf:
+    latex_engine: xelatex
+    extra_dependencies: "tcolorbox"
+header-includes   :
+  - \usepackage{tcolorbox}
+  - \usepackage{caption}
+  - \captionsetup[figure]{labelformat=empty}
 editor_options: 
@@
 library(viridis)
-# r_refs("references.bib")
+r_refs("ecaw-ref.bib")
 ```
@@
 # Introduction
+Many published research findings are non-reproducible and potentially false or misleading [@camerer_evaluating_2016; @errington_investigating_2021; @ioannidis_why_2005; @ioannidis_why_2008; @collaboration_estimating_2015].
+These shortcomings may stem from a combination of researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices.
+These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see @prasad_ending_2019).
+
+Some research disciplines have adopted standards to address these issues.
+For example, researchers conducting clinical trials regularly register outcome measures before enrolling participants, and they keep both participants and experimenters blind regarding group assignment.
+In cases where datasets already exist, however, researchers cannot easily adopt these practices.
+Researchers performing secondary data analyses generally have access to complete datasets without any blinding imposed on them.
+This may lead to exploration of diverse analyses and selective reporting based on the nature of the results.
+
+## Preregistration in the context of secondary data analyses
+Increasing the uptake and effectiveness of preregistration for secondary data analyses may require a different approach than it does for clinical trials.
+For example, the analytical space for secondary data analyses (e.g., of a longitudinal cohort dataset) is more vast than the analytical space for most clinical trials.
+Thus, the clinical trial standard of registering outcome measures, but no analysis plan, remains insufficient.
+Moreover, preregistration of secondary data analyses remains uncommon, and unlike for clinical trials (where the dataset doesn’t yet exist), there is no guarantee that researchers performing secondary data analyses haven’t accessed the dataset.
+To extend the benefits of preregistration to secondary data analyses, we propose a workflow that necessitates the preregistration of an executable analysis script before researchers can access the full dataset.
+
+## Explore and Confirm Analysis Workflow (ECAW)
+In this intervention we propose, a data management organization that controls access to a preexisting data set would first provide access to only a subset of the data a researcher requests.
+After the researcher prepares an analysis script based on the subset of data and uploads it to a registry—where it will be openly accessible and permanent (e.g., to the Open Science Framework Registry)—the data management organization then provides access to the full dataset (e.g., via a secure data environment) and the researcher can proceed as they wish.
+The exact implementation (e.g., whether the data management organization runs quality checks on the analysis scripts) would depend on the preferences of the data management organization and the research community using their dataset (see Box 1 for an hypothetical example).
+
+We call this process an Explore and Confirm Analysis Workflow (ECAW).
+Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly.
+ECAWs may increase the uptake of preregistration and provide assurance that the analyses were developed before the researchers observed the full dataset.
+Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., @bakker_ensuring_2020), ECAWs provide an executable analysis script.
+Researchers may also find ECAWs more agreeable than conventional preregistration because they allow for exploration.
+ECAWs would not fully solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration.
+
+In short, ECAWs give researchers access to enough data to develop an informed analysis, but without compromising the confirmatory nature of a final analysis on the complete dataset.
+
+\pagebreak
+
+\begin{tcolorbox}
+\textbf{Box 1. A hypothetical example of an ECAW in practice.}
+
+\includegraphics[width=1\textwidth]{./ecaw_workflow.png}
+{\fontsize{10pt}{10pt}\selectfont
+The overarching framework for ECAWs is depicted above.
+Within this framework, data management organizations would need to pre-specify several details for each step.
+Below, we provide a hypothetical example for illustrative purposes.
+The exact implementation of ECAWs would depend on the preferences of each data management organization and the community who uses their dataset.
+
+\begin{enumerate}
+  \item A research team submits to a data management organization (i) a paragraph describing the analyses they want to run, and (ii) a list of the variables they want to analyze.
+  \item The data management organization provides the research team with access to a subset of the data.
+For example, if the research team requests data for 15 variables collected from 10,000 participants, the data management organization will provide access to all 15 variables from a random subset of participants—say 1000.
+  \item The research team prepares an analysis script written in a common programming language (e.g., R or Stata) using the subset of data.
+They register this analysis script, the output from the script, and the paragraph they sent to the data management organization in Step 1, to www.osf.io/registries as an ‘Open-Ended Registration’.
+The researchers are free to run as many analyses as they would like on the subset of data.
+However, they should only register the analysis script they plan to use on the complete dataset.
+  \item The data management organization performs a basic quality check.
+They ensure that (i) the analysis script, the script output, and the paragraph are registered, and (ii) the output contains a clear number of itemized results (e.g., as clinicaltrials.gov does for outcome measures).
+If the quality check fails, the data management organization asks the researchers to update their registration until it passes.
+The research team is then given access to a dataset with all the variables requested for all participants.
+  \item The researchers can proceed as they wish.
+They can run the registered analysis script on the complete dataset, make adjustments to the analysis if desired, or not proceed at all.
+The data management organization does not perform any additional check on what analyses were run.
+Regardless of what the research team decides, a permanent version of the planned analysis is available on the OSF Registry.
+\end{enumerate}
+}
+\end{tcolorbox}
+
+\pagebreak
+
+## Examples of ECAW-like workflows
+One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (@sarafoglou_comparing_2023).
+The researchers recruited 120 teams to analyze a single observational dataset and had half the teams preregister a written analysis plan and the other half prepare an analysis plan by drafting an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded).
+Based on self-reports from the participants, the researchers found that two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.
+
+Some disciplines, such as particle physics, implement blind data analyses regularly (@klein_blind_2005; @maccoun_blind_2015).
+A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (@Forscher2020) and is currently being used for Registered Reports based on a large psychology dataset (@Schmidt2023).
+
+In medicine and clinical epidemiology, a similar workflow is implemented by the software platform OpenSAFELY (www.opensafely.org).
+This platform provides a dataset of simulated health records from which researchers can develop an analysis script.
+When ready, a researcher submits their analysis script which is automatically logged and made public in GitHub.
+The analysis then runs in a Trusted Research Environment (TRE) on data which is stored in the data centers where patients’ records already reside (i.e., the data is not copied or moved).
+This workflow keeps individual health records hidden while also documenting all analyses run on the real data.
+
+## Study objectives
+Here, we present a descriptive and exploratory survey study.
+We had no hypotheses, but we did have two specific objectives.
+(1) To gain insights on the opinions and practices of researchers who already use preexisting observational datasets, in regards to the trustworthiness and reproducibility of research.
+(2) To use these insights to inform future research—including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)—on how data management organizations can encourage rigorous and reproducible research practices.
 
 # Methods
+## Participants
+We sent an email to invite researchers on the mailing list for the UK-based Avon Longitudinal Study of Parents and Children (ALSPAC) to participate in an online survey (see Supplementary Material B).
+We partnered with ALSPAC because they manage an oft-requested dataset and expressed interest in studying ways to ensure the research stemming from their dataset is rigorous.
+ALSPAC is “a transgenerational prospective observational study investigating influences on health and development across the life course.
+It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.” (Boyd et al., 2013; Fraser et al., 2013).
+Thus, our invitation reached researchers that use observational data across the health and social sciences.
+The survey was open from 10 Oct 2022 to 1 Nov 2022.
+We sent two reminder emails, exactly one week and two weeks after the original email invitation.
+The mailing list comprises researchers whose email addresses were present on a proposal to access the ALSPAC dataset and included 1148 email addresses.
+
+## Survey
+The survey contained 6 blocks and is available at [https://osf.io/5h7gb](https://osf.io/5h7gb).
+We developed the survey with feedback from the Principal Investigator and Executive Director of ALSPAC (Nicholas Timpson and Kate Northstone).
+We aimed to include as few questions as possible (to encourage a high response rate) while still garnering substantive information on whether ECAWs are relevant and acceptable to ALSPAC users.
+
+Block 1 assessed the extent to which respondents believe that observational research using preexisting data is trustworthy and reproducible (2 questions).
+Block 2 asked respondents how often they use practices related to transparency and reducing researcher bias, including preregistration, blinded data analysis, and sharing analysis scripts (5 questions).
+Between Block 2 and Block 3, the survey described ECAWs.
+Block 3 assessed the extent to which respondents believe that ECAWs would make observational research using preexisting data more trustworthy and reproducible (2 questions).
+Block 4 directly asked respondents whether ALSPAC should run a study on ECAWs and whether they would participate (5 questions).
+Block 5 contained open-ended questions about the perceived benefits and drawbacks of ECAWs, reservations about ECAWs, whether additional incentives might be needed to use ECAW, and suggestions for other research practices or policies that data management organizations could implement to improve research quality (3 questions).
+Block 6 asked participants how concerned they are about research quality, how many relevant studies they have published, what software they use for data analysis, and for any additional comments (4 questions).
+All rating-scale questions contained the response option _“I don’t understand the question”_.
+The questions presented in Figure 3 also contained the response option _“Unsure”_.
+
+## Analyses
+We present the results for closed-ended questions as prevalence rates or counts in Figures 1-3 and Supplementary Material C.
+When reporting percentages, we collapse together all positive responses on the rating scales (e.g., _“strongly agree”_ and _“somewhat agree”_) as well as all negative responses.
+The total number of responses differ among questions due to the missing values and responses of _“I don’t understand the question”_ and _“Unsure”_.
+We narratively synthesize the responses to the open-ended questions in Table 1 (i.e., we provide a non-systematic qualitative summary).
 
 # Results
+## Survey completion
+Of 1148 emails sent, 1094 went through and 54 bounced.
+The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
+The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action.
+Thus, their list likely includes several email addresses that are no longer monitored.
+For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years.
+It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions).
+These two factors may have deflated the response rate.} The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
+This manuscript presents the results for complete surveys.
+Supplementary Material D presents the results with partially complete surveys included.
+
 ## Participants
@@
 
-We invited the ALSPAC mailing list to participate, which included 1148 email addresses.
-54 emails bounced, leaving 1094 emails that went through.
-The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
-
-Respondents published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)`) studies using preexisting observational data (Figure S1).
-They reported using the programming languages `r language_text` (Table S1)[^2].
-`r pull(filter(concerned, support == "positive"), percentage)`% (`r pull(filter(concerned, support == "positive"), n)`/`r pull(filter(concerned, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); `r pull(filter(concerned, support == "negative"), percentage)`% (`r pull(filter(concerned, support == "negative"), n)`/`r pull(filter(concerned, support == "negative"), n_sum)`) reported being less concerned.
-
-[^1]: The ALSPAC mailing list has been active for >30 years and may contain email addresses that are no longer monitored.
-For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years.
-Excluding these email addresses would increase the response rate, but we do not know by how much.
-[^2]: Participants could select multiple responses to this survey question.
-[^3]: The survey defined trustworthy as: “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust.”
-[^4]: The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.”
+Respondents had published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)`) studies using preexisting observational data (Supplementary Figure C1).
+They reported using the following programming languages or software packages: `r language_text` (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}.
+`r pull(filter(concerned, support == "positive"), percentage)`% (`r pull(filter(concerned, support == "positive"), n)`/`r pull(filter(concerned, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; `r pull(filter(concerned, support == "negative"), percentage)`% (`r pull(filter(concerned, support == "negative"), n)`/`r pull(filter(concerned, support == "negative"), n_sum)`) reported being less concerned (Supplementary Figure C2).
 
@@
 
-Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy[^3] (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible[^4] (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 2, top panel).
-At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 2, bottom panel).
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 2, top panel).
+At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
 
@@
 
-(ref:typicallyEcawPlotCaption) __Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+(ref:typicallyEcawPlotCaption) __Figure 1.
+Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
 That they merit trust”.
@@
 The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
-For the bottom panel we excluded the missing responses (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "Missing"), n)`) and responses of “I don't understand the question” (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "I don't understand the question"), n)`).
+The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_.
+The bar charts in the bottom panel excluded missing responses (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "Missing"), n)`) and responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "I don't understand the question"), n)`).
 
@@
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2A).
-About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script, method_script == "Never or almost never") %>% pull(n)`/`r filter(script, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script, method_script == "Sometimes") %>% pull(n)`/`r filter(script, method_script == "Sometimes") %>% pull(n_sum)`) (Figure 3).
-`r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst (Figure 3).
-Almost all respondents answered that they use both exploratory (`r pull(exploratory, percentage)`%; `r pull(exploratory, n)`/`r pull(exploratory, n_sum)`) and confirmatory (`r pull(confirmatory, percentage)`%; `r pull(confirmatory, n)`/`r pull(confirmatory, n_sum)`) analyses at least sometimes (Figure 3).
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2).
+About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script, method_script == "Never or almost never") %>% pull(n)`/`r filter(script, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script, method_script == "Sometimes") %>% pull(n)`/`r filter(script, method_script == "Sometimes") %>% pull(n_sum)`).
+`r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst.
+Almost all respondents answered that they use both exploratory (`r pull(exploratory, percentage)`%; `r pull(exploratory, n)`/`r pull(exploratory, n_sum)`) and confirmatory (`r pull(confirmatory, percentage)`%; `r pull(confirmatory, n)`/`r pull(confirmatory, n_sum)`) analyses at least sometimes.
 
@@
 
-(ref:methodPlotCaption) __Responses to survey questions about the research practices of participants.__
+(ref:methodPlotCaption) __Figure 2. Responses to survey questions about the research practices of participants.__
 
@@
 
-(ref:alspacPlotCaption) __Responses to survey questions about using ECAWs.__ For the 4 questions we excluded missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`), responses of "I don't understand the question" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of "Unsure" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
+(ref:alspacPlotCaption) __Figure 3.
+Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`; respectively from top to bottom), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
+Agreement with the first question may be slightly inflated due to the format of the questions in this block.
+Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions.
+Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question \protect\footnotemark[3].
+Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[4].
+
+\protect\footnotetext[3]{Another four respondents agreed with at least the first and second question, which appear contradictory.
+We did not preregister these considerations.
+More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
+\protect\footnotetext[4]{We intended for these questions to ask about ALSPAC running a trial on ECAWs.
+However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}
 
@@
 
-Table 1.
-Recurring topics in responses to the open-ended survey questions.
-The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
+__Table 1.
+Recurring topics in responses to the open-ended survey questions.__ The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
 These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents.
 A complete list of responses are viewable in the open data [LINK].
-We synthesized the response to open-ended questions into the 9 topics on the left side of this table.
-We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as typical preregistration and ECAWs.
-On the right side of the table, we provide a reflection on each topic.
+We synthesized the responses to open-ended questions into the 9 topics on the left side of this table.
+We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as conventional preregistration and ECAWs.
+On the right side of the table, we provide our reflection on each topic.
+
+```{r openEndedTable, echo=FALSE, message=FALSE, warnings=FALSE, fig.align = "center", out.width="100%", fig.id=FALSE}
+open_ended_table <- tibble(
+  subheader = c(
+    "Concerns about acceptability",
+    "Concerns about acceptability",
+    "Concerns about acceptability",
+    "Concerns about acceptability",
+    "Concerns about impact",
+    "Concerns about impact",
+    "Alternative interventions",
+    "Alternative interventions",
+    "Alternative interventions"
+    ),
+  `Respondents’ comments` = c(
+      "1. ECAWs would take too much time.",
+      "2. The subset of data would be insufficient for certain applications (e.g., for genome-wide association studies, rare outcomes, questions regarding small sub-samples, and imputation).",
+      "3. Researchers will often have to change their analysis after seeing the full dataset.",
+      "4. Researchers sometimes reuse the same dataset which they will already have full access to.",
+      "5. Researchers can still p-hack / data-dredge when using ECAWs.",
+      "6. Researchers may give findings from the subset too much weight (e.g., by not performing an analysis on the full dataset because it was null in the subset, or vice versa).",
+      "7. Data management organizations should require final analysis scripts to be shared (so that results are reproducible, and easy for other researchers to build on).",
+      "8. Data management organizations should maintain a repository that outlines each research project conducted using their dataset (to reduce duplication and facilitate replication).",
+      "9. Data management organizations could make a synthetic version of their dataset openly available. That dataset could be used for ECAWs."
+      ),
+    `Our reflection` = c(
+      "Compared to a study that is not preregistered, a study using an ECAW could very well take more time—especially at the beginning of the research process when preparing an analysis script.
+However, the advantages may outweigh the time commitment and further research could help understand this potential tradeoff.
+One survey suggests that psychology researchers believe that preregistration leads to longer project duration, but also higher research quality (Sarafoglou et al., 2022).
+Moreover, compared to conventional preregistration, some evidence suggests that an ECAW-like workflow does not take more time (Sarafoglou et al., 2023).",
+      "Our survey suggested providing access to 10% of the dataset in the first ECAW stage.
+To develop an analysis for certain research questions, this subset of data may be insufficient.
+In these cases, a larger subset could be provided (e.g., 50%) and the final analysis could be run on the other 50% of the data, rather than the full dataset.
+Alternatively, the data management organization could create and provide access to a full-sized synthetic dataset.",
+      "ECAWs do not stop researchers from changing their analysis.
+ECAWs make these changes transparent.
+This allows others to better assess risk of bias and calibrate their confidence in the study findings.
+As compared to conventional preregistration, some evidence suggests that an ECAW-like workflow leads to fewer deviations from the analysis plan (Sarafoglou et al., 2023).",
+      "ECAWs do not stop researchers from reusing the same dataset for new analyses.
+Although, as ECAWs were described in our survey, a research team could only use an ECAW once per dataset (because they will be exposed to the full dataset).\n\nData management organizations could overcome this shortcoming by providing a secure data environment that publicly logs all analyses run, but never exposes the real dataset.
+Alternatively, they could provide access to only the subset of variables needed for a specific analysis.
+This would allow researchers to run another ECAW on the same larger dataset—if it focuses on variables not used in their previous ECAW.",
+      "As is the case for conventional preregistration, ECAWs discourage, but do not necessarily prevent p-hacking and data-dredging.
+ECAWs increase transparency by making these questionable research practices detectable.",
+      "We hope that the instructions provided when implementing ECAWs make it clear that the subset of data is provided to help write an analysis script, but not to help decide which research questions to ask.",
+      "This policy could increase computational reproducibility.
+However, it may be difficult for data management organizations to implement and to ensure compliance.
+Whereas data management organizations can withhold access to the full dataset until an analysis script is registered, their influence is less direct at the publication stage.",
+      "Such a repository could be maintained in parallel with the use of ECAWs.
+It could also be used as a lighter touch intervention that may achieve some of the benefits we presume ECAWs would entail.",
+      "Several initiatives have employed synthetic datasets (e.g., OpenSAFELY, Sarafoglou et al., 2023, Schmidt et al., 2023).
+On the one hand, creating a synthetic dataset requires more technical knowledge as compared to creating a data subset, and synthetic datasets can obscure the relationships between variables, which some researchers may dislike.
+On the other hand, they reduce the risk of participant re-identification and provide a full-sized dataset.
+All considered, synthetic datasets present a reasonable alternative to data subsets."
+      )
+)
+
+open_ended_table %>% 
+  group_by(subheader) %>% 
+  gt::gt(.) %>% 
+  tab_style(
+    style = cell_text(weight = "bold"),
+    locations = list(cells_row_groups(), cells_column_labels())
+    ) %>% 
+  tab_footnote(
+    footnote = "The text in the Respondents’ comments column is synthesized from multiple responses and/or paraphrased. They are not verbatim responses.",
+    locations = cells_column_labels(
+      columns = `Respondents’ comments`
+    )
+  ) %>% 
+  tab_footnote(
+    footnote = "Sarafoglou et al.’s study used a pre-cleaned dataset.
+We are not aware of data that elucidates the time difference between ECAW-like workflows and conventional preregistration when starting from raw data.",
+    locations = cells_body(
+      columns = `Our reflection`,
+      rows = 1
+      )
+  ) %>% 
+  tab_options(table.width = pct(90)) %>% 
+  gtsave("figs/open-ended-table_figure.png", expand = c(0, 100, 0, 100), zoom = 2, vwidth = 1300)
+
+knitr::include_graphics("figs/open-ended-table_figure.png", dpi = 300)
+```
 
 ## Exploratory analyses
+Results can be explored interactively at [TEMPORARY LINK], or by running the code available at [INSERT DATA LINK].
+Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
 
 # Discussion
+The survey results suggest that a trial of ECAWs with ALSPAC could be feasible for at least three reasons.
+First, ECAWs are possible because most respondents reported performing confirmatory analyses with analysis scripts written using programming languages.
+Second, ECAWs are relevant because many respondents reported limited use of other methods to improve rigour and reproducibility—including preregistration, sharing analysis scripts, and blinding analysts.
+Moreover, although participants generally agreed that findings from observational research using pre-existing data are trustworthy and reproducible, they also believed that ECAWs would make research findings more trustworthy and more reproducible.
+Third, ECAWs appear relatively acceptable.
+For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs, and 25% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
 
-# Acknowledgements
+The open-ended responses revealed interest in policies and interventions with similar goals to ECAWs.
+These include requirements for the sharing of final analysis scripts, a database of all ongoing studies that use a particular dataset, and openly available synthetic datasets.
+
+## Limitations
+Given the response rate of 10%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices.
+Indeed, respondents themselves believed that they were more concerned about reproducibility than other researchers in their field (Supplementary Figure C2).
+ECAWs may be less acceptable among non-responders and non-response may be a mark of lack of interest in the concept.
+Nonetheless, the absolute number of responders suggests that there is an audience of researchers who might be interested to pursue this approach.
+
+The survey results are best understood as the initial thoughts of participants when introduced to the concept of ECAWs.
+The median response time was 7.2 minutes (IQR: 4.3 to 13.1) for a survey that included over 20 questions and a 500-word description of ECAWs.
+Thus, it is unlikely that respondents spent much time reflecting on ECAWs and their implications.
+
+We invited researchers from the ALSPAC mailing list which includes researchers across the fields of health and social sciences, but we did not record the specific disciplines in which the respondents were active.
+Thus, while ECAWs may be more relevant to some disciplines, the present survey does delve into the idiosyncrasies among disciplines.
+
+Data management organizations could vary widely in their implementation of ECAWs.
+For example, they could perform checks on the analysis scripts to ensure they run and they could require commented analysis scripts with a clear indication of the primary outcomes.
+Our survey results do not elucidate the best implementation of ECAWs.
+
+## Recommendations
+Some data management organizations looking to implement ECAWs will need to consider a balance between updating their data access workflow and maintaining their user numbers, engagement, and funding.
+This consideration will depend heavily on the structure of the data management organization.
+An organization managing data that is routinely collected somewhat regardless of its potential for use in research (e.g., electronic health records) may have leeway to test new workflows even if they impact user numbers.
+Other organizations—such as ALSPAC—coordinate ongoing data collection efforts that compound the value of their dataset and their continued operation remains contingent on funding cycles and data access fees.
+Even if ECAWs led to an increase in user numbers in the long-term, a temporary decrease could preclude the cost recovery systems on which their staff rely.
+In a more extreme case, the likelihood of another successful funding cycle could be impacted and compromise the project’s continuation.
+Funders and institutions interested in supporting these types of initiatives could alleviate concerns by providing targeted funding for testing these interventions and offering contingency funds to maintain cost recovery systems.
+
+With these considerations in mind, an organization like ALSPAC may benefit from first leveraging the substantial number of respondents who would opt-in to a study on ECAWs.
+Trialing ECAWs with this user group would allow organizations to collect data that may support more widespread implementation, including project completion time, study quality, and researcher satisfaction when using ECAWs.
+They could refine the ECAW pipeline with minimal concern about user numbers.
+In the situation that a data management organization is already considering implementing policies on preregistration, they may benefit from considering alternate workflows such as ECAWs, which many respondents deemed preferable to conventional preregistration.
+The open-ended responses to our survey also suggest some confusion around the purpose of ECAWs and the process of using them.
+A clear-cut module provided by data management organizations that explains the ECAW concept alongside step-by-step instructions could help address researchers’ concerns preemptively and help them adopt this workflow.
+
+The stakes are lower in cases where a static final dataset already exists and concerns about funding are absent.
+For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows (e.g., @Forscher2020; @Schmidt2023).
+Concerns about a reduction in user numbers and engagement may also be less relevant for datasets containing unique data.
+For example, a researcher trying to answer a question about health and development with ALSPAC data, may also be able to answer that question with another cohort dataset.
+However, a researcher trying to answer a question about the population of a specific country may need access to that country’s census data, regardless of the workflow required by that data management organization.
+A final consideration is that user numbers and engagement may increase if researchers feel that ECAWs increase the trustworthiness of their findings and others come to associate research from datasets using ECAWs as more open and rigorous.
+
+## Conclusion
+In this manuscript, we outlined a research workflow—ECAW—which necessitates certain open science practices and can be implemented by data management organizations.
+Responses to our survey provide information for organizations interested in developing and testing ECAWs and interventions with related goals.
+
+---
+
+## Ethical approval
+Ethical approval for the study was obtained from the ALSPAC Ethics and Law Committee and the Faculty of Life Sciences Research Ethics Committee at the University of Bristol (approval code: 12260).
+
+## Acknowledgements
+We thank Nicholas Timpson and Kate Northstone for invaluable input throughout this research project.
+We thank the ALSPAC Executive for their collaboration and willingness to run this study.
+We thank all the participants for their responses.
+
+## Funding
+Robert Thibault is supported by a general support grant awarded to METRICS from Arnold Ventures and a postdoctoral fellowship from the Canadian Institutes of Health Research.
+Alexandra Sarafoglou was supported by an Amsterdam Brain and Cognition project grant (grant ref: ABC PG 22 - January 2022) ``From rigid theory to cognitive models: a framework to study individual differences in meaning representations’’.
+The UK Medical Research Council and Wellcome (Grant ref: 217065/Z/19/Z) and the University of Bristol provide core support for ALSPAC.
+This publication is the work of the authors and Robert Thibault will serve as guarantor for the contents of this paper.
+The funders have no role in the preparation of this manuscript or the decision to publish.
+
+## Contributions
+__Conceptualization:__ Robert T. Thibault and Marcus R. Munafò.
+__Data curation:__ Marton Kovacs.
+__Formal analysis:__ Robert T. Thibault and Marton Kovacs.
+__Funding acquisition:__ Robert T. Thibault and John P. A. Ioannidis.
+__Investigation:__ Robert T. Thibault.
+__Methodology:__ Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
+__Project administration:__ Robert T. Thibault.
+__Resources:__ Robert T. Thibault.
+__Software:__ Marton Kovacs.
+__Supervision:__ Robert T. Thibault and Marcus R. Munafò.
+__Validation:__ Robert T. Thibault and Marton Kovacs.
+__Visualization:__ Robert T. Thibault and Marton Kovacs.
+__Writing - original draft:__ Robert T. Thibault.
+__Writing - review & editing:__ Robert T. Thibault, Marton Kovacs, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
+
+## Competing interests
+All authors declare no conflict of interest.
 

### 2023-04-25 manuscript/manuscript.tex
updated rmarkdown manuscript according to the review

--- 2023-03-23 manuscript/manuscript.tex
@@
 {}{\PatchFailed}
-\keywords{keywords\newline\indent Word count: X}
+\keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 3546}
 \usepackage{lineno}
@@
 \usepackage{csquotes}
+\usepackage{tcolorbox}
+\usepackage{caption}
+\captionsetup[figure]{labelformat=empty}
 \ifLuaTeX
@@
 \hypersetup{
-  pdftitle={A survey of registration practices among observational researchers using preexisting datasets},
-  pdfauthor={Robert T. Thibault1,5, Marton Kovacs2,6, Tom E. Hardwicke3, Alexandra Sarafoglou4, John P. A. Ioannidis4, \& Marcus R. Munafò1,7},
+  pdftitle={Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers},
+  pdfauthor={Robert T. Thibault1,2, Marton Kovacs3,4, Tom E. Hardwicke5, Alexandra Sarafoglou6, John P. A. Ioannidis1,7, \& Marcus R. Munafò2,8},
   pdflang={en-EN},
-  pdfkeywords={keywords},
+  pdfkeywords={Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)},
   hidelinks,
@@
 
-\title{A survey of registration practices among observational researchers using preexisting datasets}
+\title{Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers}
 \author{Robert T.
-Thibault\textsuperscript{1,5}, Marton Kovacs\textsuperscript{2,6}, Tom E.
-Hardwicke\textsuperscript{3}, Alexandra Sarafoglou\textsuperscript{4}, John P.
+Thibault\textsuperscript{1,2}, Marton Kovacs\textsuperscript{3,4}, Tom E.
+Hardwicke\textsuperscript{5}, Alexandra Sarafoglou\textsuperscript{6}, John P.
 A.
-Ioannidis\textsuperscript{4}, \& Marcus R.
-Munafò\textsuperscript{1,7}}
+Ioannidis\textsuperscript{1,7}, \& Marcus R.
+Munafò\textsuperscript{2,8}}
 \date{}
@@
 \authornote{
-
-The authors made the following contributions.
-Robert T.
-Thibault: Conceptualization, Data curation, Formal analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Supervision, Validation, Visualization, Writing - original draft, Writing - review \& editing; Marton Kovacs: Data curation, Formal analysis, Software, Validation, Visualization, Writing - review \& editing; Tom E.
-Hardwicke: Methodology, Writing - review \& editing; Alexandra Sarafoglou: Methodology, Writing - review \& editing; John P.
-A.
-Ioannidis: Methodology, Writing - review \& editing; Marcus R.
-Munafò: Conceptualization, Methodology, Supervision, Writing - review \& editing.
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Meta-Research Innovation Center at Stanford (METRICS), Stanford University.\\\textsuperscript{2} Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{3} Melbourne School of Psychological Sciences, University of Melbourne.\\\textsuperscript{4} Department of Psychology, University of Amsterdam.\\\textsuperscript{5} School of Psychological Science, University of Bristol.\\\textsuperscript{6} Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{7} Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Berlin Institute of Health, Charité -- Universitätsmedizin Berlin.\\\textsuperscript{8} MRC Integrative Epidemiology Unit at the University of Bristol.\\\textsuperscript{9} Departments of Medicine, Epidemiology and Population Health, Biomedical Data Science, and Statistics, Stanford University.}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Meta-Research Innovation Center at Stanford (METRICS), Stanford University.\\\textsuperscript{2} School of Psychological Science, University of Bristol.\\\textsuperscript{3} Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{4} Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary\\\textsuperscript{5} Melbourne School of Psychological Sciences, University of Melbourne.\\\textsuperscript{6} Department of Psychology, University of Amsterdam.\\\textsuperscript{7} Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Berlin Institute of Health, Charité -- Universitätsmedizin Berlin.\\\textsuperscript{8} MRC Integrative Epidemiology Unit at the University of Bristol.}
 
 \abstract{%
-placeholder for an abstract
+\textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis.
+An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW).
+In this workflow, a data management organization initially provides access to only a subset of their dataset to researchers who request it.
+The researchers then prepare an analysis script based on the subset of data, upload the analysis script to a registry, and then receive access to the full dataset.
+ECAWs aim to achieve similar goals as preregistration, but make access to the full dataset contingent on compliance.
+The present survey aimed to garner information from the research community where ECAWs could be applied---employing the Avon Longitudinal Study of Parents and Children (ALSPAC) as a case example.
+\hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC's transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}.
+We received 107 responses, for a 10\% response rate.
+The results suggest that---at least among our sample of respondents---ECAWs hold the potential to serve their intended purpose and appear relatively acceptable.
+For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55\% who agreed).
+However, as many as 25\% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44\% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions.
+\hfill \break \hfill \break \textbf{Preregistration:} \url{https://osf.io/g2fw5} Deviations from the preregistration are outlined in Supplementary Material A.
 }
@@
 
+Many published research findings are non-reproducible and potentially false or misleading (Camerer et al., 2016; Collaboration, 2015; Errington et al., 2021; Ioannidis, 2005, 2008).
+These shortcomings may stem from a combination of researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices.
+These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see Prasad and Cifu (2019)).
+
+Some research disciplines have adopted standards to address these issues.
+For example, researchers conducting clinical trials regularly register outcome measures before enrolling participants, and they keep both participants and experimenters blind regarding group assignment.
+In cases where datasets already exist, however, researchers cannot easily adopt these practices.
+Researchers performing secondary data analyses generally have access to complete datasets without any blinding imposed on them.
+This may lead to exploration of diverse analyses and selective reporting based on the nature of the results.
+
+\hypertarget{preregistration-in-the-context-of-secondary-data-analyses}{%
+\subsection{Preregistration in the context of secondary data analyses}\label{preregistration-in-the-context-of-secondary-data-analyses}}
+
+Increasing the uptake and effectiveness of preregistration for secondary data analyses may require a different approach than it does for clinical trials.
+For example, the analytical space for secondary data analyses (e.g., of a longitudinal cohort dataset) is more vast than the analytical space for most clinical trials.
+Thus, the clinical trial standard of registering outcome measures, but no analysis plan, remains insufficient.
+Moreover, preregistration of secondary data analyses remains uncommon, and unlike for clinical trials (where the dataset doesn't yet exist), there is no guarantee that researchers performing secondary data analyses haven't accessed the dataset.
+To extend the benefits of preregistration to secondary data analyses, we propose a workflow that necessitates the preregistration of an executable analysis script before researchers can access the full dataset.
+
+\hypertarget{explore-and-confirm-analysis-workflow-ecaw}{%
+\subsection{Explore and Confirm Analysis Workflow (ECAW)}\label{explore-and-confirm-analysis-workflow-ecaw}}
+
+In this intervention we propose, a data management organization that controls access to a preexisting data set would first provide access to only a subset of the data a researcher requests.
+After the researcher prepares an analysis script based on the subset of data and uploads it to a registry---where it will be openly accessible and permanent (e.g., to the Open Science Framework Registry)---the data management organization then provides access to the full dataset (e.g., via a secure data environment) and the researcher can proceed as they wish.
+The exact implementation (e.g., whether the data management organization runs quality checks on the analysis scripts) would depend on the preferences of the data management organization and the research community using their dataset (see Box 1 for an hypothetical example).
+
+We call this process an Explore and Confirm Analysis Workflow (ECAW).
+Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly.
+ECAWs may increase the uptake of preregistration and provide assurance that the analyses were developed before the researchers observed the full dataset.
+Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., Bakker et al.
+(2020)), ECAWs provide an executable analysis script.
+Researchers may also find ECAWs more agreeable than conventional preregistration because they allow for exploration.
+ECAWs would not fully solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration.
+
+In short, ECAWs give researchers access to enough data to develop an informed analysis, but without compromising the confirmatory nature of a final analysis on the complete dataset.
+
+\pagebreak
+
+\begin{tcolorbox}
+\textbf{Box 1. A hypothetical example of an ECAW in practice.}
+
+\includegraphics[width=1\textwidth]{./ecaw_workflow.png}
+{\fontsize{10pt}{10pt}\selectfont
+The overarching framework for ECAWs is depicted above.
+Within this framework, data management organizations would need to pre-specify several details for each step.
+Below, we provide a hypothetical example for illustrative purposes.
+The exact implementation of ECAWs would depend on the preferences of each data management organization and the community who uses their dataset.
+
+\begin{enumerate}
+  \item A research team submits to a data management organization (i) a paragraph describing the analyses they want to run, and (ii) a list of the variables they want to analyze.
+  \item The data management organization provides the research team with access to a subset of the data.
+For example, if the research team requests data for 15 variables collected from 10,000 participants, the data management organization will provide access to all 15 variables from a random subset of participants—say 1000.
+  \item The research team prepares an analysis script written in a common programming language (e.g., R or Stata) using the subset of data.
+They register this analysis script, the output from the script, and the paragraph they sent to the data management organization in Step 1, to www.osf.io/registries as an ‘Open-Ended Registration’.
+The researchers are free to run as many analyses as they would like on the subset of data.
+However, they should only register the analysis script they plan to use on the complete dataset.
+  \item The data management organization performs a basic quality check.
+They ensure that (i) the analysis script, the script output, and the paragraph are registered, and (ii) the output contains a clear number of itemized results (e.g., as clinicaltrials.gov does for outcome measures).
+If the quality check fails, the data management organization asks the researchers to update their registration until it passes.
+The research team is then given access to a dataset with all the variables requested for all participants.
+  \item The researchers can proceed as they wish.
+They can run the registered analysis script on the complete dataset, make adjustments to the analysis if desired, or not proceed at all.
+The data management organization does not perform any additional check on what analyses were run.
+Regardless of what the research team decides, a permanent version of the planned analysis is available on the OSF Registry.
+\end{enumerate}
+}
+\end{tcolorbox}
+
+\pagebreak
+
+\hypertarget{examples-of-ecaw-like-workflows}{%
+\subsection{Examples of ECAW-like workflows}\label{examples-of-ecaw-like-workflows}}
+
+One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (Sarafoglou, Hoogeveen, and Wagenmakers (2023)).
+The researchers recruited 120 teams to analyze a single observational dataset and had half the teams preregister a written analysis plan and the other half prepare an analysis plan by drafting an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded).
+Based on self-reports from the participants, the researchers found that two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.
+
+Some disciplines, such as particle physics, implement blind data analyses regularly (Klein and Roodman (2005); MacCoun and Perlmutter (2015)).
+A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al.
+(2020)) and is currently being used for Registered Reports based on a large psychology dataset (Schmidt et al.
+(2023)).
+
+In medicine and clinical epidemiology, a similar workflow is implemented by the software platform OpenSAFELY (www.opensafely.org).
+This platform provides a dataset of simulated health records from which researchers can develop an analysis script.
+When ready, a researcher submits their analysis script which is automatically logged and made public in GitHub.
+The analysis then runs in a Trusted Research Environment (TRE) on data which is stored in the data centers where patients' records already reside (i.e., the data is not copied or moved).
+This workflow keeps individual health records hidden while also documenting all analyses run on the real data.
+
+\hypertarget{study-objectives}{%
+\subsection{Study objectives}\label{study-objectives}}
+
+Here, we present a descriptive and exploratory survey study.
+We had no hypotheses, but we did have two specific objectives.
+(1) To gain insights on the opinions and practices of researchers who already use preexisting observational datasets, in regards to the trustworthiness and reproducibility of research.
+(2) To use these insights to inform future research---including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)---on how data management organizations can encourage rigorous and reproducible research practices.
+
 \hypertarget{methods}{%
 \section{Methods}\label{methods}}
+
+\hypertarget{participants}{%
+\subsection{Participants}\label{participants}}
+
+We sent an email to invite researchers on the mailing list for the UK-based Avon Longitudinal Study of Parents and Children (ALSPAC) to participate in an online survey (see Supplementary Material B).
+We partnered with ALSPAC because they manage an oft-requested dataset and expressed interest in studying ways to ensure the research stemming from their dataset is rigorous.
+ALSPAC is ``a transgenerational prospective observational study investigating influences on health and development across the life course.
+It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.'' (Boyd et al., 2013; Fraser et al., 2013).
+Thus, our invitation reached researchers that use observational data across the health and social sciences.
+The survey was open from 10 Oct 2022 to 1 Nov 2022.
+We sent two reminder emails, exactly one week and two weeks after the original email invitation.
+The mailing list comprises researchers whose email addresses were present on a proposal to access the ALSPAC dataset and included 1148 email addresses.
+
+\hypertarget{survey}{%
+\subsection{Survey}\label{survey}}
+
+The survey contained 6 blocks and is available at \url{https://osf.io/5h7gb}.
+We developed the survey with feedback from the Principal Investigator and Executive Director of ALSPAC (Nicholas Timpson and Kate Northstone).
+We aimed to include as few questions as possible (to encourage a high response rate) while still garnering substantive information on whether ECAWs are relevant and acceptable to ALSPAC users.
+
+Block 1 assessed the extent to which respondents believe that observational research using preexisting data is trustworthy and reproducible (2 questions).
+Block 2 asked respondents how often they use practices related to transparency and reducing researcher bias, including preregistration, blinded data analysis, and sharing analysis scripts (5 questions).
+Between Block 2 and Block 3, the survey described ECAWs.
+Block 3 assessed the extent to which respondents believe that ECAWs would make observational research using preexisting data more trustworthy and reproducible (2 questions).
+Block 4 directly asked respondents whether ALSPAC should run a study on ECAWs and whether they would participate (5 questions).
+Block 5 contained open-ended questions about the perceived benefits and drawbacks of ECAWs, reservations about ECAWs, whether additional incentives might be needed to use ECAW, and suggestions for other research practices or policies that data management organizations could implement to improve research quality (3 questions).
+Block 6 asked participants how concerned they are about research quality, how many relevant studies they have published, what software they use for data analysis, and for any additional comments (4 questions).
+All rating-scale questions contained the response option \emph{``I don't understand the question''}.
+The questions presented in Figure 3 also contained the response option \emph{``Unsure''}.
+
+\hypertarget{analyses}{%
+\subsection{Analyses}\label{analyses}}
+
+We present the results for closed-ended questions as prevalence rates or counts in Figures 1-3 and Supplementary Material C.
+When reporting percentages, we collapse together all positive responses on the rating scales (e.g., \emph{``strongly agree''} and \emph{``somewhat agree''}) as well as all negative responses.
+The total number of responses differ among questions due to the missing values and responses of \emph{``I don't understand the question''} and \emph{``Unsure''}.
+We narratively synthesize the responses to the open-ended questions in Table 1 (i.e., we provide a non-systematic qualitative summary).
 
@@
 
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
+\hypertarget{survey-completion}{%
+\subsection{Survey completion}\label{survey-completion}}
 
-We invited the ALSPAC mailing list to participate, which included 1148 email addresses.
-54 emails bounced, leaving 1094 emails that went through.
-The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for complete surveys and 2\% for incomplete surveys.\footnote{The ALSPAC mailing list has been active for \textgreater30 years and may contain email addresses that are no longer monitored.
-For example, we received one email reply stating that the recipient hasn't been active in research for 30 years.
-Excluding these email addresses would increase the response rate, but we do not know by how much.} The median time taken for complete survey responses was 7.22 minutes (IQR: 4.30 to 13.10).
+Of 1148 emails sent, 1094 went through and 54 bounced.
+The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for complete surveys and 2\% for incomplete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
+The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action.
+Thus, their list likely includes several email addresses that are no longer monitored.
+For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years.
+It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions).
+These two factors may have deflated the response rate.} The median time taken for complete survey responses was 7.20 minutes (IQR: 4.30 to 13.10).
+This manuscript presents the results for complete surveys.
+Supplementary Material D presents the results with partially complete surveys included.
 
-Respondents published a median of NA (IQR 2 to 25) studies using preexisting observational data (Figure S1).
-They reported using the programming languages R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Table S1)\footnote{Participants could select multiple responses to this survey question.}.
-61\% of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); 6\% reported being less concerned.
+\hypertarget{participants-1}{%
+\subsection{Participants}\label{participants-1}}
+
+Respondents had published a median of NA (IQR 2 to 25) studies using preexisting observational data (Supplementary Figure C1).
+They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}.
+61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/101) reported being less concerned (Supplementary Figure C2).
 
@@
 
-Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy\footnote{The survey defined trustworthy as: ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
-That they merit trust.''} (72\%; 77/107) and reproducible\footnote{The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.''} (79\%; 84/107) (Figure 2, top panel).
-At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 71/101) and \emph{more} reproducible (69\%; 70/102) compared to a typical study using preexisting observational data (Figure 2, bottom panel).
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 77/107) and reproducible (79\%; 84/107) (Figure 2, top panel).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 71/101) and \emph{more} reproducible (69\%; 70/102) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
 
@@
 
-\caption{\textbf{Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+\caption{\textbf{Figure 1.
+Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
 That they merit trust''.
@@
 The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
-For the bottom panel we excluded the missing responses (n = 2; 2) and responses of ``I don't understand the question'' (n = 4; 3).}\label{fig:typicallyEcawPlot}
+The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}.
+The bar charts in the bottom panel excluded missing responses (n = 2; 2) and responses of \emph{``I don't understand the question''} (n = 4; 3).}\label{fig:typicallyEcawPlot}
 \end{figure}
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (35\%; 37/107), or sometimes (24\%; 26/107) (Figure 2A).
-About half reported sharing their analysis scripts never or almost never (20\%; 21/107), or sometimes (31\%; 33/107) (Figure 3).
-75\% (80/107) reported that they never or almost never blind the data analyst (Figure 3).
-Almost all respondents answered that they use both exploratory (91\%; 98/107) and confirmatory (85\%; 92/107) analyses at least sometimes (Figure 3).
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (35\%; 37/107), or sometimes (24\%; 26/107) (Figure 2).
+About half reported sharing their analysis scripts never or almost never (20\%; 21/107), or sometimes (31\%; 33/107).
+75\% (80/107) reported that they never or almost never blind the data analyst.
+Almost all respondents answered that they use both exploratory (91\%; 98/107) and confirmatory (85\%; 92/107) analyses at least sometimes.
 
@@
 
-\caption{\textbf{Responses to survey questions about the research practices of participants.}}\label{fig:methodPlot}
+\caption{\textbf{Figure 2. Responses to survey questions about the research practices of participants.}}\label{fig:methodPlot}
 \end{figure}
@@
 
-\caption{\textbf{Responses to survey questions about using ECAWs.} The 4 questions we excluded missing values (\emph{n} = 3; 3; 3; 3), responses of ``I don't understand the question'' (\emph{n} = 0; 4; 1; 1), and responses of ``Unsure'' (\emph{n} = 2; 5; 6; 8).}\label{fig:alspacPlot}
+\caption{\textbf{Figure 3.
+Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 3; 3; 3; 3; respectively from top to bottom), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8).
+Agreement with the first question may be slightly inflated due to the format of the questions in this block.
+Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions.
+Four respondents agreed with all four statements, suggesting they may have glazed over the word ``less'' in the first question \protect\footnotemark[3].
+Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term ``study'' \protect\footnotemark[4].}\label{fig:alspacPlot}
 \end{figure}
 
-Table 1.
-Recurring topics in responses to the open-ended survey questions.
-The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
+\protect\footnotetext[3]{Another four respondents agreed with at least the first and second question, which appear contradictory.
+We did not preregister these considerations.
+More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
+\protect\footnotetext[4]{We intended for these questions to ask about ALSPAC running a trial on ECAWs.
+However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}
+
+\textbf{Table 1.
+Recurring topics in responses to the open-ended survey questions.} The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
 These questions received a total of (92) responses from (55) unique respondents.
 A complete list of responses are viewable in the open data {[}LINK{]}.
-We synthesized the response to open-ended questions into the 9 topics on the left side of this table.
-We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as typical preregistration and ECAWs.
-On the right side of the table, we provide a reflection on each topic.
+We synthesized the responses to open-ended questions into the 9 topics on the left side of this table.
+We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as conventional preregistration and ECAWs.
+On the right side of the table, we provide our reflection on each topic.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figs/open-ended-table_figure}
+
+}
+
+\caption{ }\label{fig:openEndedTable}
+\end{figure}
 
@@
 
+Results can be explored interactively at {[}TEMPORARY LINK{]}, or by running the code available at {[}INSERT DATA LINK{]}.
+Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
+
 \hypertarget{discussion}{%
@@
 
+The survey results suggest that a trial of ECAWs with ALSPAC could be feasible for at least three reasons.
+First, ECAWs are possible because most respondents reported performing confirmatory analyses with analysis scripts written using programming languages.
+Second, ECAWs are relevant because many respondents reported limited use of other methods to improve rigour and reproducibility---including preregistration, sharing analysis scripts, and blinding analysts.
+Moreover, although participants generally agreed that findings from observational research using pre-existing data are trustworthy and reproducible, they also believed that ECAWs would make research findings more trustworthy and more reproducible.
+Third, ECAWs appear relatively acceptable.
+For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, and 25\% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
+
+The open-ended responses revealed interest in policies and interventions with similar goals to ECAWs.
+These include requirements for the sharing of final analysis scripts, a database of all ongoing studies that use a particular dataset, and openly available synthetic datasets.
+
+\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}}
+
+Given the response rate of 10\%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices.
+Indeed, respondents themselves believed that they were more concerned about reproducibility than other researchers in their field (Supplementary Figure C2).
+ECAWs may be less acceptable among non-responders and non-response may be a mark of lack of interest in the concept.
+Nonetheless, the absolute number of responders suggests that there is an audience of researchers who might be interested to pursue this approach.
+
+The survey results are best understood as the initial thoughts of participants when introduced to the concept of ECAWs.
+The median response time was 7.2 minutes (IQR: 4.3 to 13.1) for a survey that included over 20 questions and a 500-word description of ECAWs.
+Thus, it is unlikely that respondents spent much time reflecting on ECAWs and their implications.
+
+We invited researchers from the ALSPAC mailing list which includes researchers across the fields of health and social sciences, but we did not record the specific disciplines in which the respondents were active.
+Thus, while ECAWs may be more relevant to some disciplines, the present survey does delve into the idiosyncrasies among disciplines.
+
+Data management organizations could vary widely in their implementation of ECAWs.
+For example, they could perform checks on the analysis scripts to ensure they run and they could require commented analysis scripts with a clear indication of the primary outcomes.
+Our survey results do not elucidate the best implementation of ECAWs.
+
+\hypertarget{recommendations}{%
+\subsection{Recommendations}\label{recommendations}}
+
+Some data management organizations looking to implement ECAWs will need to consider a balance between updating their data access workflow and maintaining their user numbers, engagement, and funding.
+This consideration will depend heavily on the structure of the data management organization.
+An organization managing data that is routinely collected somewhat regardless of its potential for use in research (e.g., electronic health records) may have leeway to test new workflows even if they impact user numbers.
+Other organizations---such as ALSPAC---coordinate ongoing data collection efforts that compound the value of their dataset and their continued operation remains contingent on funding cycles and data access fees.
+Even if ECAWs led to an increase in user numbers in the long-term, a temporary decrease could preclude the cost recovery systems on which their staff rely.
+In a more extreme case, the likelihood of another successful funding cycle could be impacted and compromise the project's continuation.
+Funders and institutions interested in supporting these types of initiatives could alleviate concerns by providing targeted funding for testing these interventions and offering contingency funds to maintain cost recovery systems.
+
+With these considerations in mind, an organization like ALSPAC may benefit from first leveraging the substantial number of respondents who would opt-in to a study on ECAWs.
+Trialing ECAWs with this user group would allow organizations to collect data that may support more widespread implementation, including project completion time, study quality, and researcher satisfaction when using ECAWs.
+They could refine the ECAW pipeline with minimal concern about user numbers.
+In the situation that a data management organization is already considering implementing policies on preregistration, they may benefit from considering alternate workflows such as ECAWs, which many respondents deemed preferable to conventional preregistration.
+The open-ended responses to our survey also suggest some confusion around the purpose of ECAWs and the process of using them.
+A clear-cut module provided by data management organizations that explains the ECAW concept alongside step-by-step instructions could help address researchers' concerns preemptively and help them adopt this workflow.
+
+The stakes are lower in cases where a static final dataset already exists and concerns about funding are absent.
+For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows (e.g., Forscher et al.
+(2020); Schmidt et al.
+(2023)).
+Concerns about a reduction in user numbers and engagement may also be less relevant for datasets containing unique data.
+For example, a researcher trying to answer a question about health and development with ALSPAC data, may also be able to answer that question with another cohort dataset.
+However, a researcher trying to answer a question about the population of a specific country may need access to that country's census data, regardless of the workflow required by that data management organization.
+A final consideration is that user numbers and engagement may increase if researchers feel that ECAWs increase the trustworthiness of their findings and others come to associate research from datasets using ECAWs as more open and rigorous.
+
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
+
+In this manuscript, we outlined a research workflow---ECAW---which necessitates certain open science practices and can be implemented by data management organizations.
+Responses to our survey provide information for organizations interested in developing and testing ECAWs and interventions with related goals.
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\hypertarget{ethical-approval}{%
+\subsection{Ethical approval}\label{ethical-approval}}
+
+Ethical approval for the study was obtained from the ALSPAC Ethics and Law Committee and the Faculty of Life Sciences Research Ethics Committee at the University of Bristol (approval code: 12260).
+
 \hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\subsection{Acknowledgements}\label{acknowledgements}}
+
+We thank Nicholas Timpson and Kate Northstone for invaluable input throughout this research project.
+We thank the ALSPAC Executive for their collaboration and willingness to run this study.
+We thank all the participants for their responses.
+
+\hypertarget{funding}{%
+\subsection{Funding}\label{funding}}
+
+Robert Thibault is supported by a general support grant awarded to METRICS from Arnold Ventures and a postdoctoral fellowship from the Canadian Institutes of Health Research.
+Alexandra Sarafoglou was supported by an Amsterdam Brain and Cognition project grant (grant ref: ABC PG 22 - January 2022) ``From rigid theory to cognitive models: a framework to study individual differences in meaning representations'\,'.
+The UK Medical Research Council and Wellcome (Grant ref: 217065/Z/19/Z) and the University of Bristol provide core support for ALSPAC.
+This publication is the work of the authors and Robert Thibault will serve as guarantor for the contents of this paper.
+The funders have no role in the preparation of this manuscript or the decision to publish.
+
+\hypertarget{contributions}{%
+\subsection{Contributions}\label{contributions}}
+
+\textbf{Conceptualization:} Robert T. Thibault and Marcus R. Munafò.
+\textbf{Data curation:} Marton Kovacs.
+\textbf{Formal analysis:} Robert T. Thibault and Marton Kovacs.
+\textbf{Funding acquisition:} Robert T. Thibault and John P. A. Ioannidis.
+\textbf{Investigation:} Robert T. Thibault.
+\textbf{Methodology:} Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
+\textbf{Project administration:} Robert T. Thibault.
+\textbf{Resources:} Robert T. Thibault.
+\textbf{Software:} Marton Kovacs.
+\textbf{Supervision:} Robert T. Thibault and Marcus R. Munafò.
+\textbf{Validation:} Robert T. Thibault and Marton Kovacs.
+\textbf{Visualization:} Robert T. Thibault and Marton Kovacs.
+\textbf{Writing - original draft:} Robert T. Thibault.
+\textbf{Writing - review \& editing:} Robert T. Thibault, Marton Kovacs, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
+
+\hypertarget{competing-interests}{%
+\subsection{Competing interests}\label{competing-interests}}
+
+All authors declare no conflict of interest.
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-bakker_ensuring_2020}{}}%
+Bakker, M., Veldkamp, C.
+L., Assen, M.
+A.
+van, Crompvoets, E.
+A., Ong, H.
+H., Nosek, B.
+A., \ldots{} Wicherts, J.
+M.
+(2020).
+Ensuring the quality and specificity of preregistrations.
+\emph{PLoS Biology}, \emph{18}(12), e3000937.
+
+\leavevmode\vadjust pre{\hypertarget{ref-camerer_evaluating_2016}{}}%
+Camerer, C.
+F., Dreber, A., Forsell, E., Ho, T.-H., Huber, J., Johannesson, M., \ldots{} Chan, T.
+(2016).
+Evaluating replicability of laboratory experiments in economics.
+\emph{Science}, \emph{351}(6280), 1433--1436.
+
+\leavevmode\vadjust pre{\hypertarget{ref-collaboration_estimating_2015}{}}%
+Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716.
+
+\leavevmode\vadjust pre{\hypertarget{ref-errington_investigating_2021}{}}%
+Errington, T.
+M., Mathur, M., Soderberg, C.
+K., Denis, A., Perfito, N., Iorns, E., \& Nosek, B.
+A.
+(2021).
+Investigating the replicability of preclinical cancer biology.
+\emph{Elife}, \emph{10}, e71601.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Forscher2020}{}}%
+Forscher, P.
+S., Noyce, A., DeBruine, L.
+M., Jones, B., Flake, J.
+K., Coles, N.
+A., \& Chartier, C.
+R.
+(2020).
+\emph{Incentivizing discovery through the {PSA001} secondary analysis challenge}.
+\url{https://psysciacc.org/2020/09/14/incentivizing-discovery-through-the-psa001-secondary-analysis-challenge/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ioannidis_why_2005}{}}%
+Ioannidis, J. P. (2005). Why most published research findings are false. \emph{PLoS Medicine}, \emph{2}(8), e124.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ioannidis_why_2008}{}}%
+Ioannidis, J. P. (2008). Why most discovered true associations are inflated. \emph{Epidemiology}, 640--648.
+
+\leavevmode\vadjust pre{\hypertarget{ref-klein_blind_2005}{}}%
+Klein, J. R., \& Roodman, A. (2005). Blind analysis in nuclear and particle physics. \emph{Annu. Rev. Nucl. Part. Sci.}, \emph{55}, 141--163.
+
+\leavevmode\vadjust pre{\hypertarget{ref-maccoun_blind_2015}{}}%
+MacCoun, R., \& Perlmutter, S. (2015). Blind analysis: {Hide} results to seek the truth. \emph{Nature}, \emph{526}(7572), 187--189.
+
+\leavevmode\vadjust pre{\hypertarget{ref-prasad_ending_2019}{}}%
+Prasad, V. K., \& Cifu, A. S. (2019). \emph{Ending medical reversal: Improving outcomes, saving lives}. Johns Hopkins University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sarafoglou_comparing_2023}{}}%
+Sarafoglou, A., Hoogeveen, S., \& Wagenmakers, E.-J.
+(2023).
+Comparing analysis blinding with preregistration in the many-analysts religion project.
+\emph{Advances in Methods and Practices in Psychological Science}, \emph{6}(1), 25152459221128319.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2023}{}}%
+Schmidt, K., Jones, A., Szabelska, A., Ebersole, C. R., Hawkins, C. B., Graham, J., \& Nosek, B. A. (2023). \emph{The ideology 2.0 study and dataset}. \url{https://osf.io/2483h/wiki/home/}; OSF.
+
 \end{CSLReferences}
@@
 
-\hypertarget{supplementary-materials-c}{%
-\section{Supplementary materials C}\label{supplementary-materials-c}}
+\hypertarget{supplementary-material-a.-deviations-from-the-preregistration.}{%
+\section{Supplementary Material A. Deviations from the preregistration.}\label{supplementary-material-a.-deviations-from-the-preregistration.}}
+
+\begin{enumerate}
+  \item We rephrased our second objective to be more accurate.
+The preregistration reads “2.
+To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices (survey Blocks 3-6).
+This objective includes assessing and refining potential interventions—such as ECAWs—and assessing their acceptability.
+”.
+The manuscript reads “(2) To use these insights to inform future research—including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)—on how data management organizations can encourage rigorous and reproducible research practices.”
+  \item Marton Kovacs was added as a contributor during data collection.
+This led to a few of the projected contributor roles outlined in the preregistration to be different from the final contributor roles outlined in the manuscript.
+  \item The preregistration stated that: “We will tabulate descriptive summary statistics for all the survey questions.” Instead of tabulating the results, the manuscript presents this data in figures.
+We feel that the figures are easier to digest as compared to tabulated data.
+  \item The preregistration stated that: “we   will   present   results   that   include   responses   from participants who did not complete the entire survey, alongside the associated response rate for each question.”.
+The results from all participants are presented in Supplementary Material D.
+However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed 107 times and partially completed 27 times, leading to a response rate of 10% for complete surveys and 12% for incomplete surveys.”
+  \item The preregistration and survey use the term “typical preregistration”. We changed this to “conventional preregistration” in the manuscript because we believe it is the more appropriate term.
+  \item We performed a brief visual inspection of the data presented in spreadsheet format.
+Three participants clicked through to the end of the survey—so they were coded as completing the survey—however, they only responded to a few of the first questions.
+One other participant appears to have provided non-sincere responses in that all their responses were neutral or skipped and they completed the survey in 146 seconds.
+We retained these four participants as partially completed responses (results in Supplementary Material D), but removed them from the main dataset.
+These decisions were not preregistered.
+  \item We did not preregister which results we would present in the abstract.
+We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 25% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
+\end{enumerate}
+
+\pagebreak
+
+\hypertarget{supplementary-material-b.-invitation-email}{%
+\section{Supplementary Material B. Invitation email}\label{supplementary-material-b.-invitation-email}}
+
+\hypertarget{original-email}{%
+\subsection{Original email}\label{original-email}}
+
+Dear ALSPAC Data User,
+
+We are working with Dr Robert Thibault, a postdoctoral scholar at Stanford University and the University of Bristol, to support his work on scientific rigour and reproducibility in observational studies.
+
+To assist him in his research we would be very grateful if you would consider completing a short survey.
+The purpose of this is to understand researcher's practices and thoughts regarding the rigour and reproducibility of observational research that uses pre-existing datasets (such as the ALSPAC resource).
+Results from this survey may be used in the future to inform initiatives for ALSPAC to better serve our users and to maximise the quality of the research using ALSPAC.
+
+The survey has 21 questions across six sections.
+It will be open until November 1 at this link: \url{https://bristolexppsych.eu.qualtrics.com/jfe/form/SV_3mVgk4lzXx4kp02}.
+It will take approximately 10-20 minutes to complete.
+The data you provide is completely anonymous and further information is available on the consent form at the start of the survey.
+
+If you have any questions or comments, please send them directly to Robert (\href{mailto:robert.thibault@bristol.ac.uk}{\nolinkurl{robert.thibault@bristol.ac.uk}}).
+
+Kind regards,
+
+The ALSPAC Executive
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\hypertarget{first-follow-up-email-sent-after-1-week}{%
+\subsection{First follow-up email (sent after 1 week)}\label{first-follow-up-email-sent-after-1-week}}
+
+Dear ALSPAC data user,
+
+Thankyou very much to those of you who have already completed the survey below.
+There is still time to complete the survey (see details in the original email below), which closes of November 1st.
+
+Please send any comments or queries directly to Robert (\href{mailto:robert.thibault@bristol.ac.uk}{\nolinkurl{robert.thibault@bristol.ac.uk}}).
+
+Many thanks,
+
+The ALSPAC Executive
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+\hypertarget{second-follow-up-email-sent-after-2-weeks}{%
+\subsection{Second follow-up email (sent after 2 weeks)}\label{second-follow-up-email-sent-after-2-weeks}}
+
+Dear ALSPAC data user,
+
+Thank you very much to those of you who have already completed the survey below.
+This will be our final email inviting you to complete the survey (see details in the original email below), which closes of November 1st.
+
+The median time to complete the survey has been 8 minutes.
+
+Please send any comments or queries directly to Robert (\href{mailto:robert.thibault@bristol.ac.uk}{\nolinkurl{robert.thibault@bristol.ac.uk}}).
+Many thanks,
+The ALSPAC Executive
+
+\pagebreak
+
+\hypertarget{supplementary-material-c.-participant-characteristics}{%
+\subsection{Supplementary Material C. Participant Characteristics}\label{supplementary-material-c.-participant-characteristics}}
 
@@
 
-\caption{Caption goes here...}\label{fig:numStudiesPlot}
+\caption{ }\label{fig:numStudiesPlot}
 \end{figure}
@@
 
-\caption{\label{tab:languageTable}What programming language or software do you use for your analyses of preexisting observational data?}
+\caption{\label{tab:languageTable}\textbf{Table C1. ``What programming language or software do you use for your analyses of preexisting observational data? (you may select multiple answers)''}}
 \resizebox{\linewidth}{!}{
@@
 
-\caption{Caption goes here...}\label{fig:concernedPlot}
+\caption{\textbf{Figure C1. Responses to survey question about number of relevant publications}}\label{fig:concernedPlot}
 \end{figure}
 
+\hypertarget{supplementary-material-d.-figures-and-tables-including-participants-who-partially-complete-the-survey}{%
+\subsection{Supplementary Material D.
+Figures and Tables including participants who partially complete the survey}\label{supplementary-material-d.-figures-and-tables-including-participants-who-partially-complete-the-survey}}
+
 \end{document}

### 2023-04-25 manuscript/supplementary_materials.Rmd
Code check (Robby)
removed participants, edited codebook generation, minor tweaks to comments.
updated rmarkdown manuscript according to the review

@@
-# Supplementary materials C
+# Supplementary Material A. Deviations from the preregistration.
 
-```{r numStudiesPlot, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=6, fig.height=3.2, fig.path='figs/', dev=c('png', 'pdf')}
+\begin{enumerate}
+  \item We rephrased our second objective to be more accurate.
+The preregistration reads “2.	To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices (survey Blocks 3-6).
+This objective includes assessing and refining potential interventions—such as ECAWs—and assessing their acceptability.
+”.
+The manuscript reads “(2) To use these insights to inform future research—including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)—on how data management organizations can encourage rigorous and reproducible research practices.”
+  \item Marton Kovacs was added as a contributor during data collection.
+This led to a few of the projected contributor roles outlined in the preregistration to be different from the final contributor roles outlined in the manuscript.
+  \item The preregistration stated that: “We will tabulate descriptive summary statistics for all the survey questions.” Instead of tabulating the results, the manuscript presents this data in figures.
+We feel that the figures are easier to digest as compared to tabulated data.
+  \item The preregistration stated that: “we   will   present   results   that   include   responses   from participants who did not complete the entire survey, alongside the associated response rate for each question.”.
+The results from all participants are presented in Supplementary Material D.
+However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed 107 times and partially completed 27 times, leading to a response rate of 10% for complete surveys and 12% for incomplete surveys.”
+  \item The preregistration and survey use the term “typical preregistration”. We changed this to “conventional preregistration” in the manuscript because we believe it is the more appropriate term.
+  \item We performed a brief visual inspection of the data presented in spreadsheet format.
+Three participants clicked through to the end of the survey—so they were coded as completing the survey—however, they only responded to a few of the first questions.
+One other participant appears to have provided non-sincere responses in that all their responses were neutral or skipped and they completed the survey in 146 seconds.
+We retained these four participants as partially completed responses (results in Supplementary Material D), but removed them from the main dataset.
+These decisions were not preregistered.
+  \item We did not preregister which results we would present in the abstract.
+We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 25% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
+\end{enumerate}
+
+\pagebreak
+
+# Supplementary Material B. Invitation email
+
+## Original email
+
+Dear ALSPAC Data User,
+
+We are working with Dr Robert Thibault, a postdoctoral scholar at Stanford University and the University of Bristol, to support his work on scientific rigour and reproducibility in observational studies.
+
+To assist him in his research we would be very grateful if you would consider completing a short survey.
+The purpose of this is to understand researcher’s practices and thoughts regarding the rigour and reproducibility of observational research that uses pre-existing datasets (such as the ALSPAC resource).
+Results from this survey may be used in the future to inform initiatives for ALSPAC to better serve our users and to maximise the quality of the research using ALSPAC.
+
+The survey has 21 questions across six sections.
+It will be open until November 1 at this link: https://bristolexppsych.eu.qualtrics.com/jfe/form/SV_3mVgk4lzXx4kp02.
+It will take approximately 10-20 minutes to complete.
+The data you provide is completely anonymous and further information is available on the consent form at the start of the survey.
+
+If you have any questions or comments, please send them directly to Robert (robert.thibault@bristol.ac.uk).
+
+Kind regards,
+
+The ALSPAC Executive
+
+---
+
+## First follow-up email (sent after 1 week)
+
+Dear ALSPAC data user,
+
+Thankyou very much to those of you who have already completed the survey below.
+There is still time to complete the survey (see details in the original email below), which closes of November 1st.
+
+Please send any comments or queries directly to Robert (robert.thibault@bristol.ac.uk).
+
+Many thanks,
+
+The ALSPAC Executive
+
+---
+
+## Second follow-up email (sent after 2 weeks)
+
+Dear ALSPAC data user,
+
+Thank you very much to those of you who have already completed the survey below.
+This will be our final email inviting you to complete the survey (see details in the original email below), which closes of November 1st.
+
+The median time to complete the survey has been 8 minutes.
+
+Please send any comments or queries directly to Robert (robert.thibault@bristol.ac.uk). 
+Many thanks,
+The ALSPAC Executive
+
+\pagebreak
+
+## Supplementary Material C. Participant Characteristics
+
+```{r numStudiesPlot, fig.align="center", out.width="100%", fig.width=6, fig.height=3.2, fig.path='figs/', dev=c('png', 'pdf')}
 # Create text for missing responses
@@
 
-(ref:table1c-caption) What programming language or software do you use for your analyses of preexisting observational data?
+(ref:table1c-caption) __Table C1. “What programming language or software do you use for your analyses of preexisting observational data? (you may select multiple answers)”__
 
-```{r concernedPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4, fig.path='figs/', dev=c('png', 'pdf')}
+```{r concernedPlot, warning=FALSE, fig.cap="(ref:concernedPlotCaption)", fig.align="center", out.width="100%", fig.width=10, fig.height=4, fig.path='figs/', dev=c('png', 'pdf')}
 # Concerned figure data preparation
@@
 ```
+
+(ref:concernedPlotCaption) __Figure C1. Responses to survey question about number of relevant publications__
+
+## Supplementary Material D. Figures and Tables including participants who partially complete the survey

### 2023-04-25 manuscript/manuscript.Rmd

@@
 keywords          : "Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)"
-wordcount         : "3546"
+wordcount         : "4020"
 bibliography      : "ecaw-ref.bib"

### 2023-04-25 manuscript/manuscript.tex

@@
 {}{\PatchFailed}
-\keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 3546}
+\keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 4020}
 \usepackage{lineno}
@@
 Of 1148 emails sent, 1094 went through and 54 bounced.
-The survey was completed 107 times and partially completed 21 times, leading to a response rate of 10\% for complete surveys and 2\% for incomplete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
+The survey was completed 103 times and partially completed 17 times, leading to a response rate of 9\% for complete surveys and 2\% for incomplete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
 The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action.
@@
 It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions).
-These two factors may have deflated the response rate.} The median time taken for complete survey responses was 7.20 minutes (IQR: 4.30 to 13.10).
+These two factors may have deflated the response rate.} The median time taken for complete survey responses was 7.40 minutes (IQR: 4.60 to 13.10).
 This manuscript presents the results for complete surveys.
@@
 
-Respondents had published a median of NA (IQR 2 to 25) studies using preexisting observational data (Supplementary Figure C1).
+Respondents had published a median of NA (IQR 2 to 26.20) studies using preexisting observational data (Supplementary Figure C1).
 They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}.
-61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/101) reported being less concerned (Supplementary Figure C2).
+62\% (62/100) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/100) reported being less concerned (Supplementary Figure C2).
 
@@
 
-Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 77/107) and reproducible (79\%; 84/107) (Figure 2, top panel).
-At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 71/101) and \emph{more} reproducible (69\%; 70/102) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 74/103) and reproducible (79\%; 81/103) (Figure 2, top panel).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
 
@@
 The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}.
-The bar charts in the bottom panel excluded missing responses (n = 2; 2) and responses of \emph{``I don't understand the question''} (n = 4; 3).}\label{fig:typicallyEcawPlot}
+The bar charts in the bottom panel excluded missing responses (n = ; ) and responses of \emph{``I don't understand the question''} (n = 3; 2).}\label{fig:typicallyEcawPlot}
 \end{figure}
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (35\%; 37/107), or sometimes (24\%; 26/107) (Figure 2).
-About half reported sharing their analysis scripts never or almost never (20\%; 21/107), or sometimes (31\%; 33/107).
-75\% (80/107) reported that they never or almost never blind the data analyst.
-Almost all respondents answered that they use both exploratory (91\%; 98/107) and confirmatory (85\%; 92/107) analyses at least sometimes.
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2).
+About half reported sharing their analysis scripts never or almost never (20\%; 21/103), or sometimes (32\%; 33/103).
+77\% (79/103) reported that they never or almost never blind the data analyst.
+Almost all respondents answered that they use both exploratory (93\%; 96/103) and confirmatory (87\%; 90/103) analyses at least sometimes.
 
@@
 \caption{\textbf{Figure 3.
-Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 3; 3; 3; 3; respectively from top to bottom), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8).
+Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 0; 0; 0; 0; respectively from top to bottom), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8).
 Agreement with the first question may be slightly inflated due to the format of the questions in this block.

### 2023-04-25 preprocessing/ecaw_raw_processed_preprocessing.Rmd

--- 2023-03-23 preprocessing/ecaw_raw_processed_preprocessing.Rmd
@@
     # dropping those questions that were open ended or conditional on other responses or irrelevant
-    # flag those responses that has any missing values
+    # flag respondents with at least one missing values from a closed-ended question
     mutate(
@@
 
-There were `r filter(raw, has_missing) %>% nrow()` respondents who had at least one missing value in the main questions of interests.
+## Partial responses
+
+We only include data from participants who completed the survey in the final analysis, as preregistered (see the full text of the preregistration [https://osf.io/g2fw5](https://osf.io/g2fw5)).
+There were `r nrow(dplyr::filter(raw, !finished))` respondents who did not complete the survey.
+
+However, these responses will be included in a supplementary analysis that will be reproted in Supplementary Material D. So we save them separately before exclusion.
+
+```{r}
+processed_all <- 
+  raw %>%
+  mutate(duration_in_mins = duration_in_seconds / 60) %>%
+  select(
+    -start_date,
+    -end_date,
+    -status,
+    -progress,
+    -finished,
+    -duration_in_seconds,
+    -recorded_date,
+    -distribution_channel,
+    -user_language
+  ) %>% 
+  # remove participants who only opened the survey but did not answer any questions. All participants who had NA for the first question (typically_trustworthy), had NAs for all questions
+  filter(!is.na(typically_trustworthy))
+
+processed_all <- 
+  processed_all
+
+readr::write_csv(processed_all, here("data/processed/ecaw_processed_all_data.csv"))
+```
+
+The full dataset that we are going to use to calculate the results in Supplement D includes `r nrow(processed_all)` responses.
+
+We create a simple codebook for this dataset.
+
+```{r}
+processed_all_codebook <-
+  read_csv(here("data/source/ecaw_221102_source_codebook.csv")) %>% 
+  select(-original_var_name) %>% 
+  rename(var_name = new_var_name) %>% 
+  filter(var_name %in% colnames(processed_all))
+
+# Adding new column descriptions
+new_row1 <- tibble(
+  description = "True if respondent has any missing responses.",
+  var_name = "has_missing"
+)
+new_row2 <- tibble(
+  description = "Time spent from opening the survey to closing it in minutes.",
+  var_name = "duration_in_mins"
+)
+
+codebook <- bind_rows(processed_all_codebook, new_row1, new_row2)
+
+readr::write_csv(processed_all_codebook, here("data/processed/ecaw_processed_all_codebook.csv"))
+```
+
+Excluding the respondents who did not finish the questionnaire.
+
+```{r}
+raw <- 
+  raw %>% 
+  dplyr::filter(finished) %>%
+  filter(!(response_id %in% c("R_2ZVO2mDq0eBPYDn",
+                              "R_3qr9uw1HGF53UiJ",
+                              "R_3iQHcAwCVxBXAEd",
+                              "R_1HqrkliX8SXIZqS")
+  )
+  )
+# We did a brief visual inspection of the data and feel these first three respondents fit better as partially completed surveys.
+Although they clicked through to the end of the survey, they did no answer questions after a certain point.
+For example, one respondent only answered the first two quetsions and clicked through to the end of the survey in 49 seconds.
+# The fourth response appears to be non-sincere. All their responses were neutral, several were skipped, and it was completed in 146 seconds. This respondent may have simply wanted to see the survey. 
+```
+
+At the end of the exclusion we had `r nrow(raw)` responses remaining for the analysis.
+There were `r filter(raw, has_missing) %>% nrow()` respondents who had at least one missing value for a closed-ended question.
 
@@
 
-On median it took `r median(raw$duration_in_seconds) / 60` to finish the survey to the participants in hour sample.
+The median time to complete the survey was `r median(raw$duration_in_seconds) / 60`.
 
@@
 
-The survey contained 4 open-ended questions (*unsure_explain*, *drawback_ecaw*, *suggestions*, *comments*) that we analyzed qualitatively as preregistered (see page X line Y of the preregistration).
+The survey contained 4 open-ended questions (*unsure_explain*, *drawback_ecaw*, *suggestions*, *comments*) that we analyzed qualitatively, as preregistered.
 
@@
 
+# Adding new column descriptions
+new_row1 <- tibble(
+  description = "True if respondent has any missing responses.",
+  var_name = "has_missing"
+)
+new_row2 <- tibble(
+  description = "Time spent from opening the survey to closing it in minutes.",
+  var_name = "duration_in_mins"
+)
+
+codebook <- bind_rows(codebook, new_row1, new_row2)
+
 # Save the codebook
-readr::write_csv(codebook, here("data/raw/ecaw_raw_codebook.csv"))
+readr::write_csv(codebook, here("data/processed/ecaw_processed_codebook.csv"))
 ```

### 2023-04-26 manuscript/manuscript.Rmd
finished converting the manuscript and cleaning up the codes

@@
 ```{r setup, include=FALSE, warning=FALSE, message=FALSE}
+# Used for rendering the manuscript rmarkdown document
 library("papaja")
+# Used for general data wrangling
 library(tidyverse)
+# Used to make exact file paths
 library(here)
+# Used for tables
 library(kableExtra)
 library(gt)
+# Used for joining figures
 library(patchwork)
 # The developer version is needed
+# Used for creating Likert ggplots
 library(likert)
+# Used for knitting rmarkdown documents
 library(knitr)
+# Used for colorblind ggplot colors
 library(viridis)
+# References
 r_refs("ecaw-ref.bib")
@@
 ```{r loadingData}
+# Loading processed datafiles
 processed <- read_csv(here::here("data/processed/ecaw_processed_data.csv"))
 processed_all <- read_csv(here::here("data/processed/ecaw_processed_all_data.csv"))
-```
-
-```{r analysisPreferences}
-# Seed for random number generation
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
 ```
@@
 Of 1148 emails sent, 1094 went through and 54 bounced.
-The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
+The survey was completed `r nrow(processed)` times and partially completed `r nrow(processed_all) - nrow(processed)` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(processed_all) / 1094 * 100)`% for at least partially complete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
 The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action.
@@
 It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions).
-These two factors may have deflated the response rate.} The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
+These two factors may have deflated the response rate.} The median time to complete the survey was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
 This manuscript presents the results for complete surveys.
@@
 
-Respondents had published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)`) studies using preexisting observational data (Supplementary Figure C1).
+Respondents had published a median of `r median(processed$n_studies, na.rm = TRUE)` (IQR `r round(quantile(processed$n_studies, .25, na.rm = TRUE))` to `r round(quantile(processed$n_studies, .75, na.rm = TRUE))`) studies using preexisting observational data (Supplementary Figure C1).
 They reported using the following programming languages or software packages: `r language_text` (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}.
@@
 The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_.
-The bar charts in the bottom panel excluded missing responses (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "Missing"), n)`) and responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "I don't understand the question"), n)`).
+The bar charts in the bottom panel excluded responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible"), n)`).
 
@@
   ungroup()
-
-# method_missing_data <-
-#   method_data %>%
-#   filter(scale %in% c("Missing", "I don't understand the question")) %>%
-#   mutate(
-#     scale = case_when(
-#       scale == "I don't understand the question" ~ "Don't understand",
-#       scale == "Missing" ~ "Missing"),
-#     text = paste0(scale, ": ", n)
-#     ) %>%
-#   group_by(type) %>%
-#   summarise(text = stringr::str_c(text, collapse = "\n"))
 
@@
 (ref:alspacPlotCaption) __Figure 3.
-Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`; respectively from top to bottom), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
+Responses to survey questions about using ECAWs.__ These bar charts exclude responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`; respectively from top to bottom), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
 Agreement with the first question may be slightly inflated due to the format of the questions in this block.

### 2023-04-26 manuscript/manuscript.tex
finished converting the manuscript and cleaning up the codes

@@
 Of 1148 emails sent, 1094 went through and 54 bounced.
-The survey was completed 103 times and partially completed 17 times, leading to a response rate of 9\% for complete surveys and 2\% for incomplete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
+The survey was completed 103 times and partially completed 20 times, leading to a response rate of 9\% for complete surveys and 11\% for at least partially complete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to.
 The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action.
@@
 It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions).
-These two factors may have deflated the response rate.} The median time taken for complete survey responses was 7.40 minutes (IQR: 4.60 to 13.10).
+These two factors may have deflated the response rate.} The median time to complete the survey was 7.40 minutes (IQR: 4.60 to 13.10).
 This manuscript presents the results for complete surveys.
@@
 
-Respondents had published a median of NA (IQR 2 to 26.20) studies using preexisting observational data (Supplementary Figure C1).
+Respondents had published a median of 10 (IQR 2 to 26) studies using preexisting observational data (Supplementary Figure C1).
 They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}.
@@
 The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}.
-The bar charts in the bottom panel excluded missing responses (n = ; ) and responses of \emph{``I don't understand the question''} (n = 3; 2).}\label{fig:typicallyEcawPlot}
+The bar charts in the bottom panel excluded responses of \emph{``I don't understand the question''} (n = 3; 2).}\label{fig:typicallyEcawPlot}
 \end{figure}
@@
 \caption{\textbf{Figure 3.
-Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 0; 0; 0; 0; respectively from top to bottom), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8).
+Responses to survey questions about using ECAWs.} These bar charts exclude responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1; respectively from top to bottom), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8).
 Agreement with the first question may be slightly inflated due to the format of the questions in this block.
@@
 
-Thankyou very much to those of you who have already completed the survey below.
-There is still time to complete the survey (see details in the original email below), which closes of November 1st.
+Thank you very much to those of you who have already completed the survey below.
+There is still time to complete the survey (see details in the original email below), which closes on November 1st.
 
@@
 Please send any comments or queries directly to Robert (\href{mailto:robert.thibault@bristol.ac.uk}{\nolinkurl{robert.thibault@bristol.ac.uk}}).
+
 Many thanks,
@@
 
+\hypertarget{participants-2}{%
+\subsubsection{Participants}\label{participants-2}}
+
+Respondents had published a median of 10 (IQR 2 to 25) studies using preexisting observational data (Supplementary Figure C1).
+They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Supplementary Table C1)\footnote[1]{Participants could select multiple responses to this survey question.}.
+61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/101) reported being less concerned (Supplementary Figure C2).
+
+\hypertarget{survey-results-1}{%
+\subsection{Survey results}\label{survey-results-1}}
+
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (71\%; 87/123) and reproducible (77\%; 93/121) (Figure D2, top panel).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/107) and \emph{more} reproducible (69\%; 74/108) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figs/typicallyEcawPlotAll-1}
+
+}
+
+\caption{\textbf{Figure 1D.
+Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+That they merit trust''.
+The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange.
+The number in the center of the data bar (gray) indicates the percentage of neutral responses.
+The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
+The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}.
+The bar charts in the bottom panel excluded missing responses (n = 12; 12) and responses of \emph{``I don't understand the question''} (n = 4; 3).}\label{fig:typicallyEcawPlotAll}
+\end{figure}
+
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (33\%; 41/123), or sometimes (21\%; 26/103) (Figure 2D).
+About half reported sharing their analysis scripts never or almost never (19\%; 23/123), or sometimes (29\%; 36/123).
+69\% (85/123) reported that they never or almost never blind the data analyst.
+Almost all respondents answered that they use both exploratory (87\%; 107/123) and confirmatory (82\%; 101/123) analyses at least sometimes.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figs/methodPlotAll-1}
+
+}
+
+\caption{\textbf{Figure 2D. Responses to survey questions about the research practices of participants.}}\label{fig:methodPlotAll}
+\end{figure}
+
+25\% (26/106) of respondents agreed (versus 44\%; 47/106 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3D).
+51\% (50/98) agreed (21\%; 21/98 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
+54\% (54/100) agreed (11\%; 11/100 disagreed) that ALSPAC should run a study on ECAWs.
+45\% (45/100) agreed (22\%; 22/100 disagreed) that they would prefer using an ECAW than using typical preregistration.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{figs/alspacPlotAll-1}
+
+}
+
+\caption{\textbf{Figure 3D.
+Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 14; 14; 14; 14; respectively from top to bottom), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 5; 2; 1), and responses of \emph{``Unsure''} (\emph{n} = 3; 6; 7; 8).
+Agreement with the first question may be slightly inflated due to the format of the questions in this block.
+Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions.
+Four respondents agreed with all four statements, suggesting they may have glazed over the word ``less'' in the first question \protect\footnotemark[2].
+Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term ``study'' \protect\footnotemark[3].}\label{fig:alspacPlotAll}
+\end{figure}
+
+\protect\footnotetext[2]{Another four respondents agreed with at least the first and second question, which appear contradictory.
+We did not preregister these considerations.
+More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
+\protect\footnotetext[3]{We intended for these questions to ask about ALSPAC running a trial on ECAWs.
+However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}
+
 \end{document}

### 2023-04-26 manuscript/supplementary_materials.Rmd
finished converting the manuscript and cleaning up the codes

@@
 Please send any comments or queries directly to Robert (robert.thibault@bristol.ac.uk).
+
 Many thanks,
@@
 ## Supplementary Material D. Figures and Tables including participants who partially complete the survey
+### Participants
+
+```{r languageDescriptivesAll, include=FALSE}
+# List of used programming languages and their counts text
+language_text_all <- 
+  processed_all %>%
+  rename(lang = programming_language) %>%
+  separate_rows(lang, sep = ",") %>%
+  count(lang) %>%
+  filter(!is.na(lang)) %>%
+  arrange(desc(n)) %>%
+  # Removing irrelevant
+  filter(lang != "irrelevant") %>%
+  mutate(lang_n = glue::glue("{lang} (n = {n})")) %>%
+  pull(lang_n) %>%
+  glue::glue_collapse(., sep = ", ", last = ", and ")
+```
+
+```{r concernedDescriptivesAll, include=FALSE}
+# Calculate concerned percentage
+concerned_all <- support_percentage(
+  processed_all,
+  concerned,
+  c(
+    "very much less concerned",
+    "less concerned",
+    "somewhat less concerned",
+    "as concerned as a typical researcher in my field",
+    "somewhat more concerned",
+    "more concerned",
+    # This level label starts with a capital letter
+    "Very much more concerned"
+  ),
+  exclude_missing = TRUE
+)
+```
+
+Respondents had published a median of `r median(processed_all$n_studies, na.rm = TRUE)` (IQR `r round(quantile(processed_all$n_studies, .25, na.rm = TRUE))` to `r round(quantile(processed_all$n_studies, .75, na.rm = TRUE))`) studies using preexisting observational data (Supplementary Figure C1).
+They reported using the following programming languages or software packages: `r language_text_all` (Supplementary Table C1)\footnote[1]{Participants could select multiple responses to this survey question.}.
+`r pull(filter(concerned_all, support == "positive"), percentage)`% (`r pull(filter(concerned_all, support == "positive"), n)`/`r pull(filter(concerned_all, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; `r pull(filter(concerned_all, support == "negative"), percentage)`% (`r pull(filter(concerned_all, support == "negative"), n)`/`r pull(filter(concerned_all, support == "negative"), n_sum)`) reported being less concerned (Supplementary Figure C2).
+
+## Survey results
+
+```{r typicallyEcawDescriptivesAll, include=FALSE}
+# Calculate typical trustworthiness and reproducibility ratings
+typically_trustworthy_all <- support_percentage(
+  processed_all,
+  typically_trustworthy,
+  c(
+    "Strongly disagree",
+    "Somewhat disagree",
+    "Neither agree nor disagree",
+    "Somewhat agree",
+    "Strongly agree"
+  ),
+  "I don't understand the question"
+)
+
+typically_reproducible_all <- support_percentage(
+  processed_all,
+  typically_reproducible,
+  c(
+    "Strongly disagree",
+    "Somewhat disagree",
+    "Neither agree nor disagree",
+    "Somewhat agree",
+    "Strongly agree"
+  ),
+  "I don't understand the question"
+)
+
+# Calculate ecaw trustworthiness and reproducibility ratings
+ecaw_trustworthy_all <- support_percentage(
+  processed_all,
+  ecaw_trustworthy,
+  c(
+    "Much less",
+    "Somewhat less",
+    "About the same",
+    "Somewhat more",
+    "Much more"
+  ),
+  "I don't understand the question"
+)
+
+ecaw_reproducible_all <- support_percentage(
+  processed_all,
+  ecaw_reproducible,
+  c(
+    "Much less",
+    "Somewhat less",
+    "About the same",
+    "Somewhat more",
+    "Much more"
+  ),
+  "I don't understand the question"
+)
+```
+
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy_all, support == "positive"), n)`/`r pull(filter(typically_trustworthy_all, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible_all, support == "positive"), n)`/`r pull(filter(typically_reproducible_all, support == "positive"), n_sum)`) (Figure D2, top panel).
+At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy_all, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible_all, support == "positive"), n)`/`r pull(filter(ecaw_reproducible_all, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+
+```{r typicallyEcawPlotAll, warning=FALSE, fig.cap="(ref:typicallyEcawPlotAllCaption)", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
+# Create typically trustworthy and reproducible plot
+# Check if there is any missing and don't understand responses separately
+# processed %>%
+#   select(contains("typically_")) %>%
+#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
+#   nrow()
+
+# Prepare plot data
+typically_plot_data_all <-
+  processed_all %>%
+  select(contains("typically_")) %>%
+  mutate(
+    across(
+      .fns = ~ factor(., levels = c(
+        "Strongly disagree",
+        "Somewhat disagree",
+        "Neither agree nor disagree",
+        "Somewhat agree",
+        "Strongly agree"
+      ))
+    )
+  ) %>%
+  rename(
+    Reproducible = typically_reproducible,
+    Trustworthy = typically_trustworthy
+  )
+
+# Create likert package data not including the missing values
+typically_plot_data_all <- likert(typically_plot_data_all)
+
+# Create figure
+typically_plot_all <-
+  plot(typically_plot_data_all, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+  labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_text(size = 14),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_blank(),
+    legend.text = element_text(size = 13.5),
+    legend.title = element_blank()
+  )
+
+# Create ecaw trustworthy and reproducible plot
+# Calculate missing responses
+ecaw_missing_all <-
+  processed_all %>%
+  select(contains("ecaw_")) %>%
+  pivot_longer(
+    cols = everything(),
+    names_to = "item",
+    values_to = "value"
+  ) %>%
+  dplyr::filter(is.na(value) | value == "I don't understand the question") %>%
+  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
+  replace_na(list(value = "Missing")) %>%
+  group_by(item, value) %>%
+  summarise(
+    n = n()
+  )
+
+# Prepare plot data
+ecaw_plot_data_all <-
+  processed_all %>%
+  select(contains("ecaw_")) %>%
+  # Transform not wanted response values to NA
+  # likert::likert drops NA values silently
+  # When var transformed to factor these values would be transformed to NA
+  # Automatically but I try to be explicit
+  mutate(
+    across(
+      everything(),
+      ~ case_when(
+        . == "I don't understand the question" ~ NA_character_,
+        TRUE ~ .
+      )
+    )
+  ) %>%
+  mutate(
+    across(
+      .fns = ~ factor(., levels = c(
+        "Much less",
+        "Somewhat less",
+        "About the same",
+        "Somewhat more",
+        "Much more"
+      ))
+    )
+  ) %>%
+  rename(
+    Reproducible = ecaw_reproducible,
+    Trustworthy = ecaw_trustworthy
+  )
+
+# Create likert package data not including the missing values
+ecaw_plot_data_all <- likert(ecaw_plot_data_all)
+
+# Create figure
+ecaw_plot_all <-
+  plot(ecaw_plot_data_all, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+  labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
+  theme(
+    title = element_text(size = 14),
+    axis.text.y = element_text(size = 14),
+    axis.text.x = element_text(size = 15),
+    axis.title.x = element_blank(),
+    legend.text = element_text(size = 13.5),
+    legend.title = element_blank(),
+    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2.3, unit = "cm")
+  )
+
+# Join the two plots together to make one figure
+typically_plot_all + ecaw_plot_all + plot_layout(ncol = 1, nrow = 2)
+```
+
+(ref:typicallyEcawPlotAllCaption) __Figure 1D.
+Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+That they merit trust”.
+The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.” For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange.
+The number in the center of the data bar (gray) indicates the percentage of neutral responses.
+The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
+The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_.
+The bar charts in the bottom panel excluded missing responses (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "Missing"), n)`) and responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "I don't understand the question"), n)`).
+
+```{r methodDescriptivesAll, include=FALSE}
+# Define the response level order for these variables
+method_levels <- c(
+  "Never or almost never",
+  "Sometimes",
+  "About half the time",
+  "Most of the time",
+  "Always or almost always",
+  "I don't understand the question",
+  "Missing"
+)
+
+# Calculate the frequency and proportion of responses for each response level
+confirmatory_all <- level_percentage(
+  processed_all,
+  method_confirmatory,
+  method_levels,
+  exclude_missing = FALSE
+) %>%
+  filter(method_confirmatory %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
+  summarise(
+    n = sum(n),
+    # Since n_sum should be the same for the whole dataset we can use unique to return one value
+    n_sum = unique(n_sum),
+    percentage = sum(percentage))
+
+exploratory_all <- level_percentage(
+  processed_all,
+  method_exploratory,
+  method_levels,
+  exclude_missing = FALSE
+) %>%
+  dplyr::filter(method_exploratory  %in% c("Sometimes", "About half the time", "Most of the time", "Always or almost always")) %>%
+  summarise(
+    n = sum(n),
+    # Since n_sum should be the same for the whole dataset we can use unique to return one value
+    n_sum = unique(n_sum),
+    percentage = sum(percentage))
+
+preregistered_all <- level_percentage(
+  processed_all,
+  method_preregistered,
+  method_levels,
+  exclude_missing = FALSE
+)
+
+blind_all <- level_percentage(
+  processed_all,
+  method_blind,
+  method_levels,
+  exclude_missing = FALSE
+)
+
+script_all <- level_percentage(
+  processed_all,
+  method_script,
+  method_levels,
+  exclude_missing = FALSE
+)
+```
+
+Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2D).
+About half reported sharing their analysis scripts never or almost never (`r filter(script_all, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script_all, method_script == "Never or almost never") %>% pull(n)`/`r filter(script_all, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script_all, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script_all, method_script == "Sometimes") %>% pull(n)`/`r filter(script_all, method_script == "Sometimes") %>% pull(n_sum)`).
+`r filter(blind_all, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind_all, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind_all, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst.
+Almost all respondents answered that they use both exploratory (`r pull(exploratory_all, percentage)`%; `r pull(exploratory_all, n)`/`r pull(exploratory_all, n_sum)`) and confirmatory (`r pull(confirmatory_all, percentage)`%; `r pull(confirmatory_all, n)`/`r pull(confirmatory_all, n_sum)`) analyses at least sometimes.
+
+```{r methodPlotAll, warning=FALSE, fig.cap="(ref:methodPlotAllCaption)", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
+# Used methods figures
+method_plot_data_all <-
+  processed_all %>%
+  select(starts_with("method_")) %>%
+  pivot_longer(
+    cols = everything(),
+    names_to = "variable",
+    values_to = "scale",
+  ) %>%
+  separate(variable, into = c("name_prefix", "type")) %>%
+  group_by(type) %>%
+  count(scale) %>%
+  ungroup() %>%
+  replace_na(list(scale = "Missing")) %>%
+  mutate(
+    scale = factor(scale, levels = c(
+      "Never or almost never",
+      "Sometimes",
+      "About half the time",
+      "Most of the time",
+      "Always or almost always",
+      "Missing",
+      "I don't understand the question"
+    )),
+    type = case_when(
+      type == "blind" ~ "Blind the data analyst",
+      type == "confirmatory" ~ "Contain confirmatory analysis",
+      type == "exploratory" ~ "Contain exploratory analysis",
+      type == "preregistered" ~ "Are preregistered",
+      type == "script" ~ "Share analysis scripts",
+    ),
+    type = factor(type, levels = c(
+      "Contain exploratory analysis",
+      "Contain confirmatory analysis",
+      "Blind the data analyst",
+      "Share analysis scripts",
+      "Are preregistered"
+    ))
+  ) %>%
+  group_by(type) %>%
+  tidyr::complete(scale, fill = list(n = 0)) %>%
+  mutate(
+    percentage = n / sum(n)
+  ) %>%
+  ungroup()
+
+method_plot_data_all %>%
+  # filter(scale %ni% c("Missing", "I don't understand the question")) %>%
+  ggplot() +
+  aes(
+    x = percentage,
+    y = type,
+    fill = scale
+  ) +
+  geom_bar(position = "stack", stat = "identity") +
+  scale_x_continuous(
+    limits = c(0, 1),
+    labels = scales::label_percent()
+  ) +
+  scale_fill_viridis(option = "A", discrete = TRUE) +
+  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) +
+  labs(
+    title = "The studies using preexisting observational data that I am involved in..."
+  ) +
+  papaja::theme_apa() +
+  theme(
+    title = element_text(size = 17),
+    axis.text = element_text(size = 15),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 14),
+    legend.title = element_blank(),
+    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
+  )
+```
+
+(ref:methodPlotAllCaption) __Figure 2D. Responses to survey questions about the research practices of participants.__
+
+```{r aslpacDescriptivesAll, include=FALSE, warning=FALSE, message=FALSE}
+willing_all <- support_percentage(
+  processed_all,
+  alspac_less_willing,
+  c(
+    "Strongly disagree",
+    "Somewhat disagree",
+    "Neither agree nor disagree",
+    "Somewhat agree",
+    "Strongly agree"
+  ),
+  c(
+    "I don't understand the question",
+    "Unsure"
+  )
+)
+
+opt_all <- support_percentage(
+  processed_all,
+  alspac_opt_in,
+  c(
+    "Strongly disagree",
+    "Somewhat disagree",
+    "Neither agree nor disagree",
+    "Somewhat agree",
+    "Strongly agree"
+  ),
+  c(
+    "I don't understand the question",
+    "Unsure"
+  )
+)
+
+study_all <- support_percentage(
+  processed_all,
+  alspac_study,
+  c(
+    "Strongly disagree",
+    "Somewhat disagree",
+    "Neither agree nor disagree",
+    "Somewhat agree",
+    "Strongly agree"
+  ),
+  c(
+    "I don't understand the question",
+    "Unsure"
+  )
+)
+
+prefer_all <- support_percentage(
+  processed_all,
+  alspac_prefer_ecaw,
+  c(
+    "Strongly disagree",
+    "Somewhat disagree",
+    "Neither agree nor disagree",
+    "Somewhat agree",
+    "Strongly agree"
+  ),
+  c(
+    "I don't understand the question",
+    "Unsure"
+  )
+)
+```
+
+`r pull(filter(willing_all, support == "positive"), percentage)`% (`r pull(filter(willing_all, support == "positive"), n)`/`r pull(filter(willing_all, support == "positive"), n_sum)`) of respondents agreed (versus `r pull(filter(willing_all, support == "negative"), percentage)`%; `r pull(filter(willing_all, support == "negative"), n)`/`r pull(filter(willing_all, support == "negative"), n_sum)` who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3D).
+`r pull(filter(opt_all, support == "positive"), percentage)`% (`r pull(filter(opt_all, support == "positive"), n)`/`r pull(filter(opt_all, support == "positive"), n_sum)`) agreed (`r pull(filter(opt_all, support == "negative"), percentage)`%; `r pull(filter(opt_all, support == "negative"), n)`/`r pull(filter(opt_all, support == "negative"), n_sum)` disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
+`r pull(filter(study_all, support == "positive"), percentage)`% (`r pull(filter(study_all, support == "positive"), n)`/`r pull(filter(study_all, support == "positive"), n_sum)`) agreed (`r pull(filter(study_all, support == "negative"), percentage)`%; `r pull(filter(study_all, support == "negative"), n)`/`r pull(filter(study_all, support == "negative"), n_sum)` disagreed) that ALSPAC should run a study on ECAWs.
+`r pull(filter(prefer_all, support == "positive"), percentage)`% (`r pull(filter(prefer_all, support == "positive"), n)`/`r pull(filter(prefer_all, support == "positive"), n_sum)`) agreed (`r pull(filter(prefer, support == "negative"), percentage)`%; `r pull(filter(prefer_all, support == "negative"), n)`/`r pull(filter(prefer_all, support == "negative"), n_sum)` disagreed) that they would prefer using an ECAW than using typical preregistration.
+
+```{r alspacPlotAll, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotAllCaption)", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
+# Check if there is any missing and don't understand responses separately
+alspac_missing_all <-
+  processed_all %>%
+  select(contains("alspac_")) %>%
+  pivot_longer(
+    cols = everything(),
+    names_to = "item",
+    values_to = "value"
+  ) %>%
+  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>%
+  mutate(
+    item = str_to_title(str_remove(item, "^[^_]*_")),
+    # We reorder it by making it a factor
+    item = factor(item, levels = c(
+      "Less_willing",
+      "Opt_in",
+      "Study",
+      "Prefer_ecaw"
+    )),
+    # By making it a factor we can complete missing levels in the data later
+    value = factor(value, levels = c(
+      "Missing",
+      "I don't understand the question",
+      "Unsure"
+    ))
+  ) %>%
+  replace_na(list(value = "Missing")) %>%
+  group_by(item, value) %>%
+  summarise(
+    n = n()
+  ) %>% 
+  # Add n = 0 if no one choose a specific factor level
+  tidyr::complete(value, fill = list(n = 0))
+
+# Prepare plot data
+alspac_plot_data_all <-
+  processed_all %>%
+  select(contains("alspac_")) %>%
+  # Transform not wanted response values to NA
+  # likert::likert drops NA values silently
+  # When var transformed to factor these values would be transformed to NA
+  # Automatically but I try to be explicit
+  mutate(
+    across(
+      everything(),
+      ~ case_when(
+        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
+        TRUE ~ .
+      )
+    )
+  ) %>%
+  mutate(
+    across(
+      .fns = ~ factor(., levels = c(
+        "Strongly disagree",
+        "Somewhat disagree",
+        "Neither agree nor disagree",
+        "Somewhat agree",
+        "Strongly agree"
+      ))
+    )
+  ) %>%
+  rename(
+    `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
+    `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
+    `ALSPAC should run a study on ECAWs.` = alspac_study,
+    `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
+  )
+
+# Create likert package data not including the missing values
+alspac_plot_likert_data_all <- likert(alspac_plot_data_all)
+
+# Create figure
+plot(alspac_plot_likert_data_all, digits = 1, text.size = 6) +
+  # TODO: have to check if this is not messing with the results
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 30), limits = rev(names(alspac_plot_data))) +
+  scale_y_continuous(labels = c("100%", "50%", "0%", "50%", "100%"), limits = c(-105, 105)) +
+  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
+  theme(
+    axis.title.x = element_blank(),
+    title = element_text(size = 15.5),
+    axis.text.y = element_text(size = 14),
+    axis.text.x = element_text(size = 16),
+    legend.text = element_text(size = 15),
+    legend.title = element_blank()
+  )
+```
+
+(ref:alspacPlotAllCaption) __Figure 3D.
+Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Missing"), n)`; respectively from top to bottom), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Unsure"), n)`).
+Agreement with the first question may be slightly inflated due to the format of the questions in this block.
+Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions.
+Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question \protect\footnotemark[2].
+Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[3].
+
+\protect\footnotetext[2]{Another four respondents agreed with at least the first and second question, which appear contradictory.
+We did not preregister these considerations.
+More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
+\protect\footnotetext[3]{We intended for these questions to ask about ALSPAC running a trial on ECAWs.
+However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}

### 2023-04-26 preprocessing/ecaw_raw_processed_preprocessing.Rmd
finished converting the manuscript and cleaning up the codes

@@
 
-# Exclusions
-## Test data
+# Excluding test data
 
@@
 ```
+
+# Flagging not complete responses
 
@@
 
-## Partial responses
-
-We only include data from participants who completed the survey in the final analysis, as preregistered (see the full text of the preregistration [https://osf.io/g2fw5](https://osf.io/g2fw5)).
-There were `r nrow(dplyr::filter(raw, !finished))` respondents who did not complete the survey.
-
-However, these responses will be included in a supplementary analysis that will be reproted in Supplementary Material D. So we save them separately before exclusion.
-
-```{r}
-processed_all <- 
-  raw %>%
-  mutate(duration_in_mins = duration_in_seconds / 60) %>%
-  select(
-    -start_date,
-    -end_date,
-    -status,
-    -progress,
-    -finished,
-    -duration_in_seconds,
-    -recorded_date,
-    -distribution_channel,
-    -user_language
-  ) %>% 
-  # remove participants who only opened the survey but did not answer any questions. All participants who had NA for the first question (typically_trustworthy), had NAs for all questions
-  filter(!is.na(typically_trustworthy))
-
-processed_all <- 
-  processed_all
-
-readr::write_csv(processed_all, here("data/processed/ecaw_processed_all_data.csv"))
-```
-
-The full dataset that we are going to use to calculate the results in Supplement D includes `r nrow(processed_all)` responses.
-
-We create a simple codebook for this dataset.
-
-```{r}
-processed_all_codebook <-
-  read_csv(here("data/source/ecaw_221102_source_codebook.csv")) %>% 
-  select(-original_var_name) %>% 
-  rename(var_name = new_var_name) %>% 
-  filter(var_name %in% colnames(processed_all))
-
-# Adding new column descriptions
-new_row1 <- tibble(
-  description = "True if respondent has any missing responses.",
-  var_name = "has_missing"
-)
-new_row2 <- tibble(
-  description = "Time spent from opening the survey to closing it in minutes.",
-  var_name = "duration_in_mins"
-)
-
-codebook <- bind_rows(processed_all_codebook, new_row1, new_row2)
-
-readr::write_csv(processed_all_codebook, here("data/processed/ecaw_processed_all_codebook.csv"))
-```
-
-Excluding the respondents who did not finish the questionnaire.
-
-```{r}
-raw <- 
-  raw %>% 
-  dplyr::filter(finished) %>%
-  filter(!(response_id %in% c("R_2ZVO2mDq0eBPYDn",
-                              "R_3qr9uw1HGF53UiJ",
-                              "R_3iQHcAwCVxBXAEd",
-                              "R_1HqrkliX8SXIZqS")
-  )
-  )
-# We did a brief visual inspection of the data and feel these first three respondents fit better as partially completed surveys.
-Although they clicked through to the end of the survey, they did no answer questions after a certain point.
-For example, one respondent only answered the first two quetsions and clicked through to the end of the survey in 49 seconds.
-# The fourth response appears to be non-sincere. All their responses were neutral, several were skipped, and it was completed in 146 seconds. This respondent may have simply wanted to see the survey. 
-```
-
-At the end of the exclusion we had `r nrow(raw)` responses remaining for the analysis.
-There were `r filter(raw, has_missing) %>% nrow()` respondents who had at least one missing value for a closed-ended question.
-
 # Distribution of completition time
@@
   )
-```
-
-# Delete not used variables
-
-```{r}
-raw <-
-  raw %>% 
-  select(
-    - start_date,
-    - end_date,
-    - status,
-    - progress,
-    - finished,
-    - duration_in_seconds,
-    - recorded_date,
-    - distribution_channel,
-    - user_language
-  )
-```
-
-# Exploring the number of responses to open-ended questions
-
-The survey contained 4 open-ended questions (*unsure_explain*, *drawback_ecaw*, *suggestions*, *comments*) that we analyzed qualitatively, as preregistered.
-
-Here, we calculate the number of responses for each open ended question.
-
-```{r}
-raw %>% 
-  select(unsure_explain, drawback_ecaw, suggestions, comments) %>% 
-  # convert to long format
-  pivot_longer(
-    cols = everything(),
-    names_to = "variable",
-    values_to = "response",
-  ) %>% 
-  # flag missing responses
-  mutate(is_na = is.na(response)) %>%
-  # count missing responses per variables
-  count(variable, is_na) %>% 
-  # keep only complete responses
-  filter(!is_na)
 ```
@@
 
+# Partial responses
+
+We only include data from participants who completed the survey in the final analysis, as preregistered (see the full text of the preregistration [https://osf.io/g2fw5](https://osf.io/g2fw5)).
+There were `r nrow(dplyr::filter(raw, !finished))` respondents who did not complete the survey.
+
+However, these responses will be included in a supplementary analysis that will be reproted in Supplementary Material D. So we save them separately before exclusion.
+
+```{r}
+processed_all <- 
+  raw %>%
+  # Delete not used variables
+  select(
+    - start_date,
+    - end_date,
+    - status,
+    - progress,
+    - finished,
+    - duration_in_seconds,
+    - recorded_date,
+    - distribution_channel,
+    - user_language
+  ) %>% 
+  # Remove participants who only opened the survey but did not answer any questions. All participants who had NA for the first question (typically_trustworthy), had NAs for all questions
+  # These respondents are included from the main analysis as well see section Excluding partial responses
+  filter(!is.na(typically_trustworthy))
+
+readr::write_csv(processed_all, here("data/processed/ecaw_processed_all_data.csv"))
+```
+
+The full dataset that we are going to use to calculate the results in Supplement D includes `r nrow(processed_all)` responses.
+
+We create a simple codebook for this dataset.
+
+```{r}
+# We re using the codebook for the source dataset as a template and append it with the new variables and their description
+processed_all_codebook <-
+  read_csv(here("data/source/ecaw_221102_source_codebook.csv")) %>% 
+  select(-original_var_name) %>% 
+  rename(var_name = new_var_name) %>% 
+  filter(var_name %in% colnames(processed_all))
+
+# Adding new column descriptions
+new_rows <- tibble(
+  description = c(
+    "True if respondent has any missing responses.",
+    "Time spent from opening the survey to closing it in minutes."
+    ),
+  var_name = c(
+    "has_missing",
+    "duration_in_mins"
+    )
+)
+
+codebook <- bind_rows(processed_all_codebook, new_rows)
+
+readr::write_csv(processed_all_codebook, here("data/processed/ecaw_processed_all_codebook.csv"))
+```
+
+# Excluding partial responses
+
+Excluding the respondents who did not finish the questionnaire for the main analysis.
+
+```{r}
+raw <- 
+  raw %>% 
+  dplyr::filter(finished) %>%
+  dplyr::filter(!(response_id %in% c("R_2ZVO2mDq0eBPYDn",
+                                     "R_3qr9uw1HGF53UiJ",
+                                     "R_3iQHcAwCVxBXAEd",
+                                     "R_1HqrkliX8SXIZqS")
+                  )
+  )
+# We did a brief visual inspection of the data and feel these first three respondents fit better as partially completed surveys.
+Although they clicked through to the end of the survey, they did no answer questions after a certain point.
+For example, one respondent only answered the first two quetsions and clicked through to the end of the survey in 49 seconds.
+# The fourth response appears to be non-sincere. All their responses were neutral, several were skipped, and it was completed in 146 seconds. This respondent may have simply wanted to see the survey. 
+```
+
+# Delete not used variables
+
+```{r}
+raw <-
+  raw %>% 
+  select(
+    - start_date,
+    - end_date,
+    - status,
+    - progress,
+    - finished,
+    - duration_in_seconds,
+    - recorded_date,
+    - distribution_channel,
+    - user_language
+  )
+```
+
+At the end of the exclusion we had `r nrow(raw)` responses remaining for the analysis.
+There were `r filter(raw, has_missing) %>% nrow()` respondents who had at least one missing value for a closed-ended question.
+
+# Exploring the number of responses to open-ended questions
+
+The survey contained 4 open-ended questions (*unsure_explain*, *drawback_ecaw*, *suggestions*, *comments*) that we analyzed qualitatively, as preregistered.
+
+Here, we calculate the number of responses for each open ended question.
+
+```{r}
+raw %>% 
+  select(unsure_explain, drawback_ecaw, suggestions, comments) %>% 
+  # convert to long format
+  pivot_longer(
+    cols = everything(),
+    names_to = "variable",
+    values_to = "response",
+  ) %>% 
+  # flag missing responses
+  mutate(is_na = is.na(response)) %>%
+  # count missing responses per variables
+  count(variable, is_na) %>% 
+  # keep only complete responses
+  filter(!is_na)
+```
+
 # Creating a simple codebook
@@
 # Adding new column descriptions
-new_row1 <- tibble(
-  description = "True if respondent has any missing responses.",
-  var_name = "has_missing"
+new_rows <- tibble(
+  description = c(
+    "True if respondent has any missing responses.",
+    "Time spent from opening the survey to closing it in minutes."
+    ),
+  var_name = c(
+    "has_missing",
+    "duration_in_mins"
     )
-new_row2 <- tibble(
-  description = "Time spent from opening the survey to closing it in minutes.",
-  var_name = "duration_in_mins"
 )
 
-codebook <- bind_rows(codebook, new_row1, new_row2)
+codebook <- bind_rows(codebook, new_rows)
 

### 2023-04-28 manuscript/manuscript.Rmd
made some cosmetic changes

@@
     email: robert.thibault@stanford.edu
-    address: Enter postal address here
   - name: Marton Kovacs
@@
 abstract: |
-  \textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis.
+  \justifying\textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis.
 An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW).
@@
     latex_engine: xelatex
-    extra_dependencies: "tcolorbox"
-header-includes   :
-  - \usepackage{tcolorbox}
-  - \usepackage{caption}
-  - \captionsetup[figure]{labelformat=empty}
+    extra_dependencies: ["tcolorbox", "relsize"]
+    includes:
+      in_header: "preamble.tex"
 editor_options: 
@@
 ---
-
+```{=latex}
+\justifying
+```
 ```{r setup, include=FALSE, warning=FALSE, message=FALSE}
@@
 
-\pagebreak
-
 \begin{tcolorbox}
-\textbf{Box 1. A hypothetical example of an ECAW in practice.}
+{\fontsize{9pt}{9pt}\selectfont\textbf{Box 1. A hypothetical example of an ECAW in practice.}}
 
 \includegraphics[width=1\textwidth]{./ecaw_workflow.png}
-{\fontsize{10pt}{10pt}\selectfont
+{\fontsize{9pt}{9pt}\selectfont
 The overarching framework for ECAWs is depicted above.
@@
 It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions).
-These two factors may have deflated the response rate.} The median time to complete the survey was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).
+These two factors may have deflated the response rate.} The median time to complete the survey was `r format(round(median(processed$duration_in_mins), 1), nsmall = 1)` minutes (IQR: `r format(round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1), nsmall = 1)` to `r format(round(quantile(processed$duration_in_mins, .75, na.rm = T), 1), nsmall = 1)`).
 This manuscript presents the results for complete surveys.
@@
 
-Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 2, top panel).
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 1, top panel).
 At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
 
-```{r typicallyEcawPlot, warning=FALSE, fig.cap="(ref:typicallyEcawPlotCaption)", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
-# Create typically trustworthy and reproducible plot
-# Check if there is any missing and don't understand responses separately
-# processed %>%
-#   select(contains("typically_")) %>%
-#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
-#   nrow()
-
+```{r typicallyEcawPlot, warning=FALSE, fig.cap="(ref:typicallyEcawPlotCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
 # Prepare plot data
@@
 
+```{=latex}
+{\smaller[1] \singlespacing
+```
 (ref:typicallyEcawPlotCaption)  __Figure 1.
@@
 The bar charts in the bottom panel excluded responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible"), n)`).
+```{=latex}
+}
+```
 
@@
 
-```{r methodPlot, warning=FALSE, fig.cap="(ref:methodPlotCaption)", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
+```{r methodPlot, warning=FALSE, fig.cap="(ref:methodPlotCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
 # Used methods figures
@@
 
+```{=latex}
+{\smaller[1] \singlespacing
+```
 (ref:methodPlotCaption) __Figure 2. Responses to survey questions about the research practices of participants.__
+```{=latex}
+}
+```
 
@@
 
-```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotCaption)", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
+```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
 # Check if there is any missing and don't understand responses separately
@@
 
+```{=latex}
+{\smaller[1] \singlespacing
+```
 (ref:alspacPlotCaption) __Figure 3.
@@
 Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[4].
+```{=latex}
+}
+```
 
@@
 
-__Table 1.
-Recurring topics in responses to the open-ended survey questions.__ The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
+\pagebreak
+
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
+___Table 1.
+Recurring topics in responses to the open-ended survey questions.___ The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
 These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents.
@@
 On the right side of the table, we provide our reflection on each topic.
-
-```{r openEndedTable, echo=FALSE, message=FALSE, warnings=FALSE, fig.align = "center", out.width="100%", fig.id=FALSE}
+```{=latex}
+\end{singlespace}
+}
+```
+
+```{r openEndedTable, echo=FALSE, message=FALSE, warnings=FALSE, fig.align="center", out.width="100%", out.height="300%", fig.id=FALSE}
 open_ended_table <- tibble(
@@
   ) %>% 
-  tab_options(table.width = pct(90)) %>% 
-  gtsave("figs/open-ended-table_figure.png", expand = c(0, 100, 0, 100), zoom = 2, vwidth = 1300)
+  tab_options(
+    table.width = pct(100)
+    # table.font.size = 12
+    ) %>% 
+  gtsave("figs/open-ended-table_figure.png", expand = c(0, 50, 0, 50), vwidth = 1300, vheight=1500)
 
@@
 
+\pagebreak
+
 ## Exploratory analyses
-Results can be explored interactively at [TEMPORARY LINK], or by running the code available at [INSERT DATA LINK].
+Results can be explored interactively by following the instructions for running the code at [INSERT DATA LINK].
 Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
@@
 ## Contributions
-__Conceptualization:__ Robert T. Thibault and Marcus R. Munafò.
-__Data curation:__ Marton Kovacs.
-__Formal analysis:__ Robert T. Thibault and Marton Kovacs.
-__Funding acquisition:__ Robert T. Thibault and John P. A. Ioannidis.
-__Investigation:__ Robert T. Thibault.
-__Methodology:__ Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
-__Project administration:__ Robert T. Thibault.
-__Resources:__ Robert T. Thibault.
-__Software:__ Marton Kovacs.
-__Supervision:__ Robert T. Thibault and Marcus R. Munafò.
-__Validation:__ Robert T. Thibault and Marton Kovacs.
-__Visualization:__ Robert T. Thibault and Marton Kovacs.
-__Writing - original draft:__ Robert T. Thibault.
+__Conceptualization:__ Robert T. Thibault and Marcus R. Munafò.\hfill\break
+__Data curation:__ Marton Kovacs.\hfill\break
+__Formal analysis:__ Robert T. Thibault and Marton Kovacs.\hfill\break
+__Funding acquisition:__ Robert T. Thibault and John P. A. Ioannidis.\hfill\break
+__Investigation:__ Robert T. Thibault.\hfill\break
+__Methodology:__ Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.\hfill\break
+__Project administration:__ Robert T. Thibault.\hfill\break
+__Resources:__ Robert T. Thibault.\hfill\break
+__Software:__ Marton Kovacs.\hfill\break
+__Supervision:__ Robert T. Thibault and Marcus R. Munafò.\hfill\break
+__Validation:__ Robert T. Thibault and Marton Kovacs.\hfill\break
+__Visualization:__ Robert T. Thibault and Marton Kovacs.\hfill\break
+__Writing - original draft:__ Robert T. Thibault.\hfill\break
 __Writing - review & editing:__ Robert T. Thibault, Marton Kovacs, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
@@
 
-# (APPENDIX) Appendix {-}
-
 ```{r child = "supplementary_materials.Rmd"}

### 2023-04-28 manuscript/manuscript.tex
made some cosmetic changes

@@
 \def\languageshorthands#1{}
+\usepackage{tcolorbox}
+\usepackage{caption}
+\usepackage{setspace}
+\usepackage{relsize}
+\usepackage[document,bottom]{ragged2e}
+\captionsetup[figure]{labelformat=empty}
+\captionsetup[table]{labelformat=empty}
+\raggedbottom
 % Manuscript styling
@@
 \usepackage{csquotes}
-\usepackage{tcolorbox}
-\usepackage{caption}
-\captionsetup[figure]{labelformat=empty}
 \ifLuaTeX
@@
 
-Correspondence concerning this article should be addressed to Robert T.
-Thibault, Enter postal address here.
-E-mail: \href{mailto:robert.thibault@stanford.edu}{\nolinkurl{robert.thibault@stanford.edu}}
+Correspondence concerning this article should be addressed to Robert T. Thibault. E-mail: \href{mailto:robert.thibault@stanford.edu}{\nolinkurl{robert.thibault@stanford.edu}}
 
@@
 \abstract{%
-\textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis.
+\justifying\textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis.
 An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW).
@@
 \maketitle
+
+\justifying
 
@@
 It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions).
-These two factors may have deflated the response rate.} The median time to complete the survey was 7.40 minutes (IQR: 4.60 to 13.10).
+These two factors may have deflated the response rate.} The median time to complete the survey was 7.4 minutes (IQR: 4.6 to 13.1).
 This manuscript presents the results for complete surveys.
@@
 
-Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 74/103) and reproducible (79\%; 81/103) (Figure 2, top panel).
+Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 74/103) and reproducible (79\%; 81/103) (Figure 1, top panel).
 At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
+{\smaller[1] \singlespacing
+
+}
+
 Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2).
@@
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
+{\smaller[1] \singlespacing
+
+}
+
 26\% (26/101) of respondents agreed (versus 45\%; 45/101 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3).
@@
 
+{\smaller[1] \singlespacing
+
+}
+
 \protect\footnotetext[3]{Another four respondents agreed with at least the first and second question, which appear contradictory.
@@
 
-\textbf{Table 1.
-Recurring topics in responses to the open-ended survey questions.} The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
+\pagebreak
+
+{\smaller[1]
+\begin{singlespace}
+
+\textbf{\emph{Table 1.
+Recurring topics in responses to the open-ended survey questions.}} The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
 These questions received a total of (92) responses from (55) unique respondents.
@@
 
+\end{singlespace}
+}
+
 \begin{figure}
 
-{\centering \includegraphics[width=1\linewidth]{figs/open-ended-table_figure}
+{\centering \includegraphics[width=1\linewidth,height=3\textheight]{figs/open-ended-table_figure}
 
@@
 
+\pagebreak
+
 \hypertarget{exploratory-analyses}{%
@@
 
-Results can be explored interactively at {[}TEMPORARY LINK{]}, or by running the code available at {[}INSERT DATA LINK{]}.
+Results can be explored interactively by following the instructions for running the code at {[}INSERT DATA LINK{]}.
 Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
@@
 
-\textbf{Conceptualization:} Robert T. Thibault and Marcus R. Munafò.
-\textbf{Data curation:} Marton Kovacs.
-\textbf{Formal analysis:} Robert T. Thibault and Marton Kovacs.
-\textbf{Funding acquisition:} Robert T. Thibault and John P. A. Ioannidis.
-\textbf{Investigation:} Robert T. Thibault.
-\textbf{Methodology:} Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
-\textbf{Project administration:} Robert T. Thibault.
-\textbf{Resources:} Robert T. Thibault.
-\textbf{Software:} Marton Kovacs.
-\textbf{Supervision:} Robert T. Thibault and Marcus R. Munafò.
-\textbf{Validation:} Robert T. Thibault and Marton Kovacs.
-\textbf{Visualization:} Robert T. Thibault and Marton Kovacs.
-\textbf{Writing - original draft:} Robert T. Thibault.
+\textbf{Conceptualization:} Robert T. Thibault and Marcus R. Munafò.\hfill\break
+\textbf{Data curation:} Marton Kovacs.\hfill\break
+\textbf{Formal analysis:} Robert T. Thibault and Marton Kovacs.\hfill\break
+\textbf{Funding acquisition:} Robert T. Thibault and John P. A. Ioannidis.\hfill\break
+\textbf{Investigation:} Robert T. Thibault.\hfill\break
+\textbf{Methodology:} Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.\hfill\break
+\textbf{Project administration:} Robert T. Thibault.\hfill\break
+\textbf{Resources:} Robert T. Thibault.\hfill\break
+\textbf{Software:} Marton Kovacs.\hfill\break
+\textbf{Supervision:} Robert T. Thibault and Marcus R. Munafò.\hfill\break
+\textbf{Validation:} Robert T. Thibault and Marton Kovacs.\hfill\break
+\textbf{Visualization:} Robert T. Thibault and Marton Kovacs.\hfill\break
+\textbf{Writing - original draft:} Robert T. Thibault.\hfill\break
 \textbf{Writing - review \& editing:} Robert T. Thibault, Marton Kovacs, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
@@
 
-\hypertarget{appendix-appendix}{%
-\appendix}
-
 \hypertarget{supplementary-material-a.-deviations-from-the-preregistration.}{%
@@
 \begin{enumerate}
+
   \item We rephrased our second objective to be more accurate.
@@
   \item We did not preregister which results we would present in the abstract.
-We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 25% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
+We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 25\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
+
 \end{enumerate}
@@
 \hypertarget{supplementary-material-c.-participant-characteristics}{%
-\subsection{Supplementary Material C. Participant Characteristics}\label{supplementary-material-c.-participant-characteristics}}
+\section{Supplementary Material C. Participant Characteristics}\label{supplementary-material-c.-participant-characteristics}}
 
@@
 
+{\smaller[1]
+\begin{singlespace}
+
+\end{singlespace}
+}
+
 \begin{figure}
@@
 
+{\smaller[1]
+\begin{singlespace}
+
+\end{singlespace}
+}
+
+\pagebreak
+
 \hypertarget{supplementary-material-d.-figures-and-tables-including-participants-who-partially-complete-the-survey}{%
-\subsection{Supplementary Material D.
+\section{Supplementary Material D.
 Figures and Tables including participants who partially complete the survey}\label{supplementary-material-d.-figures-and-tables-including-participants-who-partially-complete-the-survey}}
@@
 \hypertarget{participants-2}{%
-\subsubsection{Participants}\label{participants-2}}
+\subsection{Participants}\label{participants-2}}
 
-Respondents had published a median of 10 (IQR 2 to 25) studies using preexisting observational data (Supplementary Figure C1).
-They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Supplementary Table C1)\footnote[1]{Participants could select multiple responses to this survey question.}.
-61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/101) reported being less concerned (Supplementary Figure C2).
+Respondents had published a median of 10 (IQR 2 to 25) studies using preexisting observational data.
+They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) \footnote[1]{Participants could select multiple responses to this survey question.}.
+61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/101) reported being less concerned.
 
@@
 
+{\smaller[1]
+\begin{singlespace}
+
+\end{singlespace}
+}
+
 Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (33\%; 41/123), or sometimes (21\%; 26/103) (Figure 2D).
@@
 
+{\smaller[1]
+\begin{singlespace}
+
+\end{singlespace}
+}
+
 25\% (26/106) of respondents agreed (versus 44\%; 47/106 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3D).
@@
 
+{\smaller[1]
+\begin{singlespace}
+
+\end{singlespace}
+}
+
 \protect\footnotetext[2]{Another four respondents agreed with at least the first and second question, which appear contradictory.

### 2023-04-28 manuscript/supplementary_materials.Rmd
made some cosmetic changes

@@
 # Supplementary Material B. Invitation email
-
 ## Original email
@@
 
-## Supplementary Material C. Participant Characteristics
+# Supplementary Material C. Participant Characteristics
 
@@
 
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
 (ref:table1c-caption) __Table C1. “What programming language or software do you use for your analyses of preexisting observational data? (you may select multiple answers)”__
+```{=latex}
+\end{singlespace}
+}
+```
 
@@
 
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
 (ref:concernedPlotCaption) __Figure C1. Responses to survey question about number of relevant publications__
+```{=latex}
+\end{singlespace}
+}
+```
 
-## Supplementary Material D. Figures and Tables including participants who partially complete the survey
-### Participants
+\pagebreak
+
+# Supplementary Material D. Figures and Tables including participants who partially complete the survey
+## Participants
 
@@
 
-Respondents had published a median of `r median(processed_all$n_studies, na.rm = TRUE)` (IQR `r round(quantile(processed_all$n_studies, .25, na.rm = TRUE))` to `r round(quantile(processed_all$n_studies, .75, na.rm = TRUE))`) studies using preexisting observational data (Supplementary Figure C1).
-They reported using the following programming languages or software packages: `r language_text_all` (Supplementary Table C1)\footnote[1]{Participants could select multiple responses to this survey question.}.
-`r pull(filter(concerned_all, support == "positive"), percentage)`% (`r pull(filter(concerned_all, support == "positive"), n)`/`r pull(filter(concerned_all, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; `r pull(filter(concerned_all, support == "negative"), percentage)`% (`r pull(filter(concerned_all, support == "negative"), n)`/`r pull(filter(concerned_all, support == "negative"), n_sum)`) reported being less concerned (Supplementary Figure C2).
+Respondents had published a median of `r median(processed_all$n_studies, na.rm = TRUE)` (IQR `r round(quantile(processed_all$n_studies, .25, na.rm = TRUE))` to `r round(quantile(processed_all$n_studies, .75, na.rm = TRUE))`) studies using preexisting observational data.
+They reported using the following programming languages or software packages: `r language_text_all` \footnote[1]{Participants could select multiple responses to this survey question.}.
+`r pull(filter(concerned_all, support == "positive"), percentage)`% (`r pull(filter(concerned_all, support == "positive"), n)`/`r pull(filter(concerned_all, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; `r pull(filter(concerned_all, support == "negative"), percentage)`% (`r pull(filter(concerned_all, support == "negative"), n)`/`r pull(filter(concerned_all, support == "negative"), n_sum)`) reported being less concerned.
 
@@
 
-```{r typicallyEcawPlotAll, warning=FALSE, fig.cap="(ref:typicallyEcawPlotAllCaption)", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
+```{r typicallyEcawPlotAll, warning=FALSE, fig.cap="(ref:typicallyEcawPlotAllCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
 # Create typically trustworthy and reproducible plot
@@
 
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
 (ref:typicallyEcawPlotAllCaption) __Figure 1D.
@@
 The bar charts in the bottom panel excluded missing responses (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "Missing"), n)`) and responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "I don't understand the question"), n)`).
+```{=latex}
+\end{singlespace}
+}
+```
 
@@
 
-```{r methodPlotAll, warning=FALSE, fig.cap="(ref:methodPlotAllCaption)", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
+```{r methodPlotAll, warning=FALSE, fig.cap="(ref:methodPlotAllCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
 # Used methods figures
@@
 
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
 (ref:methodPlotAllCaption) __Figure 2D. Responses to survey questions about the research practices of participants.__
+```{=latex}
+\end{singlespace}
+}
+```
 
@@
 
-```{r alspacPlotAll, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotAllCaption)", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
+```{r alspacPlotAll, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotAllCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
 # Check if there is any missing and don't understand responses separately
@@
 
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
 (ref:alspacPlotAllCaption) __Figure 3D.
@@
 Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[3].
+```{=latex}
+\end{singlespace}
+}
+```
 

### 2023-04-28 manuscript/supplementary_materials.Rmd
changed faulty figure numbers

@@
 
-```{r numStudiesPlot, fig.align="center", out.width="100%", fig.width=6, fig.height=3.2, fig.path='figs/', dev=c('png', 'pdf')}
+```{r numStudiesPlot, fig.align="center", fig.cap="(ref:numStudiesPlotCaption)", out.width="100%", fig.width=6, fig.height=3.2, fig.path='figs/', dev=c('png', 'pdf')}
 # Create text for missing responses
@@
 #   papaja::theme_apa()
+```
+
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
+(ref:numStudiesPlotCaption) __Figure C1.__
+```{=latex}
+\end{singlespace}
+}
 ```
@@
 ```
-(ref:concernedPlotCaption) __Figure C1. Responses to survey question about number of relevant publications__
+(ref:concernedPlotCaption) __Figure C2. Responses to survey question about number of relevant publications__
 ```{=latex}

### 2023-04-28 manuscript/manuscript.Rmd
fixing errors

@@
 floatsintext      : yes
-linenumbers       : yes
+linenumbers       : no
 draft             : no
@@
       "Always or almost always",
-      "Missing",
+      # "Missing", Since there are no missing values
       "I don't understand the question"
@@
 ## Exploratory analyses
-Results can be explored interactively by following the instructions for running the code at [INSERT DATA LINK].
-Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
+Results can be explored interactively by following the instructions for running the code at https://github.com/RobbyTi/ECAW.
+Based on the sample size of `r nrow(processed)` participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
 
@@
 Third, ECAWs appear relatively acceptable.
-For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs, and 25% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
+For example, only `r pull(filter(study, support == "negative"), percentage)`% of respondents disagreed that ALSPAC should run a study on ECAWs, and `r pull(filter(willing, support == "negative"), percentage)`% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
 
@@
 The survey results are best understood as the initial thoughts of participants when introduced to the concept of ECAWs.
-The median response time was 7.2 minutes (IQR: 4.3 to 13.1) for a survey that included over 20 questions and a 500-word description of ECAWs.
+The median response time was `r format(round(median(processed$duration_in_mins), 1), nsmall = 1)` minutes (IQR: `r format(round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1), nsmall = 1)` to `r format(round(quantile(processed$duration_in_mins, .75, na.rm = T), 1), nsmall = 1)`) for a survey that included over 20 questions and a 500-word description of ECAWs.
 Thus, it is unlikely that respondents spent much time reflecting on ECAWs and their implications.

### 2023-04-28 manuscript/manuscript.tex
changed faulty figure numbers
fixing errors

@@
 \keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 4020}
-\usepackage{lineno}
-
-\linenumbers
 \usepackage{csquotes}
@@
 
-Results can be explored interactively by following the instructions for running the code at {[}INSERT DATA LINK{]}.
-Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
+Results can be explored interactively by following the instructions for running the code at \url{https://github.com/RobbyTi/ECAW}.
+Based on the sample size of 103 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
 
@@
 Third, ECAWs appear relatively acceptable.
-For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, and 25\% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
+For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, and 45\% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
 
@@
 The survey results are best understood as the initial thoughts of participants when introduced to the concept of ECAWs.
-The median response time was 7.2 minutes (IQR: 4.3 to 13.1) for a survey that included over 20 questions and a 500-word description of ECAWs.
+The median response time was 7.4 minutes (IQR: 4.6 to 13.1) for a survey that included over 20 questions and a 500-word description of ECAWs.
 Thus, it is unlikely that respondents spent much time reflecting on ECAWs and their implications.
@@
 
-\caption{ }\label{fig:numStudiesPlot}
+\caption{\textbf{Figure C1.} Responses to survey question about number of relevant publications}\label{fig:numStudiesPlot}
 \end{figure}
+
+{\smaller[1]
+\begin{singlespace}
+
+\end{singlespace}
+}
 
@@
 
-\caption{\textbf{Figure C1. Responses to survey question about number of relevant publications}}\label{fig:concernedPlot}
+\caption{\textbf{Figure C2. Responses to survey question about number of relevant publications}}\label{fig:concernedPlot}
 \end{figure}
@@
 Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (71\%; 87/123) and reproducible (77\%; 93/121) (Figure D2, top panel).
-At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/107) and \emph{more} reproducible (69\%; 74/108) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 75/107) and \emph{more} reproducible (69\%; 74/108) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
 
@@
 
-\caption{\textbf{Figure 1D.
+\caption{\textbf{Figure D1.
 Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
@@
 
-\caption{\textbf{Figure 2D. Responses to survey questions about the research practices of participants.}}\label{fig:methodPlotAll}
+\caption{\textbf{Figure D2. Responses to survey questions about the research practices of participants.}}\label{fig:methodPlotAll}
 \end{figure}
@@
 
-\caption{\textbf{Figure 3D.
+\caption{\textbf{Figure D3.
 Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 14; 14; 14; 14; respectively from top to bottom), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 5; 2; 1), and responses of \emph{``Unsure''} (\emph{n} = 3; 6; 7; 8).
@@
 
+\pagebreak
+
+\hypertarget{supplementary-material-e.-description-of-alspac-study-numbers}{%
+\section{Supplementary Material E. Description of ALSPAC study numbers}\label{supplementary-material-e.-description-of-alspac-study-numbers}}
+
+Pregnant women resident in Avon, UK with expected dates of delivery between 1st April 1991 and 31st December 1992 were invited to take part in the study.
+20,248 pregnancies have been identified as being eligible and the initial number of pregnancies enrolled was 14,541.
+Of the initial pregnancies, there was a total of 14,676 foetuses, resulting in 14,062 live births and 13,988 children who were alive at 1 year of age.
+When the oldest children were approximately 7 years of age, an attempt was made to bolster the initial sample with eligible cases who had failed to join the study originally.
+As a result, when considering variables collected from the age of seven onwards (and potentially abstracted from obstetric notes) there are data available for more than the 14,541 pregnancies mentioned above: The number of new pregnancies not in the initial sample (known as Phase I enrolment) that are currently represented in the released data and reflecting enrolment status at the age of 24 is 906, resulting in an additional 913 children being enrolled (456, 262 and 195 recruited during Phases II, III and IV respectively).
+The phases of enrolment are described in more detail in the cohort profile paper and its update (see footnote 5 below).
+The total sample size for analyses using any data collected after the age of seven is therefore 15,447 pregnancies, resulting in 15,658 foetuses.
+Of these 14,901 children were alive at 1 year of age.
+
+Of the original 14,541 initial pregnancies, 338 were from a woman who had already enrolled with a previous pregnancy, meaning 14,203 unique mothers were initially enrolled in the study.
+As a result of the additional phases of recruitment, a further 630 women who did not enrol originally have provided data since their child was 7 years of age.
+This provides a total of 14,833 unique women (G0 mothers) enrolled in ALSPAC as of September 2021.
+
+G0 partners were invited to complete questionnaires by the mothers at the start of the study and they were not formally enrolled at that time.
+12,113 G0 partners have been in contact with the study by providing data and/or formally enrolling when this started in 2010.
+3,807 G0 partners are currently enrolled.
+
 \end{document}

### 2023-04-28 manuscript/supplementary_materials.Rmd
changed numbering in D
fixing errors

@@
 ```
-(ref:numStudiesPlotCaption) __Figure C1.__
+(ref:numStudiesPlotCaption) __Figure C1.__ Responses to survey question about number of relevant publications
 ```{=latex}
@@
 Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy_all, support == "positive"), n)`/`r pull(filter(typically_trustworthy_all, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible_all, support == "positive"), n)`/`r pull(filter(typically_reproducible_all, support == "positive"), n_sum)`) (Figure D2, top panel).
-At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy_all, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible_all, support == "positive"), n)`/`r pull(filter(ecaw_reproducible_all, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy_all, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy_all, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible_all, support == "positive"), n)`/`r pull(filter(ecaw_reproducible_all, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
 
@@
 ```
-(ref:typicallyEcawPlotAllCaption) __Figure 1D.
+(ref:typicallyEcawPlotAllCaption) __Figure D1.
 Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
@@
 ```
-(ref:methodPlotAllCaption) __Figure 2D. Responses to survey questions about the research practices of participants.__
+(ref:methodPlotAllCaption) __Figure D2. Responses to survey questions about the research practices of participants.__
 ```{=latex}
@@
 ```
-(ref:alspacPlotAllCaption) __Figure 3D.
+(ref:alspacPlotAllCaption) __Figure D3.
 Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Missing"), n)`; respectively from top to bottom), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Unsure"), n)`).
@@
 However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}
+
+\pagebreak
+
+# Supplementary Material E. Description of ALSPAC study numbers
+
+Pregnant women resident in Avon, UK with expected dates of delivery between 1st April 1991 and 31st December 1992 were invited to take part in the study.
+20,248 pregnancies have been identified as being eligible and the initial number of pregnancies enrolled was 14,541.
+Of the initial pregnancies, there was a total of 14,676 foetuses, resulting in 14,062 live births and 13,988 children who were alive at 1 year of age.
+When the oldest children were approximately 7 years of age, an attempt was made to bolster the initial sample with eligible cases who had failed to join the study originally.
+As a result, when considering variables collected from the age of seven onwards (and potentially abstracted from obstetric notes) there are data available for more than the 14,541 pregnancies mentioned above: The number of new pregnancies not in the initial sample (known as Phase I enrolment) that are currently represented in the released data and reflecting enrolment status at the age of 24 is 906, resulting in an additional 913 children being enrolled (456, 262 and 195 recruited during Phases II, III and IV respectively).
+The phases of enrolment are described in more detail in the cohort profile paper and its update (see footnote 5 below).
+The total sample size for analyses using any data collected after the age of seven is therefore 15,447 pregnancies, resulting in 15,658 foetuses.
+Of these 14,901 children were alive at 1 year of age.
+
+Of the original 14,541 initial pregnancies, 338 were from a woman who had already enrolled with a previous pregnancy, meaning 14,203 unique mothers were initially enrolled in the study.
+As a result of the additional phases of recruitment, a further 630 women who did not enrol originally have provided data since their child was 7 years of age.
+This provides a total of 14,833 unique women (G0 mothers) enrolled in ALSPAC as of September 2021.
+
+G0 partners were invited to complete questionnaires by the mothers at the start of the study and they were not formally enrolled at that time.
+12,113 G0 partners have been in contact with the study by providing data and/or formally enrolling when this started in 2010.
+3,807 G0 partners are currently enrolled.

### 2023-04-28 manuscript/manuscript.Rmd
changes

@@
 \hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC’s transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}.
-We received 107 responses, for a 10% response rate.
+We received 103 responses, for a 9% response rate.
 The results suggest that—at least among our sample of respondents—ECAWs hold the potential to serve their intended purpose and appear relatively acceptable.
 For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55% who agreed).
-However, as many as 25% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions.
+However, as many as 26% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 45% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions.
 \hfill \break \hfill \break \textbf{Preregistration:} https://osf.io/g2fw5 Deviations from the preregistration are outlined in Supplementary Material A.
@@
 keywords          : "Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)"
-wordcount         : "4020"
+wordcount         : "5036"
 bibliography      : "ecaw-ref.bib"
@@
 These shortcomings may stem from a combination of researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices.
-These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see @prasad_ending_2019).
+These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions [e.g., see @prasad_ending_2019].
 
@@
 ECAWs may increase the uptake of preregistration and provide assurance that the analyses were developed before the researchers observed the full dataset.
-Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., @bakker_ensuring_2020), ECAWs provide an executable analysis script.
+Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed [e.g., @bakker_ensuring_2020], ECAWs provide an executable analysis script.
 Researchers may also find ECAWs more agreeable than conventional preregistration because they allow for exploration.
@@
 ## Examples of ECAW-like workflows
-One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (@sarafoglou_comparing_2023).
+One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset [@sarafoglou_comparing_2023].
 The researchers recruited 120 teams to analyze a single observational dataset and had half the teams preregister a written analysis plan and the other half prepare an analysis plan by drafting an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded).
@@
 
-Some disciplines, such as particle physics, implement blind data analyses regularly (@klein_blind_2005; @maccoun_blind_2015).
-A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (@Forscher2020) and is currently being used for Registered Reports based on a large psychology dataset (@Schmidt2023).
+Some disciplines, such as particle physics, implement blind data analyses regularly [@klein_blind_2005; @maccoun_blind_2015].
+A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group [@Forscher2020] and is currently being used for Registered Reports based on a large psychology dataset [@Schmidt2023].
 
@@
 ALSPAC is “a transgenerational prospective observational study investigating influences on health and development across the life course.
-It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.” (Boyd et al., 2013; Fraser et al., 2013).
+It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.” [@boyd_cohort_2013; @fraser_cohort_2013].
 Thus, our invitation reached researchers that use observational data across the health and social sciences.
@@
   filter(lang != "irrelevant") %>%
-  mutate(lang_n = glue::glue("{lang} (n = {n})")) %>%
+  mutate(lang_n = glue::glue("{lang} (_n_ = {n})")) %>%
   pull(lang_n) %>%
@@
 Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 1, top panel).
-At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+At the same time, many believed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using pre-existing observational data (Figure 1, bottom panel).
 
@@
 The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_.
-The bar charts in the bottom panel excluded responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible"), n)`).
+The bar charts in the bottom panel excluded responses of _“I don't understand the question”_ (_n_ = `r pull(filter(ecaw_missing, item == "Trustworthy"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible"), n)`;  respectively from top to bottom).
 ```{=latex}
@@
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2).
+Over half of respondents reported that their studies using pre-existing observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2).
 About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script, method_script == "Never or almost never") %>% pull(n)`/`r filter(script, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script, method_script == "Sometimes") %>% pull(n)`/`r filter(script, method_script == "Sometimes") %>% pull(n_sum)`).
 `r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst.
-Almost all respondents answered that they use both exploratory (`r pull(exploratory, percentage)`%; `r pull(exploratory, n)`/`r pull(exploratory, n_sum)`) and confirmatory (`r pull(confirmatory, percentage)`%; `r pull(confirmatory, n)`/`r pull(confirmatory, n_sum)`) analyses at least sometimes.
+Almost all respondents answered that they use both confirmatory (`r pull(confirmatory, percentage)`%; `r pull(confirmatory, n)`/`r pull(confirmatory, n_sum)`) and exploratory (`r pull(exploratory, percentage)`%; `r pull(exploratory, n)`/`r pull(exploratory, n_sum)`) analyses at least sometimes.
 
@@
 `r pull(filter(study, support == "positive"), percentage)`% (`r pull(filter(study, support == "positive"), n)`/`r pull(filter(study, support == "positive"), n_sum)`) agreed (`r pull(filter(study, support == "negative"), percentage)`%; `r pull(filter(study, support == "negative"), n)`/`r pull(filter(study, support == "negative"), n_sum)` disagreed) that ALSPAC should run a study on ECAWs.
-`r pull(filter(prefer, support == "positive"), percentage)`% (`r pull(filter(prefer, support == "positive"), n)`/`r pull(filter(prefer, support == "positive"), n_sum)`) agreed (`r pull(filter(prefer, support == "negative"), percentage)`%; `r pull(filter(prefer, support == "negative"), n)`/`r pull(filter(prefer, support == "negative"), n_sum)` disagreed) that they would prefer using an ECAW than using typical preregistration.
+Finally, `r pull(filter(prefer, support == "positive"), percentage)`% (`r pull(filter(prefer, support == "positive"), n)`/`r pull(filter(prefer, support == "positive"), n_sum)`) agreed (`r pull(filter(prefer, support == "negative"), percentage)`%; `r pull(filter(prefer, support == "negative"), n)`/`r pull(filter(prefer, support == "negative"), n_sum)` disagreed) that they would prefer using an ECAW than using conventional preregistration.
 
@@
 (ref:alspacPlotCaption) __Figure 3.
-Responses to survey questions about using ECAWs.__ These bar charts exclude responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`; respectively from top to bottom), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
+Responses to survey questions about using ECAWs.__ These bar charts exclude responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).
 Agreement with the first question may be slightly inflated due to the format of the questions in this block.
@@
 Recurring topics in responses to the open-ended survey questions.___ The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
-These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents.
+These questions received a total of `r nrow(open_ended)` responses from `r nrow(distinct(open_ended, response_id))` unique respondents.
 A complete list of responses are viewable in the open data [LINK].
@@
 ## Exploratory analyses
-Results can be explored interactively by following the instructions for running the code at https://github.com/RobbyTi/ECAW.
+Results can be explored interactively by running the code available at https://github.com/RobbyTi/ECAW.
 Based on the sample size of `r nrow(processed)` participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
@@
 Third, ECAWs appear relatively acceptable.
-For example, only `r pull(filter(study, support == "negative"), percentage)`% of respondents disagreed that ALSPAC should run a study on ECAWs, and `r pull(filter(willing, support == "negative"), percentage)`% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
+For example, only `r pull(filter(study, support == "negative"), percentage)`% of respondents disagreed that ALSPAC should run a study on ECAWs, and `r pull(filter(willing, support == "positive"), percentage)`% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
 
@@
 ## Limitations
-Given the response rate of 10%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices.
+Given the response rate of `r round(nrow(processed) / 1094 * 100)`%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices.
 Indeed, respondents themselves believed that they were more concerned about reproducibility than other researchers in their field (Supplementary Figure C2).
@@
 The stakes are lower in cases where a static final dataset already exists and concerns about funding are absent.
-For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows (e.g., @Forscher2020; @Schmidt2023).
+For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows [e.g., @Forscher2020; @Schmidt2023].
 Concerns about a reduction in user numbers and engagement may also be less relevant for datasets containing unique data.

### 2023-04-28 manuscript/manuscript.tex
changes

@@
 {}{\PatchFailed}
-\keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 4020}
+\keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 5036}
 \usepackage{csquotes}
@@
 \hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC's transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}.
-We received 107 responses, for a 10\% response rate.
+We received 103 responses, for a 9\% response rate.
 The results suggest that---at least among our sample of respondents---ECAWs hold the potential to serve their intended purpose and appear relatively acceptable.
 For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55\% who agreed).
-However, as many as 25\% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44\% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions.
+However, as many as 26\% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 45\% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions.
 \hfill \break \hfill \break \textbf{Preregistration:} \url{https://osf.io/g2fw5} Deviations from the preregistration are outlined in Supplementary Material A.
@@
 These shortcomings may stem from a combination of researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices.
-These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see Prasad and Cifu (2019)).
+These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see Prasad \& Cifu, 2019).
 
@@
 ECAWs may increase the uptake of preregistration and provide assurance that the analyses were developed before the researchers observed the full dataset.
-Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., Bakker et al.
-(2020)), ECAWs provide an executable analysis script.
+Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., Bakker et al., 2020), ECAWs provide an executable analysis script.
 Researchers may also find ECAWs more agreeable than conventional preregistration because they allow for exploration.
@@
 
-One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (Sarafoglou, Hoogeveen, and Wagenmakers (2023)).
+One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (Sarafoglou, Hoogeveen, \& Wagenmakers, 2023).
 The researchers recruited 120 teams to analyze a single observational dataset and had half the teams preregister a written analysis plan and the other half prepare an analysis plan by drafting an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded).
@@
 
-Some disciplines, such as particle physics, implement blind data analyses regularly (Klein and Roodman (2005); MacCoun and Perlmutter (2015)).
-A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al.
-(2020)) and is currently being used for Registered Reports based on a large psychology dataset (Schmidt et al.
-(2023)).
+Some disciplines, such as particle physics, implement blind data analyses regularly (Klein \& Roodman, 2005; MacCoun \& Perlmutter, 2015).
+A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al., 2020) and is currently being used for Registered Reports based on a large psychology dataset (Schmidt et al., 2023).
 
@@
 Respondents had published a median of 10 (IQR 2 to 26) studies using preexisting observational data (Supplementary Figure C1).
-They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}.
+They reported using the following programming languages or software packages: R (\emph{n} = 65), Stata (\emph{n} = 48), SPSS (\emph{n} = 17), SAS (\emph{n} = 15), Python (\emph{n} = 6), Mplus (\emph{n} = 3), Bash (\emph{n} = 2), MATLAB (\emph{n} = 1), Nextflow (\emph{n} = 1), and plink2 (\emph{n} = 1) (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}.
 62\% (62/100) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/100) reported being less concerned (Supplementary Figure C2).
@@
 Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 74/103) and reproducible (79\%; 81/103) (Figure 1, top panel).
-At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+At the same time, many believed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using pre-existing observational data (Figure 1, bottom panel).
 
@@
 The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}.
-The bar charts in the bottom panel excluded responses of \emph{``I don't understand the question''} (n = 3; 2).}\label{fig:typicallyEcawPlot}
+The bar charts in the bottom panel excluded responses of \emph{``I don't understand the question''} (\emph{n} = 3; 2; respectively from top to bottom).}\label{fig:typicallyEcawPlot}
 \end{figure}
@@
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2).
+Over half of respondents reported that their studies using pre-existing observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2).
 About half reported sharing their analysis scripts never or almost never (20\%; 21/103), or sometimes (32\%; 33/103).
 77\% (79/103) reported that they never or almost never blind the data analyst.
-Almost all respondents answered that they use both exploratory (93\%; 96/103) and confirmatory (87\%; 90/103) analyses at least sometimes.
+Almost all respondents answered that they use both confirmatory (87\%; 90/103) and exploratory (93\%; 96/103) analyses at least sometimes.
 
@@
 55\% (53/96) agreed (10\%; 10/96 disagreed) that ALSPAC should run a study on ECAWs.
-46\% (43/94) agreed (22\%; 21/94 disagreed) that they would prefer using an ECAW than using typical preregistration.
+Finally, 46\% (43/94) agreed (22\%; 21/94 disagreed) that they would prefer using an ECAW than using conventional preregistration.
 
@@
 \caption{\textbf{Figure 3.
-Responses to survey questions about using ECAWs.} These bar charts exclude responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1; respectively from top to bottom), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8).
+Responses to survey questions about using ECAWs.} These bar charts exclude responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8).
 Agreement with the first question may be slightly inflated due to the format of the questions in this block.
@@
 Recurring topics in responses to the open-ended survey questions.}} The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments.
-These questions received a total of (92) responses from (55) unique respondents.
+These questions received a total of 92 responses from 55 unique respondents.
 A complete list of responses are viewable in the open data {[}LINK{]}.
@@
 
-Results can be explored interactively by following the instructions for running the code at \url{https://github.com/RobbyTi/ECAW}.
+Results can be explored interactively by running the code available at \url{https://github.com/RobbyTi/ECAW}.
 Based on the sample size of 103 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.
@@
 Third, ECAWs appear relatively acceptable.
-For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, and 45\% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
+For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, and 26\% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.
 
@@
 
-Given the response rate of 10\%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices.
+Given the response rate of 9\%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices.
 Indeed, respondents themselves believed that they were more concerned about reproducibility than other researchers in their field (Supplementary Figure C2).
@@
 The stakes are lower in cases where a static final dataset already exists and concerns about funding are absent.
-For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows (e.g., Forscher et al.
-(2020); Schmidt et al.
-(2023)).
+For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows (e.g., Forscher et al., 2020; Schmidt et al., 2023).
 Concerns about a reduction in user numbers and engagement may also be less relevant for datasets containing unique data.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-boyd_cohort_2013}{}}%
+Boyd, A., Golding, J., Macleod, J., Lawlor, D.
+A., Fraser, A., Henderson, J., \ldots{} Davey Smith, G.
+(2013).
+Cohort profile: The {``children of the 90s''}---the index offspring of the {Avon} {Longitudinal} {Study} of {Parents} and {Children}.
+\emph{International Journal of Epidemiology}, \emph{42}(1), 111--127.
+
 \leavevmode\vadjust pre{\hypertarget{ref-camerer_evaluating_2016}{}}%
@@
 \url{https://psysciacc.org/2020/09/14/incentivizing-discovery-through-the-psa001-secondary-analysis-challenge/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-fraser_cohort_2013}{}}%
+Fraser, A., Macdonald-Wallis, C., Tilling, K., Boyd, A., Golding, J., Davey Smith, G., \ldots{} Ness, A.
+(2013).
+Cohort profile: The {Avon} {Longitudinal} {Study} of {Parents} and {Children}: {ALSPAC} mothers cohort.
+\emph{International Journal of Epidemiology}, \emph{42}(1), 97--110.
 
@@
 The results from all participants are presented in Supplementary Material D.
-However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed 107 times and partially completed 27 times, leading to a response rate of 10% for complete surveys and 12% for incomplete surveys.”
+However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed 103 times and partially completed 20 times, leading to a response rate of 9% for complete surveys and 11% for at least partially complete surveys.”
   \item The preregistration and survey use the term “typical preregistration”. We changed this to “conventional preregistration” in the manuscript because we believe it is the more appropriate term.
@@
   \item We did not preregister which results we would present in the abstract.
-We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 25\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
+We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 26\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figs/numStudiesPlot-1}
-
-}
-
-\caption{\textbf{Figure C1.} Responses to survey question about number of relevant publications}\label{fig:numStudiesPlot}
-\end{figure}
-
-{\smaller[1]
-\begin{singlespace}
-
-\end{singlespace}
-}
-
 \begin{table}[H]
@@
 
+{\centering \includegraphics[width=1\linewidth]{figs/numStudiesPlot-1}
+
+}
+
+\caption{\textbf{Figure C1. Responses to survey question about number of relevant publications.}}\label{fig:numStudiesPlot}
+\end{figure}
+
+{\smaller[1]
+\begin{singlespace}
+
+\end{singlespace}
+}
+
+\begin{figure}
+
 {\centering \includegraphics[width=1\linewidth]{figs/concernedPlot-1}
@@
 
-\caption{\textbf{Figure C2. Responses to survey question about number of relevant publications}}\label{fig:concernedPlot}
+\caption{\textbf{Figure C2. Responses to survey question about number of relevant publications.} We excluded missing responses from this figure (\emph{n} = 3).}\label{fig:concernedPlot}
 \end{figure}
@@
 
-Respondents had published a median of 10 (IQR 2 to 25) studies using preexisting observational data.
+Respondents had published a median of 10 (IQR 2 to 25) studies using pre-existing observational data.
 They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) \footnote[1]{Participants could select multiple responses to this survey question.}.
-61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/101) reported being less concerned.
+61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses pre-existing observational data; 6\% (6/101) reported being less concerned.
 
@@
 
-Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (71\%; 87/123) and reproducible (77\%; 93/121) (Figure D2, top panel).
-At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 75/107) and \emph{more} reproducible (69\%; 74/108) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+Most respondents agreed that studies that analyze pre-existing observational datasets are trustworthy (71\%; 87/123) and reproducible (77\%; 93/121) (Supplementary Figure D1, top panel).
+At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 75/107) and \emph{more} reproducible (69\%; 74/108) compared to a typical study using pre-existing observational data (Supplementary Figure D1, bottom panel).
 
@@
 \caption{\textbf{Figure D1.
-Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+Responses to the survey questions on trustworthiness and reproducibility of observational research with pre-existing data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
 That they merit trust''.
@@
 The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
-The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}.
+The bar charts in the top panel excluded missing responses (n = 0; 2) and had no selection of the option \emph{``I don't understand the question''}.
 The bar charts in the bottom panel excluded missing responses (n = 12; 12) and responses of \emph{``I don't understand the question''} (n = 4; 3).}\label{fig:typicallyEcawPlotAll}
@@
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (33\%; 41/123), or sometimes (21\%; 26/103) (Figure 2D).
+Over half of respondents reported that their studies using pre-existing observational data are preregistered never or almost never (33\%; 41/123), or sometimes (21\%; 26/103) (Supplementary Figure 2D).
 About half reported sharing their analysis scripts never or almost never (19\%; 23/123), or sometimes (29\%; 36/123).
 69\% (85/123) reported that they never or almost never blind the data analyst.
-Almost all respondents answered that they use both exploratory (87\%; 107/123) and confirmatory (82\%; 101/123) analyses at least sometimes.
+Almost all respondents answered that they use both confirmatory (82\%; 101/123) and exploratory (87\%; 107/123) analyses at least sometimes.
 
@@
 
-25\% (26/106) of respondents agreed (versus 44\%; 47/106 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3D).
+25\% (26/106) of respondents agreed (versus 44\%; 47/106 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Supplementary Figure 3D).
 51\% (50/98) agreed (21\%; 21/98 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
 54\% (54/100) agreed (11\%; 11/100 disagreed) that ALSPAC should run a study on ECAWs.
-45\% (45/100) agreed (22\%; 22/100 disagreed) that they would prefer using an ECAW than using typical preregistration.
+45\% (45/100) agreed (22\%; 22/100 disagreed) that they would prefer using an ECAW than using conventional preregistration.
 
@@
 \caption{\textbf{Figure D3.
-Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 14; 14; 14; 14; respectively from top to bottom), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 5; 2; 1), and responses of \emph{``Unsure''} (\emph{n} = 3; 6; 7; 8).
+Responses to survey questions about using ECAWs.} These bar charts exclude missing values (\emph{n} = 14; 14; 14; 14), responses of \emph{``I don't understand the question''} (\emph{n} = 0; 5; 2; 1), and responses of \emph{``Unsure''} (\emph{n} = 3; 6; 7; 8).
 Agreement with the first question may be slightly inflated due to the format of the questions in this block.
 Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions.
-Four respondents agreed with all four statements, suggesting they may have glazed over the word ``less'' in the first question \protect\footnotemark[2].
-Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term ``study'' \protect\footnotemark[3].}\label{fig:alspacPlotAll}
+Four respondents agreed with all four statements, suggesting they may have glazed over the word ``less'' in the first question.
+Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term ``study''.}\label{fig:alspacPlotAll}
 \end{figure}
@@
 }
-
-\protect\footnotetext[2]{Another four respondents agreed with at least the first and second question, which appear contradictory.
-We did not preregister these considerations.
-More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
-\protect\footnotetext[3]{We intended for these questions to ask about ALSPAC running a trial on ECAWs.
-However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}
 

### 2023-04-28 manuscript/supplementary_materials.Rmd
changes

@@
 The results from all participants are presented in Supplementary Material D.
-However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed 107 times and partially completed 27 times, leading to a response rate of 10% for complete surveys and 12% for incomplete surveys.”
+However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed `r nrow(processed)` times and partially completed `r nrow(processed_all) - nrow(processed)` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(processed_all) / 1094 * 100)`% for at least partially complete surveys.”
   \item The preregistration and survey use the term “typical preregistration”. We changed this to “conventional preregistration” in the manuscript because we believe it is the more appropriate term.
@@
   \item We did not preregister which results we would present in the abstract.
-We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 25\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
+We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only `r pull(filter(study, support == "negative"), percentage)`\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as `r pull(filter(willing, support == "positive"), percentage)`\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”
 
@@
 
+```{r languageTable}
+# Programming languages table
+# We use the number of all participants who responded to the questionnaire as the denominator
+denominator <- nrow(processed)
+
+processed %>% 
+  rename(lang = programming_language) %>% 
+  separate_rows(lang, sep = ",") %>% 
+  count(lang) %>% 
+  replace_na(list(lang = "Missing")) %>% 
+  # Removing irrelevant
+  filter(lang != "irrelevant") %>% 
+  ungroup() %>% 
+  mutate(
+    `Percentage of respondents` = round(n / denominator * 100, 0)
+  ) %>% 
+  arrange(desc(n)) %>% 
+  rename(
+    `Programming language` = lang,
+    `N` = n
+  ) %>% 
+  kbl(
+  .,
+  format = "latex",
+  col.names = c("Programming language", "N", "Percentage of respondents"),
+  booktabs = T,
+  escape = F,
+  caption = "(ref:table1c-caption)",
+  centering = T,
+  position = "H") %>%
+  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
+  row_spec(row = 0, bold = TRUE) %>%
+  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
+```
+
+```{=latex}
+{\smaller[1]
+\begin{singlespace}
+```
+(ref:table1c-caption) __Table C1. “What programming language or software do you use for your analyses of preexisting observational data? (you may select multiple answers)”__
+```{=latex}
+\end{singlespace}
+}
+```
+
 ```{r numStudiesPlot, fig.align="center", fig.cap="(ref:numStudiesPlotCaption)", out.width="100%", fig.width=6, fig.height=3.2, fig.path='figs/', dev=c('png', 'pdf')}
@@
 ```
-(ref:numStudiesPlotCaption) __Figure C1.__ Responses to survey question about number of relevant publications
-```{=latex}
-\end{singlespace}
-}
-```
-
-```{r languageTable}
-# Programming languages table
-# We use the number of all participants who responded to the questionnaire as the denominator
-denominator <- nrow(processed)
-
-processed %>% 
-  rename(lang = programming_language) %>% 
-  separate_rows(lang, sep = ",") %>% 
-  count(lang) %>% 
-  replace_na(list(lang = "Missing")) %>% 
-  # Removing irrelevant
-  filter(lang != "irrelevant") %>% 
-  ungroup() %>% 
-  mutate(
-    `Percentage of respondents` = round(n / denominator * 100, 0)
-  ) %>% 
-  arrange(desc(n)) %>% 
-  rename(
-    `Programming language` = lang,
-    `N` = n
-  ) %>% 
-  kbl(
-  .,
-  format = "latex",
-  col.names = c("Programming language", "N", "Percentage of respondents"),
-  booktabs = T,
-  escape = F,
-  caption = "(ref:table1c-caption)",
-  centering = T,
-  position = "H") %>%
-  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
-  row_spec(row = 0, bold = TRUE) %>%
-  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
-```
-
-```{=latex}
-{\smaller[1]
-\begin{singlespace}
-```
-(ref:table1c-caption) __Table C1. “What programming language or software do you use for your analyses of preexisting observational data? (you may select multiple answers)”__
+(ref:numStudiesPlotCaption) __Figure C1. Responses to survey question about number of relevant publications.__
 ```{=latex}
@@
 # Create missing values text
-concerned_missing_text <- 
+concerned_missing_n <- 
   concerned_plot_data %>% 
@@
   summarise(
-    n = n(),
-    missing_text = paste0("missing = ", n)
+    n = n()
   ) %>% 
-  pull(missing_text)
+  pull(n)
 
@@
 ```
-(ref:concernedPlotCaption) __Figure C2. Responses to survey question about number of relevant publications__
+(ref:concernedPlotCaption) __Figure C2. Responses to survey question about number of relevant publications.__ We excluded missing responses from this figure (_n_ = `r concerned_missing_n`).
 ```{=latex}
@@
 
-Respondents had published a median of `r median(processed_all$n_studies, na.rm = TRUE)` (IQR `r round(quantile(processed_all$n_studies, .25, na.rm = TRUE))` to `r round(quantile(processed_all$n_studies, .75, na.rm = TRUE))`) studies using preexisting observational data.
+Respondents had published a median of `r median(processed_all$n_studies, na.rm = TRUE)` (IQR `r round(quantile(processed_all$n_studies, .25, na.rm = TRUE))` to `r round(quantile(processed_all$n_studies, .75, na.rm = TRUE))`) studies using pre-existing observational data.
 They reported using the following programming languages or software packages: `r language_text_all` \footnote[1]{Participants could select multiple responses to this survey question.}.
-`r pull(filter(concerned_all, support == "positive"), percentage)`% (`r pull(filter(concerned_all, support == "positive"), n)`/`r pull(filter(concerned_all, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; `r pull(filter(concerned_all, support == "negative"), percentage)`% (`r pull(filter(concerned_all, support == "negative"), n)`/`r pull(filter(concerned_all, support == "negative"), n_sum)`) reported being less concerned.
+`r pull(filter(concerned_all, support == "positive"), percentage)`% (`r pull(filter(concerned_all, support == "positive"), n)`/`r pull(filter(concerned_all, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses pre-existing observational data; `r pull(filter(concerned_all, support == "negative"), percentage)`% (`r pull(filter(concerned_all, support == "negative"), n)`/`r pull(filter(concerned_all, support == "negative"), n_sum)`) reported being less concerned.
 
@@
 
-Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy_all, support == "positive"), n)`/`r pull(filter(typically_trustworthy_all, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible_all, support == "positive"), n)`/`r pull(filter(typically_reproducible_all, support == "positive"), n_sum)`) (Figure D2, top panel).
-At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy_all, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy_all, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible_all, support == "positive"), n)`/`r pull(filter(ecaw_reproducible_all, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).
+Most respondents agreed that studies that analyze pre-existing observational datasets are trustworthy (`r pull(filter(typically_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy_all, support == "positive"), n)`/`r pull(filter(typically_trustworthy_all, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible_all, support == "positive"), n)`/`r pull(filter(typically_reproducible_all, support == "positive"), n_sum)`) (Supplementary Figure D1, top panel).
+At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy_all, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy_all, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible_all, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible_all, support == "positive"), n)`/`r pull(filter(ecaw_reproducible_all, support == "positive"), n_sum)`) compared to a typical study using pre-existing observational data (Supplementary Figure D1, bottom panel).
 
@@
 # Check if there is any missing and don't understand responses separately
-# processed %>%
+# processed_all %>%
 #   select(contains("typically_")) %>%
-#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
-#   nrow()
+#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question")
+
+# Calculate missing responses
+typically_missing_all <-
+  processed_all %>%
+  select(contains("typically_")) %>%
+  pivot_longer(
+    cols = everything(),
+    names_to = "item",
+    values_to = "value"
+  ) %>%
+  dplyr::filter(is.na(value) | value == "I don't understand the question") %>%
+  mutate(item = str_to_title(str_remove(item, "^[^_]*_"))) %>%
+  replace_na(list(value = "Missing")) %>%
+  group_by(item, value) %>%
+  summarise(
+    n = n()
+  )
 
@@
 (ref:typicallyEcawPlotAllCaption) __Figure D1.
-Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
+Responses to the survey questions on trustworthiness and reproducibility of observational research with pre-existing data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate.
 That they merit trust”.
@@
 The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green.
-The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_.
+The bar charts in the top panel excluded missing responses (n = 0; `r pull(filter(typically_missing_all, item == "Reproducible" & value == "Missing"), n)`) and had no selection of the option _“I don’t understand the question”_.
 The bar charts in the bottom panel excluded missing responses (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "Missing"), n)`) and responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing_all, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing_all, item == "Reproducible" & value == "I don't understand the question"), n)`).
@@
 
-Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2D).
+Over half of respondents reported that their studies using pre-existing observational data are preregistered never or almost never (`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered_all, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered_all, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Supplementary Figure 2D).
 About half reported sharing their analysis scripts never or almost never (`r filter(script_all, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script_all, method_script == "Never or almost never") %>% pull(n)`/`r filter(script_all, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script_all, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script_all, method_script == "Sometimes") %>% pull(n)`/`r filter(script_all, method_script == "Sometimes") %>% pull(n_sum)`).
 `r filter(blind_all, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind_all, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind_all, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst.
-Almost all respondents answered that they use both exploratory (`r pull(exploratory_all, percentage)`%; `r pull(exploratory_all, n)`/`r pull(exploratory_all, n_sum)`) and confirmatory (`r pull(confirmatory_all, percentage)`%; `r pull(confirmatory_all, n)`/`r pull(confirmatory_all, n_sum)`) analyses at least sometimes.
+Almost all respondents answered that they use both confirmatory (`r pull(confirmatory_all, percentage)`%; `r pull(confirmatory_all, n)`/`r pull(confirmatory_all, n_sum)`) and exploratory (`r pull(exploratory_all, percentage)`%; `r pull(exploratory_all, n)`/`r pull(exploratory_all, n_sum)`) analyses at least sometimes.
 
@@
 
-`r pull(filter(willing_all, support == "positive"), percentage)`% (`r pull(filter(willing_all, support == "positive"), n)`/`r pull(filter(willing_all, support == "positive"), n_sum)`) of respondents agreed (versus `r pull(filter(willing_all, support == "negative"), percentage)`%; `r pull(filter(willing_all, support == "negative"), n)`/`r pull(filter(willing_all, support == "negative"), n_sum)` who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3D).
+`r pull(filter(willing_all, support == "positive"), percentage)`% (`r pull(filter(willing_all, support == "positive"), n)`/`r pull(filter(willing_all, support == "positive"), n_sum)`) of respondents agreed (versus `r pull(filter(willing_all, support == "negative"), percentage)`%; `r pull(filter(willing_all, support == "negative"), n)`/`r pull(filter(willing_all, support == "negative"), n_sum)` who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Supplementary Figure 3D).
 `r pull(filter(opt_all, support == "positive"), percentage)`% (`r pull(filter(opt_all, support == "positive"), n)`/`r pull(filter(opt_all, support == "positive"), n_sum)`) agreed (`r pull(filter(opt_all, support == "negative"), percentage)`%; `r pull(filter(opt_all, support == "negative"), n)`/`r pull(filter(opt_all, support == "negative"), n_sum)` disagreed) that they would opt-in if ALSPAC ran a study on ECAWs.
 `r pull(filter(study_all, support == "positive"), percentage)`% (`r pull(filter(study_all, support == "positive"), n)`/`r pull(filter(study_all, support == "positive"), n_sum)`) agreed (`r pull(filter(study_all, support == "negative"), percentage)`%; `r pull(filter(study_all, support == "negative"), n)`/`r pull(filter(study_all, support == "negative"), n_sum)` disagreed) that ALSPAC should run a study on ECAWs.
-`r pull(filter(prefer_all, support == "positive"), percentage)`% (`r pull(filter(prefer_all, support == "positive"), n)`/`r pull(filter(prefer_all, support == "positive"), n_sum)`) agreed (`r pull(filter(prefer, support == "negative"), percentage)`%; `r pull(filter(prefer_all, support == "negative"), n)`/`r pull(filter(prefer_all, support == "negative"), n_sum)` disagreed) that they would prefer using an ECAW than using typical preregistration.
+`r pull(filter(prefer_all, support == "positive"), percentage)`% (`r pull(filter(prefer_all, support == "positive"), n)`/`r pull(filter(prefer_all, support == "positive"), n_sum)`) agreed (`r pull(filter(prefer, support == "negative"), percentage)`%; `r pull(filter(prefer_all, support == "negative"), n)`/`r pull(filter(prefer_all, support == "negative"), n_sum)` disagreed) that they would prefer using an ECAW than using conventional preregistration.
 
@@
 (ref:alspacPlotAllCaption) __Figure D3.
-Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Missing"), n)`; respectively from top to bottom), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Unsure"), n)`).
+Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Missing"), n)`), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing_all, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing_all, item == "Prefer_ecaw" & value == "Unsure"), n)`).
 Agreement with the first question may be slightly inflated due to the format of the questions in this block.
 Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions.
-Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question \protect\footnotemark[2].
-Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[3].
+Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question.
+Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study”.
 ```{=latex}
@@
 ```
-
-\protect\footnotetext[2]{Another four respondents agreed with at least the first and second question, which appear contradictory.
-We did not preregister these considerations.
-More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
-\protect\footnotetext[3]{We intended for these questions to ask about ALSPAC running a trial on ECAWs.
-However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}
 
